---
title: "Cognitive and Resting-State EEG Predictors of Intensive Multilingual Learning, and Change in Executive Function across Training: A Bayesian Multilevel Study"
shorttitle: "Predictors and Plasticity in Multilingual Learning"
author:
  - name: "[Author One]"
    corresponding: true
    email: "author.one@example.org"
    affiliations:
      - name: "[Institution]"
  - name: "[Author Two]"
    affiliations:
      - name: "[Institution]"
abstract: |
  Adults vary widely in how quickly and how well they learn a new language. We related
  baseline executive function (working memory, inhibitory control), statistical learning
  and resting-state EEG to the trajectory of grammaticality-judgement accuracy as
  Norwegian--English bilinguals acquired one of two artificial languages. We also tested
  change in executive function across training. Trial-level accuracy was modelled in a
  joint Bayesian multilevel framework with no per-learner derived slopes. Working memory
  predicted the level of grammatical knowledge learners reached, credibly so for the
  fully sampled gender-agreement property. Inhibitory control, statistical learning and
  the resting-state measures were largely uninformative, and no neural measure added
  predictive information beyond the behavioural ones. Working-memory span itself rose
  between the first and fifth sessions, so the capacity that predicted learning also
  changed over the training period. Attainment was estimated more reliably than learning
  rate, which bears on how slope-indexed aptitude findings should be read.
keywords: [second language acquisition, working memory, statistical learning, resting-state EEG, individual alpha frequency, Bayesian multilevel models, individual differences]
bibliography: references.bib
floatsintext: true
format:
  apaquarto-pdf:
    documentmode: man
    include-in-header:
      - text: |
          % apaquarto's template decrements the table counter at the end of every
          % longtable that carries no caption of its own (apatemplate.tex, "Disable
          % longtable counter"). Quarto now places the caption on an enclosing table
          % float and marks the longtable \def\LTcaptype{none}, so no longtable
          % carries its own caption any more and that decrement cancels the increment
          % the \caption just made. Without this, every table prints as "Table 1" and
          % every cross-reference to a table resolves to 1 -- in the PDF only; the
          % HTML and DOCX outputs number correctly. Restore the lost increment.
          \makeatletter
          \AtBeginDocument{%
            \apptocmd{\endlongtable}{\ifLTpatch@caption\else\addtocounter{table}{1}\fi}{}{}%
          }
          \makeatother

          % --- Vertical rhythm -------------------------------------------------
          % Defaults left captions crowded against their table or figure, opened a
          % gap above and below every display equation that broke the flow of the
          % surrounding sentence, and put more space under a heading than over it,
          % which detaches the heading from the text it introduces.
          % caption.sty is already loaded by the apa7 class, so only its settings are
          % adjusted here; loading it again risks an option clash.
          \captionsetup{aboveskip=10pt, belowskip=10pt}
          \setlength{\abovedisplayskip}{6pt plus 2pt minus 2pt}
          \setlength{\belowdisplayskip}{6pt plus 2pt minus 2pt}
          \setlength{\abovedisplayshortskip}{3pt plus 1pt minus 1pt}
          \setlength{\belowdisplayshortskip}{3pt plus 1pt minus 1pt}
          % Heading spacing is left to the apa7 class. The titlesec package cannot parse
          % that class's redefined heading levels: it fails at \ttl@extract\subparagraph
          % with 'Argument of \subparagraph has an extra }', and the document does not
          % build with it loaded. The spacing the class applies is APA's own, which is
          % what a journal expects, and the roomy look in manuscript mode comes from the
          % required double spacing rather than from the heading skips.

          % --- Keep wide tables inside the text block --------------------------
          % A table whose natural width exceeds the measure silently runs into the
          % margin rather than being scaled; column headers then collide (the
          % convergence table printed "DivergencesPassed" and ran the longest model
          % name into the next column). Shrink any over-wide table to the measure.
          \usepackage{etoolbox}
          \usepackage{graphicx}
          \AtBeginEnvironment{longtable}{\small}
          \makeatletter
          \g@addto@macro\@floatboxreset{\centering}
          \makeatother
  apaquarto-html: default
  apaquarto-docx:
    fig-dpi: 600
execute:
  echo: false
  warning: false
  message: false
editor: source
---

```{r setup}
#| include: false
# =============================================================================
# Manuscript setup -- shared paths, results loading, inline-reporting helpers.
# Render-anytime: before the HPC models exist it prints clearly-marked placeholders;
# once results/_pooled_*.csv are present every number is populated from those
# artefacts (Marwick, Boettiger & Mullen, 2018, https://doi.org/10.1080/00031305.2017.1375986).
# =============================================================================
# Figure text size, set once and inherited by every figure through les_theme(). Each
# figure is authored at the width of the text block, so nothing is scaled on placement and
# a point here is a point on the page. While several figures were authored at 8 inches and
# placed in a 6.5-inch measure, the 9 pt previously set here arrived as roughly 7.3 pt,
# below what is legible in print.
LES_FIG_BASE_SIZE <- 10

suppressPackageStartupMessages({
  library(here)
  source(here::here("_shared", "R", "00_paths.R"))
  source(here::here("_shared", "R", "05_funnel.R"))   # les_funnel(), shared with Paper 1
  source(here::here("paper_2_plasticity", "scripts", "_config.R"))
  library(dplyr)
  library(ggplot2)
  # depictr supplies the figure theme and colour system, shared with the companion
  # paper so both carry one visual identity. Its default qualitative palette is the
  # colourblind-safe Okabe-Ito set. Optional, per the render-anytime design.
  has_depictr <- requireNamespace("depictr", quietly = TRUE)
  if (has_depictr) {
    library(depictr)
    depictr_options(base_size = LES_FIG_BASE_SIZE)
  }
})

# Figure theme: depictr when present, base ggplot2 otherwise. Legend entries get room
# between them; the default packs the factor levels tight enough that the key squares
# read as a single block.
les_theme <- function(base_size = LES_FIG_BASE_SIZE) {
  base <- if (has_depictr) depictr::theme_depictr(base_size = base_size)
          else ggplot2::theme_minimal(base_size = base_size)
  base + ggplot2::theme(
    legend.key.spacing.y = ggplot2::unit(3, "pt"),
    legend.key.spacing.x = ggplot2::unit(8, "pt"),
    legend.text  = ggplot2::element_text(margin = ggplot2::margin(r = 6)),
    legend.title = ggplot2::element_text(margin = ggplot2::margin(b = 4)))
}

# Endpoints and midpoint of the diverging scale for signed quantities (the predictor
# correlation matrix here; the scalp topography in the companion paper). Taken from
# depictr so both papers share one diverging scale.
les_div <- if (has_depictr) depictr::depictr_palette(5, "diverging")[c(5, 3, 1)] else
  c("#005B96", "#F7F7F7", "#B2182B")

# --- Collinearity among the Part A baseline predictors -------------------------
# The band powers are entered simultaneously, so each coefficient is a partial effect.
# Variance inflation factors quantify how much that costs in precision, and are computed
# here from the actual predictor frame rather than transcribed. VIF_j = 1/(1 - R^2_j) from
# regressing predictor j on the others; sqrt(VIF) is the factor by which the coefficient's
# interval is widened relative to an orthogonal predictor.
# Computed lazily on first use and then cached, so this does not depend on being placed
# after the CSV loaders in the setup chunk.
.les_vif_cache <- NULL
.les_vif <- function() {
  if (!is.null(.les_vif_cache)) return(.les_vif_cache)
  f <- paper2_results("_predictor_values.csv")
  if (!file.exists(f)) return(NULL)
  pv <- utils::read.csv(f, stringsAsFactors = FALSE)
  w <- stats::reshape(pv, idvar = "participant_lab_ID", timevar = "predictor",
                      direction = "wide")
  names(w) <- sub("^value[.]", "", names(w))
  keep <- intersect(c("digit_span", "stroop_interference", "asrt_learning",
                      "delta", "theta", "alpha", "beta", "gamma", "iaf"), names(w))
  d <- w[, keep, drop = FALSE]
  d <- d[stats::complete.cases(d), , drop = FALSE]
  if (nrow(d) < length(keep) + 2) return(NULL)
  v <- vapply(names(d), function(p) {
    r2 <- summary(stats::lm(stats::as.formula(paste(p, "~ .")), data = d))$r.squared
    1 / (1 - r2)
  }, numeric(1))
  .les_vif_cache <<- v
  v
}
vif_of <- function(p) {
  v <- .les_vif()
  if (is.null(v) || is.na(v[p])) "NN" else formatC(v[[p]], format = "f", digits = 2)
}
vif_range <- function(ps) {
  v <- .les_vif(); if (is.null(v)) return("NN")
  v <- v[intersect(ps, names(v))]; if (!length(v)) return("NN")
  paste0(formatC(min(v), format = "f", digits = 1), " to ",
         formatC(max(v), format = "f", digits = 1))
}
vif_widen <- function(ps) {
  v <- .les_vif(); if (is.null(v)) return("NN")
  v <- sqrt(v[intersect(ps, names(v))]); if (!length(v)) return("NN")
  paste0(formatC(min(v), format = "f", digits = 1), " to ",
         formatC(max(v), format = "f", digits = 1), " times")
}

if (requireNamespace("papaja", quietly = TRUE)) {
  apa_num <- papaja::apa_num; apa_p <- papaja::apa_p
} else {
  apa_num <- function(x, digits = 2, ...) formatC(x, format = "f", digits = digits)
  apa_p   <- function(p, ...) ifelse(p < .001, "< .001", sub("^0", "", formatC(p, format = "f", digits = 3)))
}

.read_csv_if <- function(path) if (file.exists(path)) utils::read.csv(path, stringsAsFactors = FALSE) else NULL
post_tbl       <- .read_csv_if(paper2_results("_pooled_posterior_summaries.csv"))
conv_tbl       <- .read_csv_if(paper2_results("_pooled_convergence.csv"))
attainment_tbl <- .read_csv_if(paper2_results("_attainment.csv"))
reliability_tbl<- .read_csv_if(paper2_results("_reliability.csv"))
sens_tbl       <- .read_csv_if(paper2_results("_prior_sensitivity.csv"))
results_ready  <- !is.null(post_tbl)

# --- Raw-data descriptives shared with Paper 1 (participant flow + demographics,
# written by paper_1_transfer/scripts/00_extract_participants.R to both papers'
# results/) plus Paper-2 predictor/outcome descriptives (10_extract_descriptives.R).
# Injected inline so no count, demographic or task specification is transcribed.
flow_tbl        <- .read_csv_if(paper2_results("_sample_flow.csv"))
parts_tbl       <- .read_csv_if(paper2_results("_participants.csv"))
task_spec_tbl   <- .read_csv_if(paper2_results("_task_specs.csv"))
predictor_desc  <- .read_csv_if(paper2_results("_predictor_descriptives.csv"))
predictor_corr  <- .read_csv_if(paper2_results("_predictor_correlations.csv"))
psd_tbl         <- .read_csv_if(paper2_results("_resting_state_psd.csv"))
acc_traj_tbl    <- .read_csv_if(paper2_results("_accuracy_trajectory_descriptives.csv"))

# Resting-state descriptives (03/10). Carries the count of recordings in which the
# aperiodic decomposition resolves an alpha peak, which is what licenses or withholds
# trust in an individual-alpha-frequency value: the argmax estimator returns the lower
# bound of its search range for a spectrum that merely decays across it, so a value at
# that bound is not necessarily a measured peak. Injected rather than transcribed.
rseeg_desc <- .read_csv_if(paper2_results("_rseeg_descriptives.csv"))
rseeg_peak <- function(what = c("resolvable", "recorded"), condition_ = "eyes_closed") {
  what <- match.arg(what)
  if (is.null(rseeg_desc) ||
      !all(c("condition", "n_participants_with_resolvable_alpha_peak") %in% names(rseeg_desc)))
    return("NN")
  r <- rseeg_desc[rseeg_desc$condition == condition_ &
                    !is.na(rseeg_desc$n_participants_with_resolvable_alpha_peak), , drop = FALSE]
  if (nrow(r) < 1) return("NN")
  as.character(if (what == "resolvable")
    r$n_participants_with_resolvable_alpha_peak[1] else r$n[1])
}

# Raw pre/post means of the three cognitive indices in their own task units
# (10_extract_descriptives.R). Part B otherwise reports only standardised posteriors,
# which leaves a reader unable to see the size of the change on the measured scale.
cog_desc <- .read_csv_if(paper2_results("_cognitive_descriptives.csv"))
cog_mean <- function(index_, session_) {
  if (is.null(cog_desc)) return("NN")
  r <- cog_desc[cog_desc$index == index_ & cog_desc$session == session_, , drop = FALSE]
  if (nrow(r) != 1 || is.na(r$mean[1])) return("NN")
  apa_num(r$mean[1], digits = 1)
}

# Per-model, per-session participant and observation counts for the Part B pre/post
# models, written by the cluster summaries job. Until that artefact lands, the
# Participants sentence prints a clearly-marked pending phrase.
prepost_tbl <- .read_csv_if(paper2_results("_prepost_ns.csv"))
prepost_n <- function(model_, session_) {
  if (is.null(prepost_tbl)) return("NN")
  r <- prepost_tbl[prepost_tbl$model == model_ & prepost_tbl$session == session_, , drop = FALSE]
  if (nrow(r) != 1 || is.na(r$n_participants[1])) return("NN")
  as.character(r$n_participants[1])
}
prepost_ns_phrase <- function() {
  if (is.null(prepost_tbl)) return("*[per-model session counts pending the summaries run]*")
  sprintf(paste0("%s at Session 1 and %s at Session 5 for digit span, %s and %s for ",
                 "Stroop interference, and %s and %s for statistical learning"),
          prepost_n("p2_cognition_prepost_digit_span", "pre"),
          prepost_n("p2_cognition_prepost_digit_span", "post"),
          prepost_n("p2_cognition_prepost_stroop_interference", "pre"),
          prepost_n("p2_cognition_prepost_stroop_interference", "post"),
          prepost_n("p2_cognition_prepost_asrt_learning", "pre"),
          prepost_n("p2_cognition_prepost_asrt_learning", "post"))
}

# Gamma-band retained-power fractions under the 30 Hz low-pass, from
# scripts/12_extract_gamma_attenuation.R. Injected so the attenuation quantiles are
# computed from the stored per-participant artefact rather than transcribed; written
# defensively so the prose falls back to a marked approximation if the artefact is
# absent or its columns change.
gamma_att_tbl <- .read_csv_if(paper2_results("_gamma_attenuation.csv"))
gamma_att <- function(what = c("median", "q25", "q75"),
                      col = "retained_fraction") {
  what <- match.arg(what)
  if (is.null(gamma_att_tbl) || !col %in% names(gamma_att_tbl)) return(NA_character_)
  v <- gamma_att_tbl[[col]]
  if (!length(v) || all(is.na(v))) return(NA_character_)
  q <- stats::quantile(v, c(.25, .5, .75), na.rm = TRUE)
  val <- switch(what, q25 = q[[1]], median = q[[2]], q75 = q[[3]])
  # Plain "%" (not "\\%"): the value lands in markdown prose, where Pandoc escapes it.
  paste0(formatC(100 * val, format = "f", digits = 1), "%")
}
gamma_att_ready <- function() !is.na(gamma_att("median"))
gamma_att_clause <- function() {
  if (gamma_att_ready()) {
    sprintf(paste0("the retained power across that band has a median of %s of what an ",
                   "unfiltered recording would have carried (interquartile range %s to %s), ",
                   "and no more than a median of %s even under the most permissive ",
                   "single-pass reading of the zero-phase filter"),
            gamma_att("median"), gamma_att("q25"), gamma_att("q75"),
            gamma_att("median", col = "retained_fraction_singlepass"))
  } else {
    paste0("the retained power across that band has a median of about a third of what an ",
           "unfiltered recording would have carried (interquartile range roughly a quarter ",
           "to a half) *[approximate; pending the gamma-attenuation artefact]*")
  }
}

# Companion clause: the empirically observed 30-45 Hz power relative to the
# extrapolated aperiodic background. Values far above the filter's transmission mean
# the residual band is dominated by energy the low-pass did not remove in proportion
# (mains spill, EMG, the noise floor).
gamma_att_obs_clause <- function() {
  if (!gamma_att_ready() ||
      !"observed_over_extrapolated" %in% names(gamma_att_tbl)) {
    return("*[observed-band ratio pending the gamma-attenuation artefact]*")
  }
  sprintf("a median of %s of the extrapolated background, against the %s the filter would pass",
          gamma_att("median", col = "observed_over_extrapolated"),
          gamma_att("median"))
}

# Computational provenance recorded at fit time by _shared/R/04_provenance.R, so the
# reported versions are those of the run that produced these results rather than of the
# machine rendering the document. Absent until the pipeline is next run.
# Split-half (Spearman-Brown) internal consistency of the three baseline cognitive
# predictors, from scripts/11_extract_predictor_reliability.R.
predrel_tbl <- .read_csv_if(paper2_results("_predictor_reliability.csv"))
predrel <- function(p) {
  if (is.null(predrel_tbl)) return("NN")
  r <- predrel_tbl[predrel_tbl$predictor == p, , drop = FALSE]
  if (nrow(r) != 1 || is.na(r$reliability_sb[1])) return("NN")
  sub("^0", "", formatC(r$reliability_sb[1], format = "f", digits = 2))
}

.prov_tbl <- .read_csv_if(paper2_results("_provenance.csv"))
prov <- function(component, fallback = "*[version pending a pipeline run]*") {
  if (is.null(.prov_tbl)) return(fallback)
  r <- .prov_tbl[.prov_tbl$component == component, , drop = FALSE]
  if (nrow(r) != 1 || is.na(r$version[1])) return(fallback)
  as.character(r$version[1])
}

# --- Part A projection-predictive selection paths (one per reference model) and the
# aperiodic-vs-raw PSIS-LOO comparison. Solution paths come from 08_projpred_selection.R
# (grouped 10-fold, validate_search = TRUE); the LOO comparison from
# 09b_compare_aperiodic_commonsample.R (09_compare_aperiodic.R is superseded).
# Injected so the selection ranking, suggested size and comparison verdict are read from
# stored artefacts rather than transcribed.
projpath_raw  <- .read_csv_if(paper2_results("p2_predictive_trajectory_projpred_path.csv"))
projpath_aper <- .read_csv_if(paper2_results("p2_predictive_trajectory_aperiodic_projpred_path.csv"))
projpath_gender <- .read_csv_if(paper2_results("p2_predictive_trajectory_gender_projpred_path.csv"))
aper_loo_tbl  <- .read_csv_if(paper2_results("_aperiodic_loo_compare.csv"))

# Which strata actually have a selection path on disk. The paper reports a gender-only
# reference model alongside the pooled one, so a reader is entitled to know whether the
# selection was run on both or only on the pooled stratum. Reading it from the artefacts
# rather than asserting it means the sentence becomes correct on its own once the missing
# stratum is fitted, instead of having to be remembered and edited.
projpred_strata <- function() {
  have <- c(pooled = !is.null(projpath_raw), gender = !is.null(projpath_gender))
  if (all(have)) return("both the pooled and the gender-agreement reference models")
  if (have[["pooled"]]) return("the pooled reference model only")
  if (have[["gender"]]) return("the gender-agreement reference model only")
  "neither reference model"
}
projpred_gender_caveat <- function() {
  if (!is.null(projpath_gender)) return("")
  paste0(" The gender-agreement stratum, which carries this paper's headline",
         " individual-difference claim, has not yet been through the same selection, so",
         " the ranking below should not be read as established for it.")
}

# The gender-stratum result, written only once that selection exists. It matters
# separately from the pooled one for two reasons: it is the stratum the paper's headline
# individual-difference claim rests on, and its search ran to the full candidate set, so
# the terms the pooled search never reached were actually evaluated there.
projpred_gender_result <- function() {
  if (is.null(projpath_gender)) return("")
  reached <- function(term) any(grepl(term, projpath_gender$predictor, fixed = TRUE))
  tail_terms <- reached("z_gamma") && reached("z_iaf")
  out <- sprintf(paste0(
    "The gender-agreement stratum, fitted to that property alone, gives the same answer:",
    " a suggested size of %s, with working memory again the first individual-difference",
    " predictor to enter, at step %s, and inhibitory control and statistical learning",
    " entering beyond it (steps %s and %s)."),
    proj_size(projpath_gender), proj_rank(projpath_gender, "z_digit_span"),
    proj_rank(projpath_gender, "z_stroop"), proj_rank(projpath_gender, "z_asrt"))
  if (tail_terms) {
    out <- paste0(out, sprintf(paste0(
      " That search ran over the full candidate set rather than projpred's default path",
      " length, so gamma power and the individual alpha frequency were evaluated and not",
      " truncated away, and they rank last of all (steps %s and %s). The full-length",
      " search was run only for this stratum, the one the individual-difference claim",
      " rests on. There, neither term added anything the retained predictors had missed,",
      " and how far beyond the suggested size a term lands is not interpreted."),
      proj_rank(projpath_gender, "z_gamma"), proj_rank(projpath_gender, "z_iaf")))
  }
  out
}

flow_n <- function(stage, col = "n_total") {
  if (is.null(flow_tbl)) return("NN")
  r <- flow_tbl[flow_tbl$stage == stage, , drop = FALSE]
  if (nrow(r) != 1 || is.na(r[[col]])) return("NN")
  as.character(r[[col]])
}
pstat <- function(metric, col = "value", digits = 1) {
  if (is.null(parts_tbl)) return("NN")
  r <- parts_tbl[parts_tbl$metric == metric, , drop = FALSE]
  if (nrow(r) != 1 || is.na(r[[col]])) return("NN")
  v <- r[[col]]
  if (is.numeric(v) && !isTRUE(all.equal(v, round(v)))) apa_num(v, digits = digits) else as.character(v)
}
# Task specification value (e.g. ASRT blocks / trials-per-block) from _task_specs.csv.
task_spec <- function(task_, spec_) {
  if (is.null(task_spec_tbl)) return("NN")
  r <- task_spec_tbl[task_spec_tbl$task == task_ & task_spec_tbl$spec == spec_, , drop = FALSE]
  if (nrow(r) < 1 || is.na(r$value[1])) return("NN")
  as.character(r$value[1])
}
# By-participant reliability (.NN style) of an effect from _reliability.csv.
# (Named relv, not rel: the tbl-reliability chunk assigns a data.frame to `rel`.)
relv <- function(model_id, effect_) {
  if (is.null(reliability_tbl)) return("NN")
  r <- reliability_tbl[reliability_tbl$model == model_id & reliability_tbl$effect == effect_, , drop = FALSE]
  if (nrow(r) != 1 || is.na(r$reliability)) return("NN")
  sub("^0", "", formatC(r$reliability, format = "f", digits = 2))
}

# projpred suggested (parsimonious) size, and the forward-search step (the `size`
# column) at which a given predictor first enters the solution path.
proj_size <- function(tbl) if (is.null(tbl) || !nrow(tbl)) "NN" else as.character(tbl$suggested_size[1])
proj_n    <- function(tbl) if (is.null(tbl) || !nrow(tbl)) "NN" else as.character(tbl$n_participants[1])
proj_rank <- function(tbl, term) {
  if (is.null(tbl)) return("NN")
  w <- which(tbl$predictor == term)
  if (!length(w)) return("NN")
  as.character(tbl$size[w[1]])
}
# aperiodic-vs-raw PSIS-LOO within a stratum: the elpd difference (aperiodic - raw),
# its SE, or the preferred-model verdict. Reads _aperiodic_loo_compare.csv.
aper_loo <- function(stratum_, what = c("verdict", "elpd", "se")) {
  what <- match.arg(what)
  if (is.null(aper_loo_tbl)) return(if (what == "verdict") "*[pending LOO comparison]*" else "NN")
  r <- aper_loo_tbl[aper_loo_tbl$stratum == stratum_, , drop = FALSE]
  if (nrow(r) != 1) return("NN")
  switch(what,
    verdict = {
      p <- as.character(r$preferred_model)
      if (grepl("neither", p, ignore.case = TRUE))
        "neither parameterisation was clearly preferred (the difference was within twice its standard error)"
      else sprintf("the %s parameterisation was preferred", p)
    },
    elpd    = apa_num(r$elpd_diff_aperiodic_minus_raw, digits = 1),
    se      = apa_num(r$se_diff, digits = 1))
}
# TRUE if a stratum's |elpd_diff| <= 2*se (neither parameterisation clearly preferred).
aper_loo_tie <- function(stratum_) {
  if (is.null(aper_loo_tbl)) return(NA)
  r <- aper_loo_tbl[aper_loo_tbl$stratum == stratum_, , drop = FALSE]
  if (nrow(r) != 1) return(NA)
  abs(r$elpd_diff_aperiodic_minus_raw) <= 2 * r$se_diff
}
# Trials per participant in a stratum. This is what makes the trial-level LOO
# comparison uninformative about participant-level predictors: the more trials each
# participant contributes, the more completely their random intercept is pinned by
# their own remaining data when one trial is held out, and the less any
# between-participant predictor can add to predicting it.
aper_loo_tpp <- function(stratum_) {
  if (is.null(aper_loo_tbl)) return("NN")
  r <- aper_loo_tbl[aper_loo_tbl$stratum == stratum_, , drop = FALSE]
  if (nrow(r) != 1 || is.na(r$n_participants[1]) || r$n_participants[1] == 0) return("NN")
  as.character(round(r$n_obs[1] / r$n_participants[1]))
}

# --- Participant-grouped K-fold comparison (09c) ------------------------------
# The same two models scored by leaving out whole PARTICIPANTS rather than single trials,
# which is the criterion the question actually calls for. Its standard error is clustered
# on participants: loo_compare's own SE treats the pointwise differences as independent,
# and under grouped CV they are not, so it understates the uncertainty severalfold. The
# ratio is reported in the artefact and quoted in the text, because the corrected and
# uncorrected numbers support different conclusions and the reader is entitled to see that.
aper_kf_tbl <- .read_csv_if(paper2_results("_aperiodic_kfold_compare.csv"))
aper_kf_ready <- function() !is.null(aper_kf_tbl) && nrow(aper_kf_tbl) >= 2
aper_kf <- function(stratum_, what = c("elpd", "se", "se_naive", "inflation", "k")) {
  what <- match.arg(what)
  if (is.null(aper_kf_tbl)) return("NN")
  r <- aper_kf_tbl[aper_kf_tbl$stratum == stratum_, , drop = FALSE]
  if (nrow(r) != 1) return("NN")
  switch(what,
    elpd      = apa_num(r$elpd_diff_aperiodic_minus_raw, digits = 1),
    se        = apa_num(r$se_diff, digits = 1),
    se_naive  = apa_num(r$se_naive_independent, digits = 1),
    inflation = apa_num(r$se_inflation_from_clustering, digits = 1),
    k         = as.character(r$k[1]))
}
# TRUE only if BOTH strata are ties on the clustered SE, so the sentence below cannot
# describe a split verdict as a single one.
aper_kf_tie <- function() {
  if (!aper_kf_ready()) return(NA)
  all(abs(aper_kf_tbl$elpd_diff_aperiodic_minus_raw) <= 2 * aper_kf_tbl$se_diff)
}
# The sentence reporting the grouped comparison. Emitted only once the artefact exists, so
# the paragraph reads correctly before the job has run and does not have to be remembered
# and edited afterwards.
aper_kf_sentence <- function() {
  if (!aper_kf_ready()) {
    return(paste0(" A direct participant-grouped elpd comparison would complement it and",
                  " is pending."))
  }
  tie <- isTRUE(aper_kf_tie())
  paste0(
    sprintf(paste0(
      "\n\nWe then ran the participant-grouped comparison directly, refitting both models on",
      " the common sample under the %s-fold participant-blocked scheme described in the",
      " Method. The aperiodic-minus-raw difference was %s in the pooled stratum and %s for",
      " gender agreement."),
      aper_kf("pooled", "k"), aper_kf("pooled", "elpd"), aper_kf("gender", "elpd")),
    sprintf(paste0(
      " These are reported with standard errors clustered on participants (%s and %s"),
      aper_kf("pooled", "se"), aper_kf("gender", "se")),
    sprintf(paste0(
      "): the usual standard error treats the pointwise contributions as independent,",
      " which under participant-blocked folds they are not, and doing so understates the",
      " uncertainty here by factors of %s and %s."),
      aper_kf("pooled", "inflation"), aper_kf("gender", "inflation")),
    if (tie) {
      sprintf(paste0(
        " On that corrected scale, neither difference reaches twice its standard error,",
        " so neither characterisation predicts a held-out participant's trajectory",
        " detectably better than the other. Both differences favour the raw bands in",
        " direction, but neither is distinguishable from zero, and we draw no",
        " conclusion from the sign. The comparison is also blunt: at these clustered",
        " standard errors only a difference beyond about %s nats in the pooled stratum",
        " and %s for gender agreement would be read as decisive, so this is a tie that",
        " bounds the comparison rather than one that establishes equivalence. That is the",
        " same verdict the trial-level comparison returned, now reached on the criterion",
        " the question calls for."),
        apa_num(2 * abs(as.numeric(aper_kf_tbl$se_diff[aper_kf_tbl$stratum == "pooled"][1])), digits = 0),
        apa_num(2 * abs(as.numeric(aper_kf_tbl$se_diff[aper_kf_tbl$stratum == "gender"][1])), digits = 0))
    } else {
      # Report the per-stratum verdict in prose. Directing the reader to a CSV would put a
      # repository file path into the Results, and the verdict is a one-clause statement.
      .kf_verdict <- function(s) {
        d <- suppressWarnings(as.numeric(aper_kf_tbl$elpd_diff_aperiodic_minus_raw[
          aper_kf_tbl$stratum == s][1]))
        e <- suppressWarnings(as.numeric(aper_kf_tbl$se_diff[aper_kf_tbl$stratum == s][1]))
        if (is.na(d) || is.na(e)) return("could not be evaluated")
        if (abs(d) > 2 * e)
          sprintf("favours the %s characterisation", if (d > 0) "aperiodic" else "raw-band")
        else "leaves the two indistinguishable"
      }
      sprintf(paste0(" On that corrected scale, the comparison %s in the pooled stratum",
                     " and %s for gender agreement."),
              .kf_verdict("pooled"), .kf_verdict("gender"))
    })
}

# Median widening of credible intervals under weakly-informative vs informative priors.
prior_ci_widening <- function(tbl = sens_tbl) {
  if (is.null(tbl)) return("*[pending sensitivity run]*")
  r <- stats::median(tbl$ci_width_weak / tbl$ci_width_inf, na.rm = TRUE)
  paste0(round(100 * (r - 1)), "%")
}

# Prior-sensitivity inline reporter: an effect's posterior median and interval under
# the weakly-informative ("weak") or the informative ("inf") prior, read from
# _prior_sensitivity.csv rather than transcribed.
sens_effect <- function(model_id, term, which = c("weak", "inf")) {
  which <- match.arg(which)
  if (is.null(sens_tbl)) return("*[pending sensitivity run]*")
  r <- sens_tbl[sens_tbl$model == model_id & sens_tbl$parameter == term, , drop = FALSE]
  if (nrow(r) != 1) return("NN")
  med <- r[[paste0("median_", which)]]
  lo  <- r[[paste0("ci_low_", which)]]
  hi  <- r[[paste0("ci_high_", which)]]
  d <- .les_digits(med, lo, hi)
  sprintf("$b = %s$, 95%% CrI $[%s, %s]$",
          apa_num(med, digits = d), apa_num(lo, digits = d), apa_num(hi, digits = d))
}
# Largest weak-prior probability of direction among the nine predictor-by-session
# interactions, across both Part A models: the value that shows the attainment-not-rate
# asymmetry survives flat priors. Deliberately excludes the design terms
# (session-by-property), whose pd is 1 by construction.
sens_max_rate_pd <- function() {
  if (is.null(sens_tbl)) return("*[pending sensitivity run]*")
  rate_terms <- paste0("z_session_time:z_",
                       c("digit_span", "stroop", "asrt",
                         "delta", "theta", "alpha", "beta", "gamma", "iaf"))
  r <- sens_tbl[sens_tbl$parameter %in% rate_terms, , drop = FALSE]
  if (!nrow(r)) return("NN")
  sub("^0", "", formatC(max(r$pd_weak, na.rm = TRUE), format = "f", digits = 2))
}

.pretty_term <- function(x) {
  x |>
    gsub("^b_", "", x = _) |>
    gsub("z_session_time", "Session", x = _) |>
    gsub("z_digit_span", "Working memory", x = _) |>
    gsub("z_stroop", "Inhibitory control", x = _) |>
    gsub("z_asrt", "Statistical learning", x = _) |>
    gsub("z_iaf", "IAF", x = _) |>
    gsub("z_alpha", "Alpha", x = _) |> gsub("z_theta", "Theta", x = _) |>
    gsub("z_beta", "Beta", x = _) |> gsub("z_delta", "Delta", x = _) |> gsub("z_gamma", "Gamma", x = _) |>
    gsub("grammatical_property", "Property:", x = _) |>
    gsub(":", " x ", x = _)
}

# Decimal precision that keeps a credible interval visibly clear of zero. A bound
# can sit very close to zero without touching it (the working-memory predictor's
# lower bound is .005), and at a fixed two decimals such an interval prints as though
# it reached zero, contradicting the credibility being claimed for it. Widen only
# where that would happen, capped at four decimals; everything else stays at two.
# Kept identical to the companion paper's helper so both report on the same rule.
.les_digits <- function(med, lo, hi, min_d = 2L, max_d = 4L) {
  # The rule is simply that no non-zero quantity may print as zero. That matters most
  # for a credible interval, where a "0.00" bound contradicts the credibility claimed
  # for it in the text, but it applies equally to a median, which would otherwise
  # appear as the signed zero "-0.00". Two decimals is the floor and four the ceiling,
  # so effects large enough to be read at two decimals are unaffected and precision
  # widens only where the alternative is a misleading or malformed number.
  # Kept identical to the companion paper's helper so both report on the same rule.
  v <- abs(c(med, lo, hi))
  v <- v[is.finite(v) & v > 0]
  if (!length(v)) return(min_d)
  d <- min_d
  while (d < max_d && round(min(v), d) == 0) d <- d + 1L
  d
}

# Inline reporter: "b = m, 95% CrI [lo, hi], pd = p%"
report_effect <- function(model_id, term, tbl = post_tbl) {
  if (is.null(tbl)) return("*[pending model fit]*")
  row <- tbl[tbl$model == model_id & tbl$parameter == term, , drop = FALSE]
  if (nrow(row) != 1) return("*[term not found]*")
  d <- .les_digits(row$median, row$ci_low, row$ci_high)
  sprintf("$b = %s$, 95%% CrI $[%s, %s]$, $p_d = %s\\%%$",
          apa_num(row$median, digits = d), apa_num(row$ci_low, digits = d),
          apa_num(row$ci_high, digits = d), apa_num(100 * row$pd, digits = 1))
}

# Probability of direction alone, as an inline percentage: for the places where the
# prose refers back to an effect already reported in full, so the pd cannot silently
# go stale on a refit.
pd_of <- function(model_id, term, tbl = post_tbl) {
  if (is.null(tbl)) return("NN")
  row <- tbl[tbl$model == model_id & tbl$parameter == term, , drop = FALSE]
  if (nrow(row) != 1 || is.na(row$pd)) return("NN")
  apa_num(100 * row$pd, digits = 1)
}
```

Adults differ markedly in the speed and the ultimate success with which they learn a
new language, and explaining that variation is a central goal of the cognitive
neuroscience of language.
Three families of predictor recur in the literature. Working memory is the
best-evidenced cognitive correlate of second-language outcomes, with a small-to-moderate
positive meta-analytic association [@linck2014] that is stronger for executive and
complex-span measures than for simple storage span. In the phonological short-term
memory literature, associations with second-language vocabulary learning are likewise
stronger for nonword repetition than for digit span [@kurokawa2026]. Statistical or
implicit learning, the ability to extract distributional regularities, relates positively
but only weakly to language ability across 97 studies, with the size of the correlation
depending on the task and the stimulus modality [@zhou2024]. In one individual-differences
study, nonadjacent statistical learning predicted comprehension of particular
native-language sentence types [@misyak2012].
Inhibitory control is the least settled. The bilingual-advantage literature finds no reliable
executive-function benefit of bilingualism [@paap2013; @dick2019], and direct tests of
inhibition against second-language outcomes are mixed, reporting null [@huensch2024;
@linckweiss2015] as well as positive [@chen2022] associations.

Alpha and theta dynamics in the resting EEG, and the individual alpha frequency in
particular, index memory and cognitive performance [@klimesch1999]. Baseline oscillatory
power has also predicted the rate of adult second-language learning directly, with faster
learners distinguished mainly by higher-frequency (beta and low-gamma) rhythms
[@prat2016; @prat2019]. In the later of those studies, only the coherence measures survived
correction for multiple comparisons, so the band-power precedent rests chiefly on the
earlier one, together with a ten-participant pilot in which resting beta power predicted an
attainment-normalised progress score in older adults [@kliesch2021]. We therefore enter
five canonical bands (delta,
theta, alpha, beta, gamma) and the individual alpha frequency as baseline predictors.
That precedent implicates higher-frequency bands recorded over right temporoparietal cortex.
Our own reduction is over posterior electrodes, so we do not expect it to transfer directly,
and the bands enter without a directional prior (see *Priors*). Although intensive language learning can reshape the underlying neural
architecture [@li2014; @martensson2012], the EEG here is measured only at baseline and used
to predict subsequent learning.

Raw band power is, moreover, an ambiguous quantity. It conflates genuine oscillatory
activity with the broadband aperiodic ($1/f$) background on which the peaks sit, and the
two components carry different information [@donoghue2020]. The aperiodic exponent has
been proposed as a non-invasive proxy for cortical excitation--inhibition balance
[@gao2017], varies systematically across individuals and with age [@voytek2015], and can
drive associations that appear on the surface to belong to band power [@ouyang2020]. A
claim that a resting rhythm predicts learning should therefore survive the separation of
the two components, and we test that directly.

A question cuts across all three families: whether an aptitude expresses itself in how
fast learners improve or in the level they reach. Most individual-difference studies index learning by a per-learner slope estimated in a first stage, a design that biases the predictor effects (see *Learning outcome and model structure*). We therefore model rate and attainment jointly at trial level and ask
which of the two the predictors act on.

This study brings these strands together in a controlled, longitudinal
artificial-language paradigm. Norwegian--English bilinguals learn one of two miniature
languages across a series of laboratory sessions. At baseline, we measure executive
function (working memory via digit span, inhibitory control via the Stroop
task [@stroop1935]), resting-state EEG and statistical learning, the last by an alternating
serial reaction-time task [@howard1997], an alternating-sequence variant of the serial
reaction-time paradigm [@nissen1987]. Across the training sessions, we track the developing
accuracy of grammaticality
judgements. Designs in which a controlled training outcome is predicted from a baseline neural recording remain rare. Most evidence linking resting EEG to cognition is cross-sectional, and the few longitudinal precedents [@prat2016; @prat2019; @kliesch2021] indexed learning by a derived rate or gain score. The present study models the full trial-level trajectory jointly, separating attainment from rate and quantifying the reliability of each. It also asks the neural question twice, of the raw band powers and of their aperiodic-adjusted decomposition, so that which characterisation of the resting spectrum the data prefer is itself treated as an empirical question.

Two questions follow. Part A asks whether the baseline cognitive and neural
measures predict the rate and attainment of learning. Part B asks whether executive
function changes between the first and fifth sessions, and whether any change is confined
to one index or broad across the battery. Part A bears on a standing question in the
individual-differences literature: whether successful adult grammar learning is
underwritten by a general advantage or by a specific capacity. A general advantage should
be evident across a battery of cognitive and neural measures assessed and modelled
identically. A specific capacity would leave one predictor associated with attainment and
its co-measured companions unrelated to it. The two accounts are
separable here to the extent that every predictor enters the same joint model of the
same outcome under priors of comparable scale. Identical modelling does not, however,
confer comparable reliability on the measures themselves, and which comparators can bear
the contrast is set out under *Limitations and future directions*. Part B has a stake of its own. Whether
cognitive capacities improve with demanding training is itself contested, since apparent gains often turn out to reflect retest artefacts and expectancy [@boot2013; @simons2016]. The three indices are therefore co-administered and used as internal controls on one another, because generic retest and expectancy effects act broadly across a battery.

# Method

## Participants

The participants are those of the companion study: Norwegian--English bilinguals
(first language Norwegian, second language English) who learned one of two miniature
languages. One participant was aged 16 at enrolment according to the language-history
questionnaire, and all others were adults. The companion paper reports the full age
distribution and the ethical approval and consent procedures covering the sample. Of the
`r flow_n("enrolled")` enrolled, `r flow_n("usable_accuracy")`
contributed grammaticality-judgement trajectories
(`r flow_n("usable_accuracy", "n_mini_english")` Mini-English,
`r flow_n("usable_accuracy", "n_mini_norwegian")` Mini-Norwegian) and
`r flow_n("battery_baseline")` a baseline cognitive battery. Of these,
`r flow_n("joint_predictor")` had both measures and were therefore candidates for the
joint predictor model, and `r flow_n("usable_rseeg")` also had usable baseline
resting-state EEG. Because the
joint model requires complete data on every predictor, the fitted raw-band model
comprises `r proj_n(projpath_raw)` of those participants. The difference is that one
participant with usable resting-state EEG had a baseline Stroop score excluded under the
task's minimum-observation rule, and listwise deletion removes them. The aperiodic variant
described below comprises `r proj_n(projpath_aper)` participants, the subset for which
the specparam re-parameterisation is defined. That decomposition returns no individual
alpha frequency for a spectrum in which it fits no alpha peak, and listwise deletion then
removes those participants too. This loss is therefore not independent of the alpha
measures themselves. The Session-1-versus-5 pre/post models of
executive function use all participants contributing data at either session
(`r prepost_ns_phrase()`), with the session effect estimated within participants through
the by-participant intercept. Participants tested only once inform the intercept but
contribute no within-participant change. @fig-flow traces
attendance and the resulting analysis samples across the six sessions.

These are the samples the models are fitted to. The descriptive displays reported below
are computed on wider frames, each named in the caption that carries it, and
those counts are not interchangeable with these. A descriptive summary of the predictors
requires only that a participant have the measure in question, not complete data on every
predictor, and in places not even the usability judgement behind the
`r flow_n("usable_rseeg")` above. No descriptive frame in this paper is smaller than the
`r proj_n(projpath_raw)` participants the raw-band model is fitted to. Of the
`r pstat("n_lhq3", "value", 0)` participants with a Language History Questionnaire record
[LHQ3; @lhq3], the mean age was `r pstat("age_years")` years
($SD = `r pstat("age_years", "sd")`$, range `r pstat("age_years", "vmin", 0)`--`r pstat("age_years", "vmax", 0)`).
Of that group, `r pstat("sex_female", "value", 0)` were female,
`r pstat("sex_male", "value", 0)` male and
`r pstat("sex_nonbinary", "value", 0)` non-binary, and `r pstat("sex_undisclosed", "value", 0)`
did not disclose their gender. Most were right-handed
(`r pstat("hand_right", "value", 0)` of `r pstat("hand_reported_n", "value", 0)` who
reported handedness). English had been acquired early (age of first exposure
$M = `r pstat("eng_aoa_listen")`$ years) and used for many years, with high self-rated
proficiency ($M = `r pstat("eng_l2_proficiency", "value", 2)`$ on a 0--1 scale). The study was
conducted with institutional research-ethics approval (reference to be added) and written
informed consent from all participants.

```{r}
#| label: fig-flow
#| fig-cap: "Participant Flow"
#| apa-note: "Each stage spans the full enrolled cohort. The solid block is the participants still present and the faint block completing the bar is those no longer present, so cumulative attrition reads as filled against unfilled area. Tapering bands connect consecutive stages, each transition annotated with the number lost and the percentage retained. In the upper tier, the solid block is split by mini-language. The predictor-model stages are not recorded by language and are drawn in neutral grey. The upper tier tracks attendance across all six sessions, since this paper draws on all of them: the cognitive battery at Sessions 1 and 5, the resting-state recording at Session 2, and the learning trajectory across the four judgement sessions. The lower tier is the predictor-model sample proper, which is nested, and its final stage is the sample the raw-band model is fitted to, after listwise deletion on the full predictor set. Bands are drawn only within a tier, because the analysis samples pool across sessions and are not a subset of final-session attendance. The repository's flow artefacts reproduce and cross-check every count shown."
#| fig-width: 6.5
#| fig-height: 5.4
if (!is.null(flow_tbl)) {
  fl <- flow_tbl[!is.na(flow_tbl$n_total), ]
  gv <- function(st) { v <- fl$n_total[fl$stage == st]; if (length(v)) v[1] else NA_integer_ }
  att <- fl[grepl("^(enrolled|attended)", fl$stage), ]
  att <- att[order(att$order), ]
  d <- rbind(
    data.frame(label = sub("\\s*\\(Paper [12]\\)", "", att$label), n = att$n_total,
               n_a = att$n_mini_english, n_b = att$n_mini_norwegian,
               tier = "Enrolment and attendance", note = NA_character_,
               stringsAsFactors = FALSE),
    data.frame(
      label = c("Contributed a learning trajectory",
                "…and a usable baseline battery",
                "…and usable baseline resting-state EEG",
                "…and complete data on every predictor (fitted raw-band model)"),
      # The final stage is the model's own fitted sample, read from the projpred
      # solution path rather than from the flow artefact, which records candidacy
      # stages and not the listwise deletion the joint model applies. Without it the
      # funnel stopped one stage short of the sample every Part A estimate rests on.
      n     = c(gv("trajectory"), gv("joint_predictor"), gv("usable_rseeg"),
                suppressWarnings(as.integer(proj_n(projpath_raw)))),
      n_a = NA_integer_, n_b = NA_integer_,
      tier = "Predictor-model sample",
      note = c(NA_character_,
               sprintf("%s had a usable baseline battery in total", gv("battery_baseline")),
               NA_character_,
               sprintf("the aperiodic variant is fitted to %s", proj_n(projpath_aper))),
      stringsAsFactors = FALSE))
  les_funnel(d, base_size = LES_FIG_BASE_SIZE)
} else {
  cat("*Participant-flow diagram will appear once `_sample_flow.csv` is present.*")
}
```

## Cognitive and neural indices

The cognitive battery was administered at Sessions 1 (baseline) and 5 (post) as
browser-based tasks on the Gorilla experiment platform [@anwylirvine2020]. Records from
individuals who completed the online battery without enrolling in the study were excluded
at extraction, since they have no consent linkage to the participant key. Working memory
was measured by digit span (forward and backward, list lengths 3--8
with four items each, `r task_spec("digit_span", "scored_trials")` scored trials, the index
being the number correct). Inhibitory
control was measured by a four-colour Stroop task
(`r task_spec("stroop", "test_trials")` test trials,
`r task_spec("stroop", "congruent_trials")` of them congruent, the index being the
incongruent-minus-congruent mean reaction time). Higher Stroop scores
therefore index poorer inhibitory control, so a positive coefficient for this predictor
would mean poorer inhibition predicting higher accuracy. Statistical learning was
measured by an alternating serial reaction-time (ASRT) task [@howard1997], the
alternating-sequence variant of the serial reaction-time paradigm [@nissen1987], of
`r task_spec("asrt", "n_blocks")` blocks of `r task_spec("asrt", "trials_per_block")`
trials in which pattern and random trials alternate. The index is the
low-minus-high-probability-triplet mean reaction-time difference, with higher scores
indexing stronger learning. Reaction times were
trimmed to 50--5000 ms with within-cell outliers beyond three standard deviations removed.

Resting-state EEG (rs-EEG) was recorded once, at Session 2, with a 32-channel Brain Products LiveAmp
(500 Hz) in separate eyes-closed and eyes-open recordings of about
five minutes each, each delimited by its own onset and offset markers. Preprocessing in
BrainVision Analyzer followed the order: visual inspection of the continuous record,
re-referencing to the average of the mastoids (TP9, TP10), band-pass filtering with
zero-phase-shift Butterworth filters (high-pass 0.1 Hz, low-pass 30 Hz, fourth order, no
notch), ocular correction by independent component analysis, and segmentation into the two
resting blocks. Artefact handling was by visual inspection only, with no automatic
amplitude, gradient or low-activity criteria. Essentially all marked intervals fell in the task portions of the recording and not in the resting blocks. The full five minutes was therefore retained for every participant in both conditions, and retention does not vary across the
sample. Because no dedicated electro-oculogram electrodes were recorded, the ocular
correction used surrogate channels derived from the montage. Power spectral density was estimated over eight occipito-parietal
electrodes (O1, Oz, O2, P3, Pz, P4, P7, P8) by Welch's method [@welch1967] (2-s Hann
windows, 50% overlap) and integrated into band power for delta (1--4 Hz), theta (4--8 Hz),
alpha (8--13 Hz), beta (13--30 Hz) and gamma (30--45 Hz). The individual alpha frequency
(IAF) was the frequency of maximum power in the 7--13 Hz range [@klimesch1999].
@corcoran2018 recommend a smoothed peak-alpha or centre-of-gravity estimator in
preference to this argmax. The specparam IAF recovered in the aperiodic decomposition
serves as the cross-check here.

That argmax has a failure mode that matters here, because the one suggestive neural
result rests on this measure. For a spectrum that merely decays across the search range,
with no peak in it, the argmax returns the lower bound of the range, so a value at 7 Hz is
not necessarily a measured peak. The aperiodic decomposition resolves an alpha peak in
`r rseeg_peak("resolvable")` of the `r rseeg_peak("recorded")` eyes-closed recordings. For
the remainder, the individual alpha frequency should be read as unresolved. No analysis
excluding those recordings is reported here, so the
individual-alpha-frequency association below is provisional in that respect.

The gamma band requires an explicit caveat, because it lies entirely above the offline
low-pass corner. With the low-pass at 30 Hz, the whole 30--45 Hz band sits at or beyond the
filter's $-3$ dB point. The filter's transmission was weighted by each participant's own
extrapolated aperiodic background (the broadband $1/f$ component on which the oscillatory
peaks sit), and the quantiles were taken across every recorded resting-state spectrum
rather than across a fitted model's sample. On that basis, `r gamma_att_clause()`.
What survives is further contaminated. No notch filter was applied, and appreciable mains
energy at 50 Hz persists through the low-pass, so part of the residual band is line noise.
Consistent with that, the power observed across
30--45 Hz far exceeds what the filter would transmit of the aperiodic background
(`r gamma_att_obs_clause()`), so most of the residual band is not filtered neural
background at all. The gamma predictor is therefore not a valid index of
gamma-band activity, and we do not interpret it. We retain it in the reported models only
so that the predictor set is not silently altered after the fact. Its null result should be disregarded, and on no account read as evidence that gamma-band activity is unrelated to learning.

We considered recomputing the band from the unfiltered continuous recordings, whose acquisition bandwidth of DC--131 Hz retains the frequencies of interest, and decided against it, because the resulting measure would not be interpretable either.
Surface 30--45 Hz activity overlaps completely with the spectrum of cranial
electromyogram, which extends from roughly 20 Hz upwards. Recordings made under
complete neuromuscular blockade show differences of one to two orders of magnitude in
scalp power above 20 Hz relative to the unparalysed state
[@whitham2007; @muthukumaraswamy2013]. Separating neuronal from myogenic contributions in
this range requires either explicit modelling of that separation or dedicated
electromyogram channels [@hipp2013], and this montage provides neither. Recovering the
band would therefore replace a filter-attenuated measure with a muscle-dominated one, and
would additionally change a pre-specified predictor after the analysis had been run. We report the band as unmeasurable in this dataset and treat a properly acquired gamma measure as a design requirement for future work, since nothing here can retrieve it.

The eyes-closed values serve as the neural predictors, since the alpha rhythm and its peak
are most reliably estimated with the eyes closed. The reduction reproduces the expected
posterior alpha-blocking on eye opening, which we treat as a check on the recording.
Had it been absent, it would have indicated a problem with the montage or the reference
before any predictor built from these data could be trusted. All continuous predictors were
$z$-scored, so each coefficient is expressed per standard deviation of its predictor and the
predictors are directly comparable with one another.

## Learning outcome and model structure (Part A)

The learning outcome is the longitudinal trajectory of grammaticality-judgement accuracy,
modelled jointly at trial level. The two-stage alternative
is what most aptitude studies do: fit a slope per learner, then regress
those slopes on the predictors. It biases the predictor effects through regression
dilution, because a slope estimated from few noisy trials is treated as if it were measured
without error, and it understates uncertainty for the same reason [@gelman2005;
@humberg2024]. We therefore fit a single Bayesian multilevel Bernoulli model of trial-level
accuracy. The learning *rate* is a by-participant random slope for session and the
predictors enter as interactions with session [@barr2013], so that under partial pooling a
learner whose slope is poorly determined contributes proportionately less to the predictor
estimate [@gelmanhill2007]:
$$
\text{correct} \sim \text{Session} \times \text{Property}
 + \text{Session} \times (\text{WM} + \text{Inhibition} + \text{StatLearning} + \text{rs-EEG})
 + (1 + \text{Session} \mid \text{participant}).
$$

Session was coded as a continuous standardised variable so that the rate is a single
slope. A linear effect of session is the parsimonious default. A generalised additive
mixed-model (GAMM) form was held in reserve should the trajectory prove nonlinear
[@winterwieling2016; @soskuthy2017], since averaging across learners can distort the
apparent functional form of a learning curve [@heathcote2000; @murre2011]. That reserve
analysis was not in the end fitted, and no formal test of linearity is reported, so the
session term stands as a linear approximation to the observed trajectories.
@fig-accuracy-trajectory is where its adequacy can
be judged, and the gender-agreement trajectories in particular are not monotone.

We report both the learning rate (the predictor-by-session
interactions) and the attainment (the model-implied end-of-training accuracy), and assess
the reliability of the by-participant rate and attainment before interpreting either as an
individual difference [@hedge2018; @hedge2021]. These are single-occasion, model-based
separation reliabilities estimated within this sample, computed as the variance of the
shrunken by-participant posterior means relative to the model's between-participant
variance, $\mathrm{Var}(\hat{u}_i)/\tau^2$ (by the normal-theory identity, approximately
$1 - \overline{\mathrm{Var}(u_i \mid \text{data})}/\tau^2$).
They are not test-retest reliabilities, which would require a second measurement wave [@rouderhaaf2019].
Because session is standardised, a predictor's main effect is its association with accuracy at the average session. We label the main effects as overall accuracy and reserve *attainment* for the model-implied end-of-training quantity. Pooled
accuracy is shaped by properties of differing difficulty entering cumulatively across
sessions: gender agreement from Session 2, differential object marking from Session 3 and
verb--object number agreement from Session 4 (see the companion paper for the design).
As a sensitivity analysis, Part A is therefore additionally re-fit restricted to gender
agreement, the property with full session coverage and the most stable trajectory.

## Pre/post models of executive function (Part B)

Part B fits a Gaussian pre/post
model of each cognitive index (Session 1 versus Session 5) with a by-participant intercept and a
$\pm0.5$ pre/post contrast. Because the study has no no-training control group, a pre/post change is confounded with the generic threats that attend any uncontrolled retest [@shadish2002]. We add a joint model across the three $z$-scored indices in which each acts as an internal control for the others. With digit span as the reference level, the
session effect is its pre/post change and each measure-by-session term is how much less that
measure changed. The model can therefore identify a change concentrated in one index and set it
against the broad shift that practice, regression to the mean or expectancy would produce.
The three indices are standardised within measure and enter on their recorded polarity,
which is not aligned across them: higher digit span is better performance, higher Stroop
interference worse. A generic improvement therefore moves the two in opposite signed
directions, so the digit-span-versus-Stroop contrast confounds specificity with that
opposition and should be read as an upper bound on the separation. Because resting-state EEG was recorded only at baseline, no neural pre/post is
estimated. The rs-EEG measures enter Part A as baseline predictors (the rs-EEG term
above).

## Priors

Priors are informative and standardised on the logit scale, established from the
literature reviewed in the Introduction and following a selective, evidence-tiered scheme. The cited behavioural associations are *attainment* correlations, so each small informative prior is placed on a predictor's main effect and shrunk towards zero on its interaction with session. Working memory and statistical learning receive small
positive priors [@linck2014; @kurokawa2026; @zhou2024; @misyak2012]. Inhibitory control
receives a near-zero regularising prior, reflecting its weak-to-null evidence base
[@huensch2024; @linckweiss2015; @chen2022], and the resting-state EEG predictors receive
near-zero regularising priors. Although the precedent points to a higher-frequency advantage
[@prat2016; @prat2019], those effect sizes were obtained over a different scalp region and
for a different outcome, learning rate. The bands are therefore treated as an extension of
that question to a new region and outcome, and no directional prior is encoded. The
consequence is asymmetric: a credible band-power effect over posterior sites would have
counted for the precedent, but a null obtained under regularising priors at this sample
size cannot count against it. Variance components receive generous
half--Student-$t(3,0,2.5)$ priors and correlations an $\mathrm{LKJ}(2)$ prior
[@gelman2006; @lewandowski2009]. In a sensitivity analysis, the Part A predictors were re-fit under weakly informative priors. As intended, the regularising priors on the
resting-state bands shrank them towards zero and tightened their intervals (across all
coefficients, the weakly informative intervals were about `r prior_ci_widening()` wider
at the median). No band-power predictor was credible under either prior. The individual
alpha frequency, however, is prior-sensitive. Under the weakly informative prior, its
gender-agreement interval excluded zero
(`r sens_effect("p2_predictive_trajectory_gender", "z_iaf", "weak")`), whereas under the
regularising prior it did not
(`r sens_effect("p2_predictive_trajectory_gender", "z_iaf", "inf")`). The regularised estimate is treated as primary, and we read the association as suggestive. The working-memory main effect was, if anything, larger under the
weakly informative prior, and its pooled interval then excluded zero
(`r sens_effect("p2_predictive_trajectory", "z_digit_span", "weak")`). The
predictor-by-session interactions likewise remained non-credible under weakly informative
priors (largest $p_d$ = `r sens_max_rate_pd()`), so the attainment-not-rate asymmetry reported below is a property of the data and does not depend on the prior specification. The Part B pre/post
models use weakly informative priors throughout
[@schad2021; @depaoli2017]. The weakly informative sensitivity fits place
$\mathcal{N}(0, 1.5)$ on the intercept and on every population-level coefficient, with the
variance and correlation priors unchanged.

The full specification is as follows, so that the models can be rebuilt from the manuscript
alone. In the Part A models, the intercept has a $\mathcal{N}(1, 1)$ prior on the logit scale
(judgement accuracy expected above chance) and any population-level coefficient without a
term-specific prior has $\mathcal{N}(0, 1)$. By-participant standard deviations have
half--Student-$t(3, 0, 2.5)$ and the random-effect correlation matrix $\mathrm{LKJ}(2)$. The
term-specific priors are, for the main effect and its interaction with session respectively:
working memory $\mathcal{N}(0.10, 0.25)$ and $\mathcal{N}(0.05, 0.15)$; statistical learning
$\mathcal{N}(0.10, 0.20)$ and $\mathcal{N}(0.05, 0.15)$; inhibitory control
$\mathcal{N}(0, 0.15)$ and $\mathcal{N}(0, 0.10)$; and every resting-state predictor,
whether a raw band power, the individual alpha frequency or a quantity derived from the
specparam spectral decomposition described below, $\mathcal{N}(0, 0.25)$ and
$\mathcal{N}(0, 0.15)$. The Part B models use
$\mathcal{N}(0, 1)$ on the intercept and on coefficients, and half--Student-$t(3, 0, 2.5)$
on both the by-participant standard deviation and the residual standard deviation. The
joint model, whose by-participant structure includes a random pre/post slope, additionally
places an $\mathrm{LKJ}(2)$ prior on the random-effect correlation.

```{r prior-spec-check}
#| results: asis
# Render-time consistency check for the prior specification stated in prose above.
# The fit scripts are the source of truth; once the cluster job writes
# results/_prior_specification.csv (brms::prior_summary() rows per model), this chunk
# compares the artefact against the prose values, kept here in one named vector, and
# prints a visible warning marker on any divergence. Silent when the artefact is
# absent or everything matches.
.prior_spec_tbl <- .read_csv_if(paper2_results("_prior_specification.csv"))
if (!is.null(.prior_spec_tbl) &&
    all(c("class", "coef", "prior") %in% names(.prior_spec_tbl))) {
  .norm <- function(x) gsub("\\s+", "", tolower(x))
  # The prose values for the Part A pooled model: coefficient-specific slabs, the
  # intercept, and the default population-level prior (empty coef).
  # ".default" stands for the class-b prior with no coefficient (an empty coef in
  # prior_summary()); an empty string cannot be used as a name in a named vector.
  .expected_b <- c(
    "z_digit_span"                = "normal(0.1, 0.25)",
    "z_session_time:z_digit_span" = "normal(0.05, 0.15)",
    "z_asrt"                      = "normal(0.1, 0.2)",
    "z_session_time:z_asrt"       = "normal(0.05, 0.15)",
    "z_stroop"                    = "normal(0, 0.15)",
    "z_session_time:z_stroop"     = "normal(0, 0.1)",
    "z_alpha"                     = "normal(0, 0.25)",
    "z_session_time:z_alpha"      = "normal(0, 0.15)",
    ".default"                    = "normal(0, 1)")
  ps <- .prior_spec_tbl
  if ("model" %in% names(ps)) ps <- ps[ps$model == "p2_predictive_trajectory", , drop = FALSE]
  .mism <- character(0)
  for (cf in names(.expected_b)) {
    is_default <- identical(cf, ".default")
    rows <- ps[ps$class == "b" &
                 (if (is_default) (is.na(ps$coef) | ps$coef == "") else ps$coef == cf), ,
               drop = FALSE]
    if (nrow(rows) < 1) next  # term absent from the artefact; nothing to compare
    if (!identical(.norm(rows$prior[1]), .norm(.expected_b[[cf]])))
      .mism <- c(.mism, sprintf("`%s`: fitted `%s` vs prose `%s`",
                                if (is_default) "(default b)" else cf,
                                rows$prior[1], .expected_b[[cf]]))
  }
  ri <- ps[ps$class == "Intercept", , drop = FALSE]
  if (nrow(ri) >= 1 && !identical(.norm(ri$prior[1]), .norm("normal(1, 1)")))
    .mism <- c(.mism, sprintf("Intercept: fitted `%s` vs prose `normal(1, 1)`", ri$prior[1]))
  if (length(.mism))
    cat("\n\n**[PRIOR-SPECIFICATION MISMATCH -- the fitted priors differ from the",
        " specification stated above: ", paste(.mism, collapse = "; "),
        ". Reconcile before submission.]**\n\n", sep = "")
}
```

## Sampler, diagnostics and inference

Models were fitted with `brms` [@burkner2017; @burkner2018] as a front end to Stan
[@carpenter2017], using the same sampler configuration as the companion paper: four chains
of 10,000 iterations each, the first 2,000 discarded as warm-up, giving 32,000 post-warm-up
draws, with `adapt_delta` = 0.95 and `max_treedepth` = 12. Sampling used within-chain
`reduce_sum` threading at two threads per chain. Because `reduce_sum` takes the grouping of
its partial sums from the thread count, the recorded draws depend on that setting as well as
on the seed. The convergence gate was
rank-normalised $\widehat{R} < 1.01$, bulk- and tail-ESS above 400, and zero divergent
transitions [@vehtari2021]. Posterior predictive checks were written to the repository at
fit time [@gabry2019], and inference was by posterior medians, 95% credible intervals and
the probability of direction ($p_d$) [@makowski2019; @kruschke2018]. All credible intervals are
equal-tailed quantile intervals, bounded by the 2.5th and 97.5th percentiles of the
posterior draws. An ESS of 400 is the threshold at which the $\widehat{R}$ and ESS
diagnostics are themselves trustworthy [@vehtari2021]. It is a weaker guarantee than the
stability of an interval *limit*, which is estimated far less efficiently than a posterior
mean, so borderline coefficients should be read with that in mind [@kruschke2021]. As in the companion paper, we
describe an effect as credible when its 95% credible interval excludes zero, and otherwise read the probability of direction as a continuous index of directional certainty. We do not treat it as a dichotomous test. The projection-predictive and aperiodic analyses narrow the predictor space. We apply no blanket multiplicity correction, which would inflate Type II error without an inferential gain [@rubin2021]. Binary accuracy is modelled with a
Bernoulli (logit) likelihood throughout. Analyses were carried out in R `r prov("R")`
[@rcoreteam] with `brms` `r prov("brms")` as a front end to CmdStan `r prov("CmdStan")`,
`projpred` `r prov("projpred")` for the projection-predictive selection and `loo`
`r prov("loo")` for the Pareto-smoothed importance-sampling leave-one-out (PSIS-LOO)
comparison, under the fixed seed
`r prov("seed:LES_SEED", "20260607")`, in a reproducible `here()`-anchored
compendium [@marwick2018]. The models were fitted on a Linux high-performance computing
cluster (`r prov("platform", "*[platform pending a pipeline run]*")`), where the software
environment is supplied by environment modules and not by a container. The versions given
above are therefore those of the fitting run, and not necessarily those of the machine that
renders this document. The Stan models were compiled with GCC 14.2.0, the compiler the R
module supplies,
at an optimisation level pinned to `-O1` in CmdStan's `make/local`, because that compiler
version raises an internal compiler error on Stan's `reduce_sum` templates at `-O2` and
above. The pin governs whether the models compile and not the draws
they produce. The repository's cluster environment script applies it automatically, so a
rebuild on the same cluster inherits it and a rebuild elsewhere under the same compiler needs
it set by hand. Figures are drawn with `ggplot2` [@wickham2016] using the
theme and colour system of `depictr` [@depictr], whose default qualitative palette is the
colourblind-safe Okabe--Ito set. The companion paper uses the same system, so the two
share one visual identity.

## Predictor selection and a 1/f-adjusted robustness check

Two further analyses probe the neural predictors. First, raw resting-state band power
conflates genuine oscillatory activity with the aperiodic $1/f$ background, and that
background is itself a substantive individual difference [@donoghue2020; @ouyang2020]. We
re-parameterised each participant's baseline resting spectrum into an aperiodic component
(exponent and offset), $1/f$-adjusted oscillatory (theta, alpha, beta) power and the
specparam-estimated individual alpha frequency, using a specparam
decomposition [@donoghue2020], and re-fit Part A with these de-confounded quantities
in place of the raw bands. The decomposition is our own R implementation of that algorithm
and does not call the Python `fooof` package. It was fitted over 2--40 Hz with at most six peaks,
peak widths constrained to 1--12 Hz, a minimum peak height of 0.05 log$_{10}$ power above the
flattened spectrum, a peak threshold of two standard deviations of that spectrum, and the
aperiodic component in fixed (knee-free) mode. This aperiodic variant is a robustness
check of the raw-band model, motivated by the periodic/aperiodic conflation. It is not a
response to the raw-band result, and it does not amount to a second independent pass over
the predictor space. Like the rest of the analysis plan, it was specified after data
collection (see *Transparency and preregistration*). The two reference models are compared
by leave-one-out cross-validation [@vehtari2017], so the choice between the two neural
characterisations rests on a quantified comparison.

A trial-level criterion cannot settle that comparison on its own. The two reference
models differ only in participant-level resting-state predictors, which are constant
within a participant. With about `r aper_loo_tpp("pooled")` trials per participant in the pooled stratum, a held-out trial's by-participant intercept is already pinned by that participant's remaining trials, so a between-participant predictor can add little to
predicting it. Leave-one-*trial*-out PSIS-LOO is therefore precise, but about a quantity
that is not the one at issue, and it would return a near-zero difference for informative
and uninformative resting-state predictors alike. Blocking the folds on the grouping
factor is the recommended remedy wherever such dependence exists, even when the fitted
model already accounts for the dependence with a random effect, as ours does
[@roberts2017]. We therefore also compare the models by K-fold cross-validation with
folds blocked on participants (K = 10, the same fold assignment for both models,
participants dealt round-robin over a seeded shuffle). We report the difference in
expected log predictive density (elpd) with a standard error clustered on participants.
The blocking follows @roberts2017 and
the predictive score @vehtari2017. The clustered standard error is our own construction.
Because the pointwise contributions are not independent within a participant, we sum them
within participant before taking the standard error across participants. A difference is
read as decisive only when it exceeds twice this clustered standard error, and the naive
independent-observation standard error is reported alongside so the size of the clustering
correction is visible.

Second, the joint Part A model deliberately retains every candidate predictor for
inference, but one also wants to know which subset accounts for its predictive
performance. We answer this with projection-predictive variable selection
[@piironen2017; @piironen2020],
which projects the rich reference posterior onto nested submodels and ranks predictors
by the predictive performance retained after projection. Each submodel inherits the
reference model's regularisation and is not re-estimated from scratch. Because
the reference model is a multilevel Bernoulli model, we use the latent projection
[@catalina2021], the appropriate and recommended projection for this model class. The
forward search is cross-validated in full, with every selection step re-run within each
fold, to pay for the selection-induced optimism of naive fit-then-select pipelines. The
folds are formed by participant, so that no participant's correlated trials
straddle the train/test split [@roberts2017]. Ten such participant-grouped folds were
used, with twenty clustered posterior draws during the search and 400 draws for
performance evaluation, which was carried out on the response (0/1) scale under the
latent projection. The suggested submodel size is projpred's standard
heuristic: the smallest size whose cross-validated elpd lies within one standard error of
the reference model's [@piironen2020]. A predictor that the minimal submodel does
not retain is not thereby shown to be null. The selection summarises which predictors
account for the reference model's predictive performance, and the full Part A posterior
remains the inferential statement [@piironen2020]. The raw-band and aperiodic reference
models' selections are reported side by side.

## Transparency and preregistration

The study design, materials and session protocol were preregistered before data
collection (<https://osf.io/tjr54>). The preregistration fixed the design, the tasks and
the six-session schedule but did not specify a statistical analysis plan. The Bayesian
multilevel models, the informative priors, the specparam/aperiodic re-parameterisation and
the projection-predictive selection reported here were therefore specified after the data
were collected, and they are exploratory rather than confirmatory
[@nosek2018; @wagenmakers2012]. The
preregistration also motivated the Session-1 battery as an index of attention relevant to
the selection of a source of transfer. The present study reframes the same three tasks as
working-memory, inhibitory-control and statistical-learning measures, used both as
predictors of learning and as pre/post outcomes. That re-conceptualisation is a departure
from the registered rationale and is likewise exploratory.

Data collection was preregistered for two sites, Norway and Spain. The present analyses
concern only the Norwegian cohort (first-language Norwegian, second-language English). The
Spanish-site data are not analysed here, and the findings generalise only
to first-language-Norwegian learners. Finally, the preregistered protocol required participants to score above 80% on a post-training behavioural test in each grammar session (with a second attempt allowed) before they could proceed. This performance-conditional gate shapes which
participant-sessions contribute learning data and is part of the attrition described above.

# Results

```{r convergence}
#| echo: false
#| results: asis
# The static lead-in and the generated count are emitted together, as one paragraph. Split
# across a sentence of prose and a chunk they rendered as two paragraphs of one sentence
# each, which reads as a stub.
if (!results_ready) {
  cat("::: {.callout-note title=\"Reproducible draft\"}\n",
      "Model fits are not yet present. Every figure and statistic below populates ",
      "automatically once the pooled result files are produced by the modelling ",
      "pipeline.\n:::\n", sep = "")
} else {
  np <- sum(conv_tbl$passed, na.rm = TRUE)
  # The divergence clause is built conditionally so a future run with divergent
  # transitions cannot print a self-contradictory paragraph.
  div_clause <- if (all(!is.na(conv_tbl$n_divergent)) && all(conv_tbl$n_divergent == 0)) {
    "and no divergent transitions"
  } else {
    sprintf("and %d divergent transitions in total (see @tbl-convergence)",
            sum(conv_tbl$n_divergent, na.rm = TRUE))
  }
  cat(sprintf(paste0(
    "Convergence diagnostics for every model reported below are given in ",
    "@tbl-convergence, in the appendix, as realised values, since a realised ",
    "value distinguishes a model that converged comfortably from one that converged ",
    "marginally. All %d models were assessed against the same gate and %d passed it, ",
    "with rank-normalised $\\widehat{R}$ below 1.01, effective sample sizes above 400 ",
    "in both the bulk and the tail, %s.\n"),
    nrow(conv_tbl), np, div_clause))
}
```


```{r}
#| results: asis
# The pooled convergence artefact is written per fitted model, so any model that has not
# been run is silently absent rather than shown as failing. State that explicitly instead
# of letting the table imply the grid is complete.
if (!is.null(conv_tbl)) {
  .expected <- c("p2_predictive_trajectory", "p2_predictive_trajectory_gender",
                 "p2_predictive_trajectory_aperiodic", "p2_predictive_trajectory_aperiodic_gender",
                 "p2_cognition_prepost_digit_span", "p2_cognition_prepost_stroop_interference",
                 "p2_cognition_prepost_asrt_learning", "p2_cognition_prepost_joint")
  .missing <- setdiff(.expected, conv_tbl$model)
  if (length(.missing))
    cat("\nDiagnostics are not yet available for ", length(.missing),
        " of the models reported elsewhere in this paper (",
        paste0("`", .missing, "`", collapse = ", "),
        "), whose fits predate the current diagnostic artefact.\n", sep = "")
}
```

## Descriptive overview

Before the models, we show the observed data they summarise. @fig-accuracy-trajectory
plots the raw learning trajectory, which is also the direct check on the
linear-in-session assumption the models make. @fig-predictor-corr shows the baseline
predictors' intercorrelations and @tbl-predictor-desc gives the same predictors
numerically. The per-measure distributions (@fig-predictor-dist) and the resting-state
spectra behind the band-power and individual-alpha-frequency measures (@fig-rs-psd) are
given in the online supplementary material (@apx-supplementary), which is deposited with
the analysis code in this paper's repository. These displays cover three different
frames, as their notes state. @fig-predictor-dist plots the stored records for each measure one
panel at a time, so the number of points differs between panels. @fig-predictor-corr and
@tbl-predictor-desc cover the participants who have both a baseline cognitive battery and
an eyes-closed resting-state spectrum, and @fig-rs-psd covers every recorded spectrum. All
three are wider than the `r proj_n(projpath_raw)` participants of the fitted raw-band
model, so their counts describe the measures rather than the models.

One feature of the learning trajectory bears on how *rate* and *attainment* should be read,
since both terms would otherwise suggest improvement. Mean accuracy is highest
at each property's first session and lower at its last, in every property and both
mini-languages. The session term is credibly negative both pooled
(`r report_effect(LES_P2_MODELS[["predictive"]], "z_session_time")`) and within gender
agreement (`r report_effect(LES_P2_MODELS[["predictive_gender"]], "z_session_time")`). In
the pooled model, this partly reflects the design: properties of increasing difficulty
enter cumulatively across sessions, so later sessions sample harder material (see *Learning
outcome and model structure*). Within gender agreement, which is sampled throughout, that
explanation does not apply, and we record the decline as an unexplained feature of these
data. Throughout what follows, *rate* is the slope of this trajectory and *attainment* its
model-implied end point, whichever way the trajectory runs.

```{r}
#| label: fig-accuracy-trajectory
#| fig-cap: "Observed Grammaticality-Judgement Accuracy Across Sessions, by Grammatical Property and Mini-Language"
#| apa-note: "Thin lines are individual participants, and the ribbon is a 95% confidence interval computed between participants, the sampling unit being participants and not trials. This is the raw trajectory the joint model summarises, and the direct check on the linear-in-session assumption."
#| fig-width: 6.5
#| fig-height: 3.4
if (!is.null(acc_traj_tbl)) {
  gc <- acc_traj_tbl[acc_traj_tbl$level == "group_cell" & !is.na(acc_traj_tbl$grammatical_property), ]
  bp <- acc_traj_tbl[acc_traj_tbl$level == "participant_cell" &
                       !is.na(acc_traj_tbl$grammatical_property), ]
  if (!nrow(bp)) bp <- NULL
  # Order the panels by the session in which each property was introduced (gender
  # agreement in Session 2, differential object marking in Session 3, verb-object
  # number agreement in Session 4). The column is character, so without an explicit
  # factor the panels would fall alphabetically and scramble the design. This must
  # precede the ggplot()/geom_line() calls below, which capture their data by value.
  .prop_order <- c("gender_agreement", "differential_object_marking",
                   "verb_object_number_agreement")
  prop_lab <- as_labeller(c(gender_agreement = "Gender agreement",
                            differential_object_marking = "Differential object marking",
                            verb_object_number_agreement = "Verb-object number agreement"))
  gc$grammatical_property <- factor(gc$grammatical_property, levels = .prop_order)
  if (!is.null(bp)) bp$grammatical_property <- factor(bp$grammatical_property,
                                                      levels = .prop_order)
  p <- ggplot(gc, aes(session, mean_accuracy, colour = mini_language))
  if (!is.null(bp)) p <- p + geom_line(data = bp,
                                       aes(session, mean_accuracy,
                                           group = interaction(participant_lab_ID, mini_language)),
                                       colour = "grey80", linewidth = 0.2, alpha = 0.5)
  p + geom_ribbon(aes(ymin = mean_accuracy - 1.96 * se, ymax = mean_accuracy + 1.96 * se,
                      fill = mini_language), alpha = 0.2, colour = NA) +
    geom_line(aes(linetype = mini_language), linewidth = 0.7) + geom_point(size = 1) +
    facet_wrap(~ grammatical_property, labeller = prop_lab) +
    # Okabe-Ito blue and vermillion, the same pair fig-predictor-forest uses, so every
    # hard-coded scale in the paper draws on the colourblind-safe set the Method names.
    scale_colour_manual(values = c(`Mini-English` = "#0072B2", `Mini-Norwegian` = "#D55E00"),
                        aesthetics = c("colour", "fill")) +
    scale_linetype_manual(values = c(`Mini-English` = "solid", `Mini-Norwegian` = "22")) +
    labs(x = "Session", y = "Proportion correct", colour = NULL, fill = NULL, linetype = NULL) +
    les_theme() + theme(legend.position = "top")
} else cat("*Raw accuracy trajectory will appear once `_accuracy_trajectory_descriptives.csv` is present.*")
```

```{r}
#| label: fig-predictor-corr
#| fig-cap: "Pearson Correlations Among the Baseline Predictors"
#| apa-note: "Correlations are computed on the z-scored predictors, over the participants who have both a baseline cognitive battery and an eyes-closed resting-state spectrum, pairwise-complete, so the n behind each cell varies slightly. Only the lower triangle is shown, the redundant mirror and the unit diagonal being omitted. The resting-state band powers are expected to be positively intercorrelated through the shared aperiodic 1/f background, which motivates the specparam robustness check and is why the projpred ranking must be read against this structure."
#| fig-width: 5.6
#| fig-height: 5
if (!is.null(predictor_corr)) {
  ord <- c("digit_span", "stroop_interference", "asrt_learning",
           "delta", "theta", "alpha", "beta", "gamma", "iaf")
  corr_labs <- c(digit_span = "Digit span", stroop_interference = "Stroop",
                 asrt_learning = "ASRT", delta = "Delta", theta = "Theta",
                 alpha = "Alpha", beta = "Beta", gamma = "Gamma", iaf = "IAF")
  strip0 <- function(x, d = 2) sub("^(-?)0\\.", "\\1.", formatC(x, format = "f", digits = d))
  pc <- predictor_corr
  # Keep one triangle only: drop the redundant upper mirror and the r = 1 diagonal.
  pc <- pc[match(as.character(pc$var1), ord) < match(as.character(pc$var2), ord), ]
  pc$var1 <- factor(pc$var1, levels = ord); pc$var2 <- factor(pc$var2, levels = rev(ord))
  ggplot(pc, aes(var1, var2, fill = r)) +
    geom_tile(colour = "white", linewidth = 0.5) +
    geom_text(aes(label = strip0(r)), size = 3) +
    scale_fill_gradient2(low = les_div[1], mid = les_div[2], high = les_div[3], midpoint = 0,
                         limits = c(-1, 1), labels = function(x) strip0(x, 1)) +
    scale_x_discrete(labels = corr_labs) + scale_y_discrete(labels = corr_labs) +
    labs(x = NULL, y = NULL, fill = "r") +
    les_theme() +
    # Blank the specific grid elements, not just the parent: a theme that sets
    # panel.grid.major.x/y explicitly (as depictr's does) is not cleared by setting
    # panel.grid alone, and the grid would then show through the masked triangle.
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          panel.grid = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank())
} else cat("*Predictor correlation matrix will appear once `_predictor_correlations.csv` is present.*")
```

```{r}
#| label: tbl-predictor-desc
#| tbl-cap: "Descriptive Statistics for the Baseline Predictors"
#| apa-note: "The frame is the participants who have both a baseline cognitive battery and an eyes-closed resting-state spectrum, which is wider than the sample any Part A model is fitted to, and the n column gives the number contributing to each row. Each row carries its own unit. Because a band power is the spectral density integrated over the band, its unit is squared microvolts and not squared microvolts per hertz."
if (!is.null(predictor_desc) && requireNamespace("knitr", quietly = TRUE)) {
  pd <- predictor_desc
  num <- vapply(pd, is.numeric, logical(1)); pd[num] <- lapply(pd[num], round, 2)
  # Map the pipeline's snake_case keys and lowercase headers to reader-facing labels,
  # using the same names the figures use, and carry the unit on each row so the table
  # is readable without the Method. Done defensively: if the artefact's columns ever
  # change, the raw frame is printed rather than the chunk failing.
  .pdl <- c(digit_span = "Working memory: digit span (items correct)",
            stroop_interference = "Inhibitory control: Stroop interference (ms)",
            asrt_learning = "Statistical learning: ASRT difference (ms)",
            delta = "Delta power (µV²)", theta = "Theta power (µV²)",
            alpha = "Alpha power (µV²)", beta = "Beta power (µV²)",
            gamma = "Gamma power (µV², not interpreted)",
            iaf = "Individual alpha frequency (Hz)")
  if (all(c("predictor", "n", "mean", "sd", "median", "min", "max") %in% names(pd))) {
    lab <- unname(.pdl[pd$predictor])
    pd <- data.frame(Predictor = ifelse(is.na(lab), pd$predictor, lab),
                     n = pd$n, M = pd$mean, SD = pd$sd, Mdn = pd$median,
                     Min = pd$min, Max = pd$max, check.names = FALSE)
  }
  knitr::kable(pd)
} else if (is.null(predictor_desc)) cat("*Predictor descriptive table pending `_predictor_descriptives.csv`.*")
```

## Part A: predictors of the learning trajectory

Each predictor enters twice: as a main effect, indexing the overall level of accuracy, and
in interaction with session, indexing the learning *rate*. Because session is standardised,
the main effect is the association at the average session (see Method). Working memory was
the clearest predictor. Higher baseline digit span went with higher overall accuracy,
credibly so for the fully sampled gender-agreement property
(`r report_effect(LES_P2_MODELS[["predictive_gender"]], "z_digit_span")`) and positively,
though with an interval that narrowly included zero, in the pooled model
(`r report_effect(LES_P2_MODELS[["predictive"]], "z_digit_span")`). Inhibitory control
(`r report_effect(LES_P2_MODELS[["predictive"]], "z_stroop")`) and statistical learning
(`r report_effect(LES_P2_MODELS[["predictive"]], "z_asrt")`) were not credibly related to
accuracy, in line with their weak-to-null evidence base.

Of the baseline resting-state measures, the band powers were uninformative
(@fig-predictor-forest), but that null must be read against how strongly they are
intercorrelated. Entering the bands simultaneously means each coefficient is the unique
contribution of that band with the others held constant, and the four low-frequency bands
share a great deal of variance (@fig-predictor-corr). Their variance inflation factors are
`r vif_range(c("delta","theta","alpha","beta"))`, which widens their posterior intervals by
roughly `r vif_widen(c("delta","theta","alpha","beta"))` relative to an orthogonal
predictor. Their null coefficients are therefore weakly identified and uninformative about whether those rhythms relate to learning, and they are not directly
comparable to the zero-order band--outcome associations reported in the precedent
literature. Two of the neural predictors escape this: gamma
(VIF `r vif_of("gamma")`) and the individual alpha frequency (VIF `r vif_of("iaf")`) are
close to orthogonal to the rest, so their estimates stand on their own. The one
suggestive signal was the individual alpha frequency, whose association with
gender-agreement accuracy was
`r report_effect(LES_P2_MODELS[["predictive_gender"]], "z_iaf")`. The cross-check the
Method promised points the same way without settling it. In the aperiodic
re-parameterisation of the same stratum, fitted to a smaller sample, the
specparam-estimated individual alpha frequency was
`r report_effect("p2_predictive_trajectory_aperiodic_gender", "z_specparam_iaf")`, and the
aperiodic exponent, the quantity the decomposition adds, was
`r report_effect("p2_predictive_trajectory_aperiodic_gender", "z_aperiodic_exponent")`. The learning-*rate*
interactions were uniformly weak (for working memory,
`r report_effect(LES_P2_MODELS[["predictive"]], "z_session_time:z_digit_span")`), consistent with priors that placed their small informative mass on attainment and shrank the rate terms towards zero.

```{r p2-selection}
#| results: asis
if (!is.null(projpath_raw)) {
  cat(sprintf(paste0(
    "To establish which predictors account for the reference model's predictive ",
    "performance, projection-predictive selection was run on %s and on the aperiodic ",
    "re-parameterisation, and it returned the same parsimonious answer throughout.%s ",
    "In the raw-band model, the suggested size was %s: the design scaffold ",
    "(by-participant intercept and session slope, session, grammatical property and their ",
    "interaction) together with working memory (digit span). Digit span was the first ",
    "individual-difference predictor selected, entering at step %s of the forward search, ",
    "exactly the suggested size. Inhibitory control (step %s), statistical learning (step %s) ",
    "and the resting-state band predictors entered only beyond the suggested size, so they ",
    "are not among the terms the search retained. The forward search was capped at projpred's ",
    "default path length, so the lowest-ranked terms, including the individual alpha ",
    "frequency, reported above as a suggestive baseline signal, fell below that cap ",
    "and were not positively ruled out in that stratum.\n\n%s\n\n",
    "The aperiodic re-parameterisation reproduced this ",
    "ordering (suggested size %s), with the aperiodic exponent and offset entering at steps ",
    "%s and %s, well beyond the suggested size, so the $1/f$ background added no predictive ",
    "information the raw bands had missed. That search was capped at the same default path length, ",
    "so the adjusted beta power and the specparam individual alpha frequency were not reached ",
    "in it either, and the same caveat applies to them as to the raw model's tail terms. ",
    "This is a statement about predictive parsimony, not ",
    "evidence that the unselected predictors have no effect [@piironen2020]."),
    projpred_strata(), projpred_gender_caveat(),
    proj_size(projpath_raw), proj_rank(projpath_raw, "z_digit_span"),
    proj_rank(projpath_raw, "z_stroop"), proj_rank(projpath_raw, "z_asrt"),
    projpred_gender_result(),
    proj_size(projpath_aper), proj_rank(projpath_aper, "z_aperiodic_exponent"),
    proj_rank(projpath_aper, "z_aperiodic_offset")))
  if (!is.null(aper_loo_tbl)) {
    both_tie <- isTRUE(aper_loo_tie("pooled")) && isTRUE(aper_loo_tie("gender"))
    verdict <- if (both_tie) {
      "in each case smaller than twice its standard error"
    } else {
      sprintf("%s in the pooled stratum and %s for gender agreement",
              aper_loo("pooled", "verdict"), aper_loo("gender", "verdict"))
    }
    cat(sprintf(paste0(
      "\n\nThe two parameterisations were also compared by PSIS-LOO on the sample they ",
      "share. Because the specparam re-parameterisation is defined for %s of the %s participants, ",
      "the raw-band model was refit on that common subsample so the two are compared on ",
      "identical observations. The aperiodic-minus-raw difference in expected log predictive ",
      "density was %s (SE %s) in the pooled stratum and %s (SE %s) for gender agreement, %s ",
      "[@vehtari2017]. As noted in the Method, this leave-one-trial-out criterion cannot ",
      "detect participant-level predictors, so the near-zero difference is not read as ",
      "evidence that the two characterisations are equivalent.%s"),
      proj_n(projpath_aper), proj_n(projpath_raw),
      aper_loo("pooled", "elpd"), aper_loo("pooled", "se"),
      aper_loo("gender", "elpd"), aper_loo("gender", "se"), verdict,
      aper_kf_sentence()))
  }
  cat("\n")
} else {
  cat("*Projection-predictive selection results appear here once the projpred ",
      "solution-path CSVs are present in `paper_2_plasticity/results/`.*\n", sep = "")
}
```

```{r p2-derived}
#| results: asis
if (results_ready && !is.null(attainment_tbl) && !is.null(reliability_tbl)) {
  af <- function(m) { r <- attainment_tbl[attainment_tbl$model == m, ]
    if (nrow(r) == 1) sprintf("%.2f (95%% CrI [%.2f, %.2f])", r$median, r$ci_low, r$ci_high) else "NA" }
  # Reliabilities are bounded above by 1, so the leading zero is stripped, matching relv()
  # and predrel() elsewhere in the document rather than printing the same quantity two ways.
  rf <- function(m, e) { r <- reliability_tbl[reliability_tbl$model == m & reliability_tbl$effect == e, ]
    if (nrow(r) == 1) sub("^0", "", sprintf("%.2f", r$reliability)) else "NA" }
  cat(sprintf(paste0("Model-implied end-of-training attainment was %s pooled and %s for gender ",
              "agreement. The by-participant level at the average session (the intercept) was ",
              "estimated with high reliability (pooled %s, gender-only %s). The learning-rate slope ",
              "was estimated less reliably (pooled %s, ",
              "gender-only %s), and because the gender-only rate is the least reliably estimated of ",
              "these quantities we do not interpret rate-based individual-difference claims from that ",
              "model.\n"),
      af("p2_predictive_trajectory"), af("p2_predictive_trajectory_gender"),
      rf("p2_predictive_trajectory", "Intercept"), rf("p2_predictive_trajectory_gender", "Intercept"),
      rf("p2_predictive_trajectory", "z_session_time"), rf("p2_predictive_trajectory_gender", "z_session_time")))
}
```

@tbl-reliability gives these outcome reliabilities in full, and
@tbl-predictor-reliability the internal consistency of the three baseline cognitive
indices. Both matter, because the association between a predictor and an outcome is
bounded by the reliability of each. Two of the three cognitive indices are moreover
difference scores, a form whose reliability is typically lower than that of its
components [@draheim2019; @hedge2018].

```{r}
#| label: fig-predictor-forest
#| fig-cap: "Part A Predictor Effects on Overall Accuracy and Learning Rate"
#| apa-note: "Points are posterior medians with 95% equal-tailed credible intervals for each baseline predictor's association with overall accuracy (the main effect, the association at the average session) and with learning rate (the predictor-by-session interaction), in the pooled and gender-agreement-only models. Estimates are on the logit scale: main effects are the change in log-odds of a correct response per SD of the predictor, interaction terms the change in the per-session slope per SD. The two panels share a common x axis so the families are directly comparable, and the rate-interaction terms carry deliberately tighter priors than the main effects. Predictors are ordered working memory, inhibitory control and statistical learning (for the Stroop-based inhibitory-control predictor, higher scores mean poorer inhibition), then the resting-state measures. Intervals excluding zero indicate effects whose direction is well determined."
#| fig-width: 6.5
#| fig-height: 6
if (results_ready) {
  preds <- c("z_digit_span", "z_stroop", "z_asrt",
             "z_delta", "z_theta", "z_alpha", "z_beta", "z_gamma", "z_iaf")
  rate_terms <- paste0("z_session_time:", preds)
  fp <- post_tbl |>
    filter(model %in% c("p2_predictive_trajectory", "p2_predictive_trajectory_gender"),
           parameter %in% c(preds, rate_terms)) |>
    mutate(Component = ifelse(grepl("^z_session_time:", parameter),
                              "Learning rate (x session)", "Overall accuracy (main effect)"),
           Predictor = factor(.pretty_term(gsub("^z_session_time:", "", parameter)),
                              levels = rev(.pretty_term(preds))),
           Model = ifelse(grepl("_gender$", model), "Gender only", "Pooled"))
  # Both are character, so without explicit levels the facets and the legend fall
  # alphabetically: the rate panel ahead of the main effects, and the gender-only
  # sensitivity model ahead of the primary pooled one, in both cases reversing the order
  # the caption and the surrounding text use.
  fp$Component <- factor(fp$Component,
                         levels = c("Overall accuracy (main effect)", "Learning rate (x session)"))
  fp$Model     <- factor(fp$Model, levels = c("Pooled", "Gender only"))
  if (nrow(fp)) {
    ggplot(fp, aes(x = median, y = Predictor, colour = Model, shape = Model)) +
      geom_vline(xintercept = 0, linetype = 2, colour = "grey50") +
      geom_pointrange(aes(xmin = ci_low, xmax = ci_high),
                      position = position_dodge(width = 0.5)) +
      facet_wrap(~ Component) +
      scale_colour_manual(values = c(Pooled = "#0072B2", `Gender only` = "#D55E00")) +
      scale_shape_manual(values = c(Pooled = 16, `Gender only` = 17)) +
      labs(x = "Posterior estimate (log-odds)", y = NULL, colour = NULL, shape = NULL) +
      les_theme(10) + theme(legend.position = "top")
  }
}
```

```{r}
#| label: tbl-reliability
#| tbl-cap: "By-Participant Reliability of the Learning Trajectory"
#| apa-note: "Columns give the group standard deviation, mean posterior standard error, and reliability of the intercept (a participant's level at the average session, since session is standardised) and of the session (rate) slope, for the pooled and gender-agreement-only models. The reliability column is the empirical-Bayes form, the variance of the shrunken by-participant posterior means over the model's between-participant variance. The group SD and mean posterior SE columns are shown for context, not as the terms of a formula."
if (!is.null(reliability_tbl)) {
  # Reliability is bounded above by 1, so its leading zero is stripped, matching relv()
  # and the Results prose. Group SD and the mean posterior SE are unbounded and keep theirs.
  .strip0 <- function(x) sub("^(-?)0\\.", "\\1.", formatC(x, format = "f", digits = 2))
  rel <- reliability_tbl |>
    transmute(Model = ifelse(grepl("_gender$", model), "Gender only", "Pooled"),
              Quantity = ifelse(effect == "Intercept", "Overall accuracy (intercept)", "Learning rate (session)"),
              `Group SD` = round(group_sd, 2),
              `Mean posterior SE` = round(mean_posterior_se, 2),
              Reliability = .strip0(reliability))
  knitr::kable(rel)
}
```


## Part B: change in executive function over training

Comparing the first and fifth sessions, working-memory span increased over training,
`r report_effect("p2_cognition_prepost_digit_span", "session_prepost")`. The Stroop
interference effect
(`r report_effect("p2_cognition_prepost_stroop_interference", "session_prepost")`) and the
statistical-learning effect
(`r report_effect("p2_cognition_prepost_asrt_learning", "session_prepost")`) showed no
credible change. In the joint model that lets each index control the others, the pattern
pointed the same way. The joint-model digit-span change was
`r report_effect("p2_cognition_prepost_joint", "session_prepost")`, an interval that,
unlike the single-index model's, includes zero. The estimate is less precise in the
joint specification, which shares one residual standard deviation across the three
measures and adds a by-participant change slope. Relative to that change, the
pre/post shift was smaller for the Stroop interference effect
(`r report_effect("p2_cognition_prepost_joint", "session_prepost:measurestroop_interference")`)
and for the statistical-learning effect
(`r report_effect("p2_cognition_prepost_joint", "session_prepost:measureasrt_learning")`).
These specificity contrasts have credible intervals that include zero (probabilities of
direction `r pd_of("p2_cognition_prepost_joint", "session_prepost:measurestroop_interference")`%
and `r pd_of("p2_cognition_prepost_joint", "session_prepost:measureasrt_learning")`%), so
the evidence that the gain is working-memory-specific is suggestive without being decisive.
@fig-prepost shows these effects.

Two features of the comparison qualify it further. First, the three indices are $z$-scored
within measure and enter the joint model on their recorded polarity, and that polarity is
not aligned across them. A higher digit span is better performance, a higher Stroop
interference score worse (see *Cognitive and neural indices*). A generic practice
improvement would therefore push digit span up and Stroop interference down, in opposite
signed directions, so the digit-span-versus-Stroop contrast is in part an artefact of that
opposition, and it overstates the separation correspondingly. The ASRT comparison is not affected in the same way, since a higher ASRT
score also means better performance and its polarity therefore matches digit span's.
Second, the Stroop index has very low internal consistency here (see *Limitations
and future directions*), which leaves it a noisy control. In task units,
the digit-span change was from `r cog_mean("digit_span", 1)` items correct at Session
1 to `r cog_mean("digit_span", 5)` at Session 5, against
`r cog_mean("stroop_interference", 1)` to `r cog_mean("stroop_interference", 5)` ms of
Stroop interference and `r cog_mean("asrt_learning", 1)` to
`r cog_mean("asrt_learning", 5)` ms of ASRT learning.

```{r}
#| label: fig-prepost
#| fig-cap: "Pre/Post (Session 1 to 5) Change in the Executive-Function Indices"
#| apa-note: "Points are single-index session effects and, for the joint model, the digit-span change and the contrasts of each measure's change against digit span (the reference level). Filled points mark effects whose 95% credible interval excludes zero; open points do not."
#| fig-width: 6.5
#| fig-height: 3.5
if (results_ready) {
  pp <- post_tbl |>
    filter((model %in% c("p2_cognition_prepost_digit_span",
                         "p2_cognition_prepost_stroop_interference",
                         "p2_cognition_prepost_asrt_learning") & parameter == "session_prepost") |
           (model == "p2_cognition_prepost_joint" &
              parameter %in% c("session_prepost",
                               "session_prepost:measurestroop_interference",
                               "session_prepost:measureasrt_learning"))) |>
    mutate(Effect = dplyr::case_when(
      model == "p2_cognition_prepost_digit_span" ~ "Digit span change",
      model == "p2_cognition_prepost_stroop_interference" ~ "Stroop change",
      model == "p2_cognition_prepost_asrt_learning" ~ "ASRT change",
      parameter == "session_prepost" ~ "Joint: digit-span change",
      parameter == "session_prepost:measurestroop_interference" ~ "Joint: Stroop vs digit span",
      parameter == "session_prepost:measureasrt_learning" ~ "Joint: ASRT vs digit span"),
      Block = ifelse(grepl("^Joint", Effect), "Joint model", "Single-index models"))
  pp$Block  <- factor(pp$Block, levels = c("Single-index models", "Joint model"))
  pp$Effect <- factor(pp$Effect, levels = c(
    "Joint: ASRT vs digit span", "Joint: Stroop vs digit span", "Joint: digit-span change",
    "ASRT change", "Stroop change", "Digit span change"))
  # Filled points mark effects whose 95% interval excludes zero, matching the
  # convention used in the companion paper's ERP forest.
  pp$Credible <- (pp$ci_low > 0) | (pp$ci_high < 0)
  ggplot(pp, aes(x = median, y = Effect)) +
    geom_vline(xintercept = 0, linetype = 2, colour = "grey50") +
    geom_pointrange(aes(xmin = ci_low, xmax = ci_high, shape = Credible)) +
    scale_shape_manual(values = c(`FALSE` = 1, `TRUE` = 16), guide = "none") +
    facet_grid(Block ~ ., scales = "free_y", space = "free_y") +
    labs(x = "Pre/post change (SD units)", y = NULL) +
    les_theme(10)
}
```

# Discussion

This study asked two questions: whether baseline cognitive and neural measures predict
how well adults learn a new grammar (Part A), and whether executive function changes over
the training period (Part B). The answer to the first is selective. Working memory
predicted how well learners did, while inhibitory control,
statistical learning and the resting-state measures added little, and the predictors acted
on the level learners reached, leaving their rate of change unexplained. The answer to the
second falls on the same
capacity: working-memory span was the one index that rose over training.

## Predictors of overall accuracy and learning rate

Working memory was the clearest cognitive predictor of how well learners did, credibly so for gender agreement and in the same direction, though less certainly, when the properties were pooled. This fits the literature placing working memory as the
best-evidenced cognitive correlate of second-language outcomes [@linck2014]. Inhibitory
control and statistical learning, whose evidence base is weak or contested
[@huensch2024; @linckweiss2015; @chen2022; @zhou2024], added little. Because digit span taps storage and not the executive or complex span that carries most of the working-memory--language association [@linck2014], a complex-span measure might well have shown a larger association. If so, the digit-span estimate is unlikely to overstate the working-memory contribution. Digit span is also the weaker of
the phonological short-term memory tasks [@kurokawa2026, on vocabulary outcomes]. That is a conjecture about what another measure would have shown, and it does not bound the estimate reported here.

The baseline resting-state measures
were largely uninformative. Only the individual alpha frequency, a stable, trait-like
measure [@grandy2013trait] related to general cognitive ability
[@klimesch1999; @grandy2013], hinted at a
positive association with overall accuracy, and only in the gender-agreement model, not the
pooled one. It stands as a lead for confirmatory work, and
one to be held lightly. In the same stratum, the projection-predictive search ranked the
individual alpha frequency last, and the estimate is prior-sensitive (see *Priors*). In a
small number of recordings, that measure is also a boundary value with no resolved peak
behind it (see *Cognitive and neural indices*). The null for the band powers contrasts
with reports that resting beta and low-gamma power predict L2 learning *rate* [@prat2016;
@prat2019]. We did estimate the rate analogue. The band-by-session interactions are in the
model and none was credible, though under weakly informative priors their intervals are
wide enough not to exclude effects of the published size. The divergence therefore tracks
our posterior reduction where theirs was right-temporoparietal, together with our
conservative priors and our precision. The two sets of results differ in scope without
contradicting each other. On the gamma half of that comparison, we can say nothing at all.
Our 30--45 Hz band lies above the 30 Hz offline low-pass, so the
measure is not a valid index of gamma-band power and its null is uninformative by
construction (see *Cognitive and neural indices*). Testing the low-gamma part of that result [@prat2016; @prat2019] would need a recording made without a 30 Hz low-pass and with the means to separate myogenic from neuronal activity. The aperiodic re-parameterisation sharpens the neural
null. Had the flat band coefficients merely reflected the
conflation of oscillatory power with the $1/f$ background [@donoghue2020], de-confounding
them should have uncovered signal. That is not what happened. The aperiodic exponent and offset entered the
selection well beyond the suggested size. Under participant-grouped cross-validation,
neither characterisation predicted a held-out learner's trajectory detectably better than
the other.
Against a largely cross-sectional literature relating the aperiodic exponent to cognitive
function, this bounds the question without answering it. At baseline, in this sample, the exponent carried no detectable incremental signal for a longitudinal training outcome, a test cross-sectional designs cannot run. That the two characterisations
could not be distinguished at this sample size also means the band-power nulls above are
not obviously attributable to
periodic/aperiodic conflation, though a comparison this blunt (see the minimum detectable
difference quoted in the Results) cannot establish that they are equivalent.

The predictors
acted on the *level* learners reached, with no detectable purchase on the *rate*. The
predictor-by-session interactions were weak, with the statistical-learning-by-session
term the only one approaching credibility in the raw-band models, and only in the
gender-agreement model. That model's
rate is the least reliably estimated quantity here (reliability
`r relv("p2_predictive_trajectory_gender", "z_session_time")`), so, per the rule stated in the Results, we do not interpret it. In the aperiodic re-parameterisation of that stratum, the adjusted-alpha-by-session term comes at least as close, and we do not interpret it either. The attainment-not-rate asymmetry is not an artefact of the priors' design, a small informative mass on
attainment shrunk towards zero on rate, since it held when the rate terms were re-fit under
weakly informative priors (see *Priors*).

The overall pattern is the one a circumscribed advantage would produce. Had a broad aptitude driven learning, it should have appeared in the statistical-learning index, which is reliably measured and was assessed and modelled identically. In the event, the signal was concentrated in working memory, and the projection-predictive selection retained nothing beyond it. The neural and inhibitory-control comparisons carry much less weight in that argument than the modelling alone suggests: the low-frequency bands are weakly identified, the gamma predictor is not a valid measure, and the Stroop index is too unreliable to have registered an association (see *Limitations and future directions*). The contrast the data can draw is therefore between working memory and statistical learning, the two indices measured well enough to be compared.

The joint model's chief advantage is that it separates attainment from rate. Estimating the by-participant
intercept and slope inside one multilevel model, with their interactions with the predictors,
propagates the uncertainty a two-stage slope analysis would discard [@gelman2005;
@humberg2024]. The by-participant level was estimated with high reliability and the pooled
learning rate less so
(`r relv("p2_predictive_trajectory", "z_session_time")` against `r relv("p2_predictive_trajectory", "Intercept")`),
with the gender-only rate more modest again (`r relv("p2_predictive_trajectory_gender", "z_session_time")`),
so rate-based claims from that model warrant more caution [@hedge2018]. This asymmetry
bears on how aptitude findings indexed by a per-learner slope should be read. If the rate
is in general the less reliably estimated quantity (as it was here even under trial-level
joint modelling, the most favourable case), then slope-indexed predictor effects will be
attenuated and unstable in proportion to that unreliability. Disagreement among them is
then to be expected [@hedge2018; @humberg2024].

## Change in executive function over training

A behavioural change also emerged over training. Working-memory span rose between the first and fifth
sessions, while inhibitory control and statistical learning did not. This comparison lacks a
no-training control group, so on its own a pre/post gain is open to the familiar threats of
an uncontrolled retest: practice effects, regression to the mean, maturation and expectancy
[@shadish2002; @calamia2012; @barnett2005]. Two features of the analysis bound those threats.
First, the gain appeared concentrated in one index. Of three indices measured and retested
identically, only working memory moved. In a joint model across the battery, the pre/post change was larger for digit span than for the co-administered Stroop and statistical-learning measures, in the
same direction and with probabilities of direction of
`r pd_of("p2_cognition_prepost_joint", "session_prepost:measurestroop_interference")`% and
`r pd_of("p2_cognition_prepost_joint", "session_prepost:measureasrt_learning")`%. Those specificity contrasts do
not themselves clear the credibility threshold, since their 95% intervals include zero, and
the joint-model digit-span change does not either, so the concentration remains provisional. The
two co-administered measures nonetheless serve as internal controls. The threats of an
uncontrolled retest act broadly across a battery, so a change concentrated in a single
index is not what they predict. Both controls are imperfect. The Stroop index is of
very low internal consistency here, which leaves it a noisy comparator, and its polarity runs
opposite to digit span's, which inflates that contrast (see *Part B: change in executive
function over training*). Second, that same working-memory measure was the one baseline predictor of
overall accuracy, so the predictive and pre/post analyses converge on it. None of this licenses a
causal claim. Practice effects are not uniform across tasks [@scharfen2018], and even active
control conditions do not fully rule out differential expectancy [@boot2013; @simons2016], so
attributing the gain to the multilingual training specifically would require a matched
no-training group. The result is consistent with a training-related, working-memory-specific change, and no more than that.

## Limitations and future directions

The first limitation is precision. Although the models are fitted to tens of thousands of
trials, the predictor effects are *between*-participant quantities, so their effective sample
size is the number of participants. Only
`r proj_n(projpath_raw)` participants contribute between-participant variance to the
raw-band model and `r proj_n(projpath_aper)` to the aperiodic variant, and eighteen
between-participant coefficients (nine main effects and nine interactions with time) are
estimated from them. That is why even the clearest predictor effect is reported with a wide
interval, and it bounds how strongly any single predictor claim can be made. Effects that do
clear a 95% interval at this sample size are also, on average, overestimates of their true
magnitude, so the working-memory coefficient should be read as evidence of a positive association, not as a calibrated estimate of its size.

A second limitation is that the study is silent on *neural* plasticity. Resting-state EEG was recorded only at
baseline, so the rs-EEG measures bear only on the prediction question. Establishing whether intensive multilingual training reshapes
resting-state activity would need a matched post-training recording, and the behavioural
pre/post comparison reported above cannot stand in for one. A change in working-memory span
is a change in what the task measures, and says nothing about whether the resting rhythms
underlying that capacity have shifted.

A third limitation is the reliability of the predictors themselves, which bounds any
association they can enter into. Because reporting the reliability of the outcome while
leaving the predictors unquantified would be one-sided, we estimated the internal
consistency of each baseline index by permutation split-half with a Spearman--Brown
correction. The result separates two measures that are usually discussed together as
difference scores [@hedge2018]. Statistical learning is estimated highly reliably
(`r predrel("asrt_learning")`) and working-memory span adequately so
(`r predrel("digit_span")`), but Stroop interference is not. Its split-half reliability is
`r predrel("stroop_interference")`, which is close to zero. Almost none of its
between-participant variance is systematic, so the null association between inhibitory control and learning reported here is uninformative, and no
association could have been detected with this measure however large the true effect. That
applies to Part A and to Part B, where the Stroop index is a noisy internal control on the
pre/post comparison. The statistical-learning null carries no such excuse. The ASRT is
estimated highly reliably, so its near-zero coefficient cannot be put down to measurement
error. Even so, failing to
detect an association is not the same as demonstrating its absence. Under a prior centred above zero, the pooled
estimate is `r report_effect(LES_P2_MODELS[["predictive"]], "z_asrt")`. The interval's
upper bound is close to the working-memory median we treat as our clearest effect, so these
data cannot exclude a statistical-learning association of the size they attribute to
working memory. What they establish is that no such association was detected on a
well-measured index, which is why it, and not the neural or inhibitory-control nulls, is the
comparator the specificity argument rests on.

A further scope condition is that no language-specific aptitude test was administered.
All predictors index domain-general capacities, so the contrast our data can draw is
between a single circumscribed capacity and a broad domain-general advantage. It does not
extend to the language-specific aptitude of the Modern Language Aptitude Test (MLAT) and
LLAMA tradition [@li2015]. Whether such an aptitude measure would outpredict working
memory here is an open question.

Finally, the artificial-language design trades ecological validity for control. The gain is
a grammar no participant has met before, so prior exposure cannot confound the
learning curve, and that control is what makes the individual-difference question
tractable at all. The cost is that the three properties enter at different points and
differ in difficulty, so the pooled trajectory averages over learners who are at different
stages of different problems. The pooled analysis is therefore best read alongside the
gender-only one, restricted to the single property sampled from the first grammar session
onwards, which the predictor pattern survived.

## Conclusion

Two lines of evidence point to working memory. It was the one baseline index that predicted
how much grammatical knowledge learners acquired, and the one index that rose across the
training period. Statistical learning, measured at least as reliably, did neither, and the
inhibitory-control measure was too unreliable here to test either way. Both results are
therefore about a single capacity, though the evidence that the pre/post gain is confined
to it stops short of being decisive.

# Data and code availability

All extraction, modelling and reporting scripts, the computational environment
specification and the machine-written summary tables that populate this manuscript are
available in this paper's repository at
<https://github.com/pablobernabeu/LESS-cognitive-predictors>, together with the online
supplementary material. The raw electroencephalographic recordings are too large to
distribute through that repository and are deposited on OSF at <https://osf.io/tq7vy>. The
supplement holds
the baseline predictor distributions and the resting-state spectra, which are descriptive
companions to the analyses reported here and are not needed to interpret them. The
environment specification has two parts, because no single artefact covers the whole of it:
an `renv` lockfile pinning the R packages of the fitting library, snapshotted on the cluster,
and a provenance record written at fit time. The compiler and the optimisation pin described
above belong to neither, being properties of the cluster module, which
is why they are stated in prose. The provenance record covers the software versions in force
when the Part A predictor models were fitted, so those fits can be traced to the environment
that produced them. The pre/post models, the projection-predictive selection and the
cross-validation comparisons run from the same environment, but they write no record of
their own.
`r if (is.null(.prov_tbl) || !nrow(.prov_tbl)) "That record has not yet been written for this paper, which is why the version fields under *Sampler, diagnostics and inference* print as pending. They populate from it automatically once the Part A predictor models are next fitted." else paste0("It was written when those models were fitted, on ", substr(.prov_tbl$recorded_utc[1], 1, 10), ", and it supplies the version fields under *Sampler, diagnostics and inference*.")`

{{< pagebreak >}}

# Convergence diagnostics {#apx-convergence-diagnostics-p2}

Every model whose posterior estimates are reported in the main text was assessed against the same convergence criteria. @tbl-convergence gives the realised per-model values behind the summary in the Results. The common-sample refits behind the cross-validation comparisons are not listed. They exist only to score the two parameterisations on identical observations, carry no reported coefficient and do not pass through the diagnostic artefact.

```{r}
#| label: tbl-convergence
#| tbl-cap: "Convergence Diagnostics for Every Model Whose Posterior Estimates Are Reported"
#| apa-note: "Columns give the maximum rank-normalised split R-hat and the minimum bulk and tail effective sample sizes, each taken over all parameters. Part A models come first, then the Part B pre/post models. The weak-prior rows are the prior-sensitivity refits described under Priors."
if (!is.null(conv_tbl) && requireNamespace("knitr", quietly = TRUE)) {
  # Row labels: the CSV carries pipeline identifiers, which are not reader-facing, and it
  # is written in whatever order the summaries job produced. Map to prose names and order
  # rows to follow the manuscript (Part A, then Part B). An identifier with no entry in
  # the map keeps its stripped form rather than being dropped, so a future model still
  # appears.
  .cml <- c(
    predictive_trajectory                     = "Part A: pooled",
    predictive_trajectory_weakprior           = "Part A: pooled (weak priors)",
    predictive_trajectory_gender              = "Part A: gender agreement",
    predictive_trajectory_gender_weakprior    = "Part A: gender agreement (weak priors)",
    predictive_trajectory_gender_iafres       = "Part A: gender agreement, IAF-resolved participants only",
    predictive_trajectory_aperiodic           = "Part A: pooled, aperiodic",
    predictive_trajectory_aperiodic_gender    = "Part A: gender agreement, aperiodic",
    cognition_prepost_digit_span              = "Part B: digit span",
    cognition_prepost_stroop_interference     = "Part B: Stroop interference",
    cognition_prepost_asrt_learning           = "Part B: ASRT learning",
    cognition_prepost_joint                   = "Part B: joint across the three indices",
    cognition_prepost_joint_signaligned       = "Part B: joint across the three indices, sign-aligned")
  conv_tbl <- conv_tbl[order(match(sub("^p2_", "", conv_tbl$model), names(.cml)),
                             conv_tbl$model), , drop = FALSE]
  # Reordering leaves the source row numbers as row names, which kable would print as an
  # unlabelled leading column of pipeline row indices.
  rownames(conv_tbl) <- NULL
  ct <- conv_tbl |>
    dplyr::transmute(Model = ifelse(is.na(.cml[sub("^p2_", "", model)]),
                                    sub("^p2_", "", model),
                                    unname(.cml[sub("^p2_", "", model)])),
                     `max R-hat` = round(max_rhat, 3),
                     `min ESS (bulk)` = round(min_ess_bulk),
                     `min ESS (tail)` = round(min_ess_tail),
                     Divergences = n_divergent,
                     Passed = ifelse(passed, "yes", "no"))
  # Drop a column that carries no information. Every model here has zero divergent
  # transitions and every one passes, so those two columns were a wall of "0" and "yes"
  # restating the sentence above the table, and they were what pushed the header past the
  # measure. They are dropped only when they are genuinely constant, so a future run in
  # which a model diverges or fails still shows it rather than hiding it.
  if (all(!is.na(ct$Divergences)) && all(ct$Divergences == 0)) ct$Divergences <- NULL
  if (all(ct$Passed == "yes"))                ct$Passed      <- NULL
  knitr::kable(ct)
} else {
  cat("*Convergence diagnostics will appear here once the models are fitted.*")
}
```

# Reliability of the baseline predictors {#apx-predictor-reliability}

@tbl-predictor-reliability gives the split-half internal consistency of the three baseline cognitive indices. The values, and what they imply for the inhibitory-control null, are discussed under *Limitations and future directions*.

```{r}
#| label: tbl-predictor-reliability
#| tbl-cap: "Internal-Consistency Reliability of the Three Baseline Cognitive Predictors at Session 1"
#| apa-note: "Reliability is estimated by permutation split-half: each participant's trials are repeatedly split in half, the index is recomputed on each half, the halves are correlated across participants, and the mean correlation is corrected to full test length by the Spearman-Brown formula. Two of the three indices (Stroop interference, ASRT learning) are difference scores."
if (!is.null(predrel_tbl) && requireNamespace("knitr", quietly = TRUE)) {
  .prl <- c(digit_span = "Working memory (digit span)",
            stroop_interference = "Inhibitory control (Stroop interference)",
            asrt_learning = "Statistical learning (ASRT)")
  # Correlations and reliabilities are bounded by 1, so their leading zeros are stripped,
  # matching predrel() and relv() in the running prose.
  .strip0 <- function(x) sub("^(-?)0\\.", "\\1.", formatC(x, format = "f", digits = 2))
  pr <- predrel_tbl |>
    dplyr::transmute(Predictor = unname(.prl[predictor]),
                     `Participants` = n_participants,
                     `Mean half-test r` = .strip0(r_half),
                     `Reliability (Spearman-Brown)` = .strip0(reliability_sb),
                     `Random splits` = n_splits)
  knitr::kable(pr)
} else {
  cat("*Predictor reliability will appear once `scripts/11_extract_predictor_reliability.R` has run.*")
}
```

# Supplementary materials {#apx-supplementary}

The two displays below are supplied to the journal as online supplementary material
and are deposited with the analysis code in this paper's repository. They are
descriptive companions to the intercorrelations and numerical summaries in the main
text, so they are not required to interpret the findings, and they fall outside the
five figures of the article proper. @fig-predictor-dist shows the distribution of each
baseline predictor over the stored records for that measure, and @fig-rs-psd the
resting-state spectra behind the band-power and individual-alpha-frequency predictors.

```{r}
#| label: fig-predictor-dist
#| fig-cap: "Distributions of the Baseline Predictors"
#| apa-note: "Violins with individual points show working memory (digit span), Stroop interference and ASRT statistical-learning scores, and the eyes-closed resting-state band powers and individual alpha frequency. Each panel is drawn from the stored records for that measure, so the number of points differs between panels and exceeds the model samples reported under Participants. Difference-score predictors (Stroop, ASRT) are shown on their raw scale, and each panel's unit is given in its label. Box plots mark the median and interquartile range. The absolute band powers are strongly right-skewed. They nonetheless enter the models z-scored on this raw scale, with no transform applied. Points at the 7 Hz floor of the individual-alpha-frequency panel are values at the lower bound of the estimator's search range, not necessarily measured peaks (see Cognitive and neural indices)."
#| fig-width: 6.5
#| fig-height: 4.4
pred_vals <- .read_csv_if(paper2_results("_predictor_values.csv"))
if (!is.null(pred_vals)) {
  ord <- c("digit_span", "stroop_interference", "asrt_learning",
           "delta", "theta", "alpha", "beta", "gamma", "iaf")
  pred_vals$predictor <- factor(pred_vals$predictor, levels = ord)
  # Units differ across predictors and the y scales are free, so each panel states its
  # own unit: the cognitive indices are a count and two reaction-time differences, the
  # band powers a power integrated over frequency, and the individual alpha frequency a
  # frequency. The band unit is uV^2, NOT uV^2/Hz: 03_extract_resting_state_eeg.R defines
  # band power as the trapezoidal integral of the PSD over the band, which cancels the
  # per-Hz denominator. fig-rs-psd, which plots the density itself, keeps uV^2/Hz.
  pred_lab <- as_labeller(c(
    digit_span = "Digit span (items correct)", stroop_interference = "Stroop (ms)",
    asrt_learning = "ASRT (ms)", delta = "Delta (µV²)",
    theta = "Theta (µV²)", alpha = "Alpha (µV²)",
    beta = "Beta (µV²)", gamma = "Gamma (µV²)",
    iaf = "IAF (Hz)"))
  ggplot(pred_vals, aes(x = "", y = value)) +
    geom_violin(fill = "grey90", colour = "grey60", linewidth = 0.3) +
    geom_jitter(width = 0.12, size = 0.5, alpha = 0.5, colour = "grey25") +
    geom_boxplot(width = 0.10, fill = "white", colour = "grey30", linewidth = 0.3, outlier.shape = NA) +
    facet_wrap(~ predictor, scales = "free_y", nrow = 3, labeller = pred_lab) +
    labs(x = NULL, y = NULL) + les_theme() +
    theme(axis.text.x = element_blank(), panel.grid.major.x = element_blank())
} else cat("*Predictor distributions will appear once `_predictor_values.csv` is present.*")
```

```{r}
#| label: fig-rs-psd
#| fig-cap: "Grand-Average Resting-State Power Spectral Density Over the Eight Occipito-Parietal Electrodes"
#| apa-note: "Eyes-closed is shown against eyes-open. Both axes are logarithmic, so the aperiodic 1/f background appears as a straight line. The curve is the geometric mean across every recorded resting-state spectrum, a wider frame than the fitted models use, and the shaded ribbon is the interquartile range. Grey lines mark the interior delta/theta/alpha/beta/gamma band edges (4, 8, 13, 30 Hz), and the alpha band (8--13 Hz) is highlighted. The darker region above 30 Hz lies beyond the offline low-pass corner: the nominal gamma band (30--45 Hz) falls entirely within it, which is why that predictor is not interpreted. The eyes-closed posterior alpha increase (alpha-blocking on eye opening) is the expected Berger effect, which serves here as a check on the recording and the reduction."
#| fig-width: 6.5
#| fig-height: 3.4
if (!is.null(psd_tbl)) {
  ga <- psd_tbl |>
    dplyr::group_by(condition, freq_hz) |>
    dplyr::summarise(lo = stats::quantile(power, .25, na.rm = TRUE),
                     hi = stats::quantile(power, .75, na.rm = TRUE),
                     power = exp(mean(log(power), na.rm = TRUE)),
                     .groups = "drop")
  cond_lab <- c(eyes_closed = "Eyes closed", eyes_open = "Eyes open")
  ggplot(ga, aes(freq_hz, power, colour = condition)) +
    annotate("rect", xmin = 8, xmax = 13, ymin = 0, ymax = Inf, alpha = 0.08, fill = "grey40") +
    # Shade the region above the 30 Hz offline low-pass corner. The nominal gamma band
    # (30-45 Hz) lies entirely inside it, so the figure shows directly why that predictor
    # is not interpretable rather than leaving the point to the text.
    annotate("rect", xmin = 30, xmax = Inf, ymin = 0, ymax = Inf, alpha = 0.13, fill = "grey55") +
    annotate("text", x = 37, y = Inf, label = "above low-pass", vjust = 1.6,
             size = 2.8, colour = "grey30") +
    geom_vline(xintercept = c(4, 8, 13, 30), colour = "grey80", linewidth = 0.2) +
    geom_ribbon(aes(ymin = lo, ymax = hi, fill = condition), alpha = 0.15, colour = NA) +
    geom_line(aes(linetype = condition), linewidth = 0.6) +
    scale_x_log10(breaks = c(1, 2, 4, 8, 13, 20, 30, 45), labels = c(1, 2, 4, 8, 13, 20, 30, 45)) +
    scale_y_log10() +
    scale_colour_manual(values = c(eyes_closed = "#D55E00", eyes_open = "#0072B2"), labels = cond_lab) +
    scale_fill_manual(values = c(eyes_closed = "#D55E00", eyes_open = "#0072B2"), guide = "none") +
    scale_linetype_manual(values = c(eyes_closed = "solid", eyes_open = "22"), labels = cond_lab) +
    labs(x = "Frequency (Hz)", y = expression(Power~(mu*V^2/Hz)), colour = NULL, linetype = NULL) +
    les_theme() + theme(legend.position = "top")
} else cat("*Resting-state spectra will appear once `_resting_state_psd.csv` is present.*")
```
