---
title: "Cognitive and Resting-State EEG Predictors of Intensive Multilingual Learning, and Change in Executive Function across Training: A Bayesian Multilevel Study"
shorttitle: "Predictors and Plasticity in Multilingual Learning"
author:
  - name: Pablo Bernabeu
    corresponding: true
    orcid: 0000-0003-1083-2460
    email: pcbernabeu@gmail.com
    affiliations:
      # The affiliation that carries the postal address is listed first because the two
      # outputs build the correspondence note differently: the Word and HTML route takes
      # the corresponding author's FIRST affiliation (frontmatter.lua), while the PDF
      # route takes whichever affiliation carries an `address` (title.tex). Listing
      # Oxford first is what makes them agree, and Bilingualism: Language and Cognition
      # requires the corresponding author's postal address on the title page.
      - id: oxford
        name: University of Oxford
        department: Department of Education
        address: 15 Norham Gardens
        city: Oxford
        postal-code: OX2 6PY
        country: United Kingdom
      - id: clabl
        name: UiT The Arctic University of Norway
        department: Center for Language, Brain and Learning (C-LaBL)
        city: Tromsø
        country: Norway
    roles:
      - role: Conceptualization
      - role: Methodology
      - role: Software
      - role: Formal analysis
      - role: Data curation
      - role: Visualization
      - role: Writing – original draft
      - role: Writing – review & editing
  - name: Christina Athanasiadi
    affiliations:
      - ref: clabl
    roles:
      - role: Data curation
      - role: Project administration
      - role: Writing – review & editing
  - name: Stella L. Pischinger
    affiliations:
      - ref: clabl
    roles:
      - role: Data curation
      - role: Writing – review & editing
  - name: Gabriella Silva
    orcid: 0000-0002-5349-7036
    affiliations:
      - ref: nebrija
    roles:
      - role: Conceptualization
      - role: Methodology
      - role: Investigation
      - role: Writing – review & editing
  - name: My Ngoc Giang Hoang
    affiliations:
      - ref: clabl
    roles:
      - role: Data curation
      - role: Writing – review & editing
  - name: Vincent DeLuca
    orcid: 0000-0002-2275-209X
    affiliations:
      - ref: clabl
    roles:
      - role: Conceptualization
      - role: Methodology
      - role: Supervision
      - role: Writing – review & editing
  - name: Jason Rothman
    orcid: 0009-0003-3936-6384
    affiliations:
      - id: nebrija
        name: Nebrija University
        department: Nebrija Research Center in Cognition (CINC)
        city: Madrid
        country: Spain
      - id: lancaster
        name: Lancaster University
        department: Department of Linguistics and English Language
        city: Lancaster
        country: United Kingdom
      - ref: clabl
    roles:
      - role: Conceptualization
      - role: Funding acquisition
      - role: Resources
      - role: Supervision
      - role: Writing – review & editing
  - name: Iva Ivanova
    affiliations:
      - id: utep
        name: University of Texas at El Paso
        department: Department of Psychology
        city: El Paso
        region: Texas
        country: United States
    roles:
      - role: Conceptualization
      - role: Methodology
      - role: Writing – review & editing
  - name: Claudia Poch
    orcid: 0000-0002-9357-4948
    affiliations:
      - ref: nebrija
    roles:
      - role: Methodology
      - role: Validation
  - name: Jorge González Alonso
    orcid: 0000-0001-5047-3226
    affiliations:
      - ref: clabl
      - ref: nebrija
    roles:
      - role: Conceptualization
      - role: Methodology
      - role: Funding acquisition
      - role: Project administration
      - role: Supervision
      - role: Writing – review & editing
author-note:
  correspondence-note: >
    Correspondence concerning this article should be addressed to Pablo Bernabeu,
    Department of Education, University of Oxford, 15 Norham Gardens, Oxford OX2 6PY,
    United Kingdom, Email: pcbernabeu@gmail.com; or Jorge González Alonso, Nebrija
    Research Center in Cognition (CINC), Nebrija University, Calle de Asura 90, 28043
    Madrid, Spain, Email: jgonzalezal@nebrija.es.
  disclosures:
    study-registration: "The study design, tasks and session protocol were preregistered on the Open Science Framework (https://osf.io/tjr54)."
    data-sharing: "Data, code and materials are available at the locations given under Data and code availability."
    related-report: "The electrophysiological and behavioural signatures of transfer in the same sample are reported in a companion manuscript submitted to the Journal of Neurolinguistics."
    conflict-of-interest: "Competing interests: The authors declare none."
    financial-support: "This work was supported by the Research Council of Norway (grant 326487-FORSKER21) and by the Comunidad de Madrid (grant 2020-T1/HUM-20037), both awarded to Jorge González Alonso."
    gratitude: >-
      We thank Mona Kirkness Fossum for assistance with translating materials;
      Gaute Berglund, Anders Gabrielsen and Tekabe Legesse Feleke for assistance with session
      piloting; Teodóra Vékony for advice on the alternating serial reaction
      time task; Björn Lundquist for assistance with implementation in OpenSesame; and
      Antonio Fábregas, Natalia Mitrofanova, Yulia Rodina, Marit Westergaard, Merete
      Anderssen and Fatih Bayram for advice and assistance with the design and translation of
      materials and participant recruitment. Models were fitted on High-Performance Computing
      (HPC) clusters at Sigma2 (Norway) and the Advanced Research Computing facility of the
      University of Oxford.
      Generative AI statement: the authors used Claude Code (version 2.1, Anthropic),
      running primarily the Claude Opus and Claude Fable models and accessed through
      Anthropic's Claude service (https://claude.com/claude-code) between July 2026 and
      September 2026, to support the research, the analyses and the writing. It drafted
      and revised text and code and checked the manuscript against its analysis artefacts.
      The authors reviewed all of its output and take responsibility for the content.
    credit-introduction: "Author roles were classified using the Contributor Role Taxonomy (CRediT; https://credit.niso.org/) as follows:"
    credit-authors:
      - name: Pablo Bernabeu
        roles: "Conceptualization, Methodology, Software, Formal analysis, Data curation, Visualization, Writing – original draft, Writing – review & editing."
      - name: Christina Athanasiadi
        roles: "Data curation, Project administration, Writing – review & editing."
      - name: Stella L. Pischinger
        roles: "Data curation, Writing – review & editing."
      - name: Gabriella Silva
        roles: "Conceptualization, Methodology, Investigation, Writing – review & editing."
      - name: My Ngoc Giang Hoang
        roles: "Data curation, Writing – review & editing."
      - name: Vincent DeLuca
        roles: "Conceptualization, Methodology, Supervision, Writing – review & editing."
      - name: Jason Rothman
        roles: "Conceptualization, Funding acquisition, Resources, Supervision, Writing – review & editing."
      - name: Iva Ivanova
        roles: "Conceptualization, Methodology, Writing – review & editing."
      - name: Claudia Poch
        roles: "Methodology, Validation, Writing – review & editing."
      - name: Jorge González Alonso
        roles: "Conceptualization, Methodology, Funding acquisition, Project administration, Supervision, Writing – review & editing."
abstract: |
  Adults vary widely in how well they learn a new language. We related baseline working
  memory, inhibitory control, statistical learning and resting-state
  electroencephalography (EEG) to grammaticality-judgement accuracy as Norwegian--English
  bilinguals learned one of two artificial languages. We also tested change in executive
  function across training. Trial-level accuracy was modelled jointly in one Bayesian
  multilevel model, without per-learner derived slopes. Working memory was positively associated with overall
  judgement accuracy, credibly so in the fully sampled gender-agreement property and
  with a pooled interval that narrowly included zero. Inhibitory control and statistical
  learning showed no credible association, the intercorrelated band powers could not be
  resolved individually, and the projection-predictive selection retained no neural
  measure. Working-memory span rose between the first and fifth sessions, so the
  capacity that predicted learning also changed over training. The by-participant level
  was estimated more reliably than the learning rate, which bears on how slope-indexed
  aptitude findings are read.
keywords: [second language acquisition, working memory, statistical learning, resting-state EEG, individual alpha frequency, Bayesian multilevel models, individual differences]
bibliography: references.bib
floatsintext: true
format:
  apaquarto-pdf:
    documentmode: man
    pdf-engine: lualatex
    include-in-header:
      - text: |
          % apaquarto's template decrements the table counter at the end of every
          % longtable that carries no caption of its own (apatemplate.tex, "Disable
          % longtable counter"). Quarto now places the caption on an enclosing table
          % float and marks the longtable \def\LTcaptype{none}, so no longtable
          % carries its own caption any more and that decrement cancels the increment
          % the \caption just made. Without this, every table prints as "Table 1" and
          % every cross-reference to a table resolves to 1 -- in the PDF only; the
          % HTML and DOCX outputs number correctly. Restore the lost increment.
          \makeatletter
          \AtBeginDocument{%
            \apptocmd{\endlongtable}{\ifLTpatch@caption\else\addtocounter{table}{1}\fi}{}{}%
          }
          \makeatother

          % --- Vertical rhythm -------------------------------------------------
          % Defaults left captions crowded against their table or figure, opened a
          % gap above and below every display equation that broke the flow of the
          % surrounding sentence, and put more space under a heading than over it,
          % which detaches the heading from the text it introduces.
          % caption.sty is already loaded by the apa7 class, so only its settings are
          % adjusted here; loading it again risks an option clash.
          \captionsetup{aboveskip=10pt, belowskip=10pt}
          % Keep multi-line displays compact without changing the manuscript's
          % double spacing in ordinary prose.
          \AtBeginDocument{%
            \setlength{\jot}{0pt}%
            \setlength{\abovedisplayskip}{4pt}%
            \setlength{\belowdisplayskip}{4pt}%
            \setlength{\abovedisplayshortskip}{2pt}%
            \setlength{\belowdisplayshortskip}{2pt}%
          }
          % Keep paragraph spacing stable rather than letting apa7's flushbottom
          % setting create page-specific gaps in a double-spaced manuscript.
          \raggedbottom
          % Heading spacing is left to the apa7 class. The titlesec package cannot parse
          % that class's redefined heading levels: it fails at \ttl@extract\subparagraph
          % with 'Argument of \subparagraph has an extra }', and the document does not
          % build with it loaded. The spacing the class applies is APA's own, which is
          % what a journal expects, and the roomy look in manuscript mode comes from the
          % required double spacing rather than from the heading skips.

          % --- Keep wide tables inside the text block --------------------------
          % A table whose natural width exceeds the measure silently runs into the
          % margin rather than being scaled; column headers then collide (the
          % convergence table printed "DivergencesPassed" and ran the longest model
          % name into the next column). Shrink any over-wide table to the measure.
          \usepackage{etoolbox}
          \usepackage{graphicx}
          \usepackage{needspace}
           \AtBeginEnvironment{longtable}{\footnotesize\setlength{\tabcolsep}{3pt}}
          \makeatletter
          \g@addto@macro\@floatboxreset{\centering}
          \makeatother
    # cairo_pdf embeds the fonts it uses and renders the Unicode minus, micro and
    # multiplication signs the figures contain; R's default pdf() device references
    # Helvetica without embedding it, which submission systems flag.
    knitr:
      opts_chunk:
        dev: "cairo_pdf"
  apaquarto-html: default
  apaquarto-docx:
    fig-dpi: 600
execute:
  echo: false
  warning: false
  message: false
editor: source
---

```{r setup}
#| include: false
# =============================================================================
# Manuscript setup -- shared paths, results loading, inline-reporting helpers.
# Render-anytime: before the HPC models exist it prints clearly-marked placeholders;
# once results/_pooled_*.csv are present every number is populated from those
# artefacts (Marwick, Boettiger & Mullen, 2018, https://doi.org/10.1080/00031305.2017.1375986).
# The results loading and the reporting helpers live in manuscript_setup.R, sourced at
# the end of this chunk, so that supplement.qmd renders from the same definitions.
# =============================================================================
# Figure text size, set once and inherited by every figure through les_theme(). Each
# figure is authored at the width of the text block, so nothing is scaled on placement and
# a point here is a point on the page. While several figures were authored at 8 inches and
# placed in a 6.5-inch measure, the 9 pt previously set here arrived as roughly 7.3 pt,
# below what is legible in print.
LES_FIG_BASE_SIZE <- 10

suppressPackageStartupMessages({
  library(here)
  source(here::here("_shared", "R", "00_paths.R"))
  source(here::here("_shared", "R", "05_funnel.R"))   # les_funnel(), shared with Paper 1
  source(here::here("paper_2_plasticity", "scripts", "_config.R"))
  # les_theme(), les_div, the number formatters, .read_csv_if(), prov(), flow_n() and
  # pstat(), shared with Paper 1 so both manuscripts report by one set of rules.
  source(here::here("_shared", "R", "06_manuscript_helpers.R"))
  library(dplyr)
  library(ggplot2)
  # depictr supplies the figure theme and colour system, shared with the companion
  # paper so both carry one visual identity. Its default qualitative palette is the
  # colourblind-safe Okabe-Ito set. Optional, per the render-anytime design.
  has_depictr <- requireNamespace("depictr", quietly = TRUE)
  if (has_depictr) {
    library(depictr)
    depictr_options(base_size = LES_FIG_BASE_SIZE)
  }
  source(here::here("_shared", "R", "07_task_durations.R"))
})

source(here::here("paper_2_plasticity", "manuscript_setup.R"))
```

Adults differ markedly in the speed and the ultimate success with which they learn a
new language, and explaining that variation is a central goal of the cognitive
neuroscience of language.
Three families of predictor recur in the literature. Working memory is the
best-evidenced domain-general cognitive correlate of second-language outcomes,
language-specific aptitude aside [@li2015], with a small-to-moderate positive
meta-analytic association [@linck2014] that is stronger for executive and complex-span
measures than for simple storage span and, for vocabulary learning, stronger for nonword
repetition than for digit span [@kurokawa2026]. Statistical or implicit learning, the
ability to extract distributional regularities, relates positively but only weakly to
language ability across 97 studies, with the size of the correlation depending on the
task and the stimulus modality [@zhou2024], and nonadjacent statistical learning has
predicted comprehension of particular native-language sentence types [@misyak2012]. Two
artificial-language training studies that tracked their predictors across five sessions
found the contribution of statistical learning concentrated in the earliest exposure,
after which explicit learning aptitude [@kenanidis2026] or cognitive control [@chen2022]
carried more of the variance. Inhibitory control is the least settled. The
bilingual-advantage literature finds no reliable executive-function benefit of
bilingualism [@paap2013; @dick2019], and direct tests of inhibition against
second-language outcomes have returned null associations, for phonological processing
[@huensch2024], for classroom learning gains [@linckweiss2015] and for
artificial-grammar learning, where a flanker index of inhibition predicted at no stage
[@chen2022].

Alpha and theta dynamics in the resting electroencephalogram (EEG) index memory and
cognitive performance
[@klimesch1999]. The individual alpha frequency is a stable, trait-like measure
[@grandy2013trait] that has correlated positively with general cognitive ability
[@grandy2013], so a positive association with learning is the direction the
cross-sectional literature would suggest. That association did not hold in a larger
sample, where the frequency shared variance with processing speed alone [@ociepka2022],
and the prior placed on it here is not directional (see *Priors*). Baseline
oscillatory power has
also predicted the rate of adult second-language learning directly, with faster learners
distinguished mainly by higher-frequency (beta and low-gamma) rhythms [@prat2016;
@prat2019]. In the later of those studies, only the coherence measures survived
correction for multiple comparisons, so the band-power precedent rests chiefly on the
earlier one and on a 10-participant pilot in older adults, in which resting low-beta
power predicted progress relative to the room left for improvement at baseline
[@kliesch2021]. We therefore enter five canonical bands (delta, theta, alpha, beta,
gamma) and the individual alpha frequency as baseline predictors. That precedent
implicates higher-frequency bands over right temporoparietal cortex, whereas our
reduction is over posterior electrodes, so the bands enter without a directional prior
(see *Priors*). Although intensive language learning can reshape the underlying neural
architecture [@li2014; @martensson2012], the EEG here is measured only at baseline and
used to predict subsequent learning.

Raw band power is, moreover, an ambiguous quantity. It conflates genuine oscillatory
activity with the broadband aperiodic ($1/f$) background on which the peaks sit, and the
two components carry different information [@donoghue2020]. The aperiodic exponent has
been proposed as a non-invasive proxy for cortical excitation--inhibition balance
[@gao2017], varies systematically across individuals and with age [@voytek2015], and can
drive associations that appear on the surface to belong to band power [@ouyang2020].
Separating the two components therefore gives a test that can fail in either direction.
An apparent band-power association should survive it, and an association carried by the
aperiodic background should appear only once it is made. We run both.

A question cuts across all three families: whether an aptitude expresses itself in how
fast learners improve or in the level they reach. Most individual-difference studies
index learning by a per-learner slope estimated in a first stage, a design that treats an
error-laden slope as an error-free quantity (see *Learning outcome and model structure*).
We therefore model rate and level jointly at trial level and ask which of the two the
predictors act on.

This study brings these strands together in a controlled, longitudinal
artificial-language paradigm. Norwegian--English bilinguals learn one of two miniature
languages across a series of laboratory sessions. At baseline, we measure working
memory, inhibitory control, statistical learning and resting-state EEG, and across the
training sessions we track the developing accuracy of grammaticality judgements. Designs
that predict a controlled training outcome from a baseline neural
recording remain uncommon, and the longitudinal precedents [@prat2016; @prat2019;
@prat2020; @kliesch2021; @kliesch2022] indexed learning by a derived rate or gain score.
The present study models the full trial-level trajectory jointly, separating the level
reached from the rate of change and quantifying the reliability of each, and it asks the
neural question of both the raw band powers and their aperiodic-adjusted decomposition,
so that which characterisation of the resting spectrum the data prefer is itself an
empirical question.

Two questions follow. Part A asks whether the baseline cognitive and neural
measures predict the rate and attainment of learning. Part B asks whether executive
function changes between the first and fifth sessions, and whether any change is confined
to one index or broad across the battery. Part A bears on a question that recurs in the
individual-differences literature [@linck2014; @li2015]: whether success in adult
grammar learning reflects a broad cognitive advantage or a single capacity. Because no
language-specific aptitude test was administered, the contrast drawn here is between one
domain-general capacity and several (see *Limitations and future directions*). A broad
advantage should be evident across a battery of cognitive and neural measures assessed
and modelled identically. A single capacity would leave one predictor associated with
overall accuracy and its co-measured companions unrelated to it. Which memory or control
ability predicts grammar learning has been found to depend on the stage of learning, in
an artificial language [@morganshort2014], and on the learning condition and the
difficulty of the rule, in second-language grammar [@rivera2023]. The question is
therefore put here to a battery assessed once at baseline and modelled across the whole
trajectory, and the two accounts are separable to the extent that every predictor enters
the same joint model under priors of comparable scale, subject to the reliability of each
measure (see *Limitations and future directions*). Part B has a stake of its own. Whether cognitive capacities improve with demanding training is contested. Apparent
gains often turn out to reflect retest artefacts and expectancy [@boot2013; @simons2016],
and for language learning specifically an 11-week randomised trial in older adults
found no gain in working memory, memory or intelligence over a relaxation control
[@berggren2020]. The three indices, administered together at both sessions, are used as
internal controls on one another, because generic retest and expectancy effects act
broadly across a battery.

# Method

## Participants

The participants are those of the companion study [@lesstransfer2026], drawn from the
longitudinal artificial-language project reported by @gonzalezalonso2025:
Norwegian--English bilinguals (first language Norwegian, second language English) who
learned one of two miniature languages. `r if (nzchar(minor_age())) paste0("One participant was aged ", minor_age(), " at enrolment according to the language-history questionnaire, and all others were adults.") else ""` Of the
`r flow_n("enrolled")` enrolled, `r flow_n("usable_accuracy")`
contributed grammaticality-judgement trajectories
(`r flow_n("usable_accuracy", "n_mini_english")` Mini-English,
`r flow_n("usable_accuracy", "n_mini_norwegian")` Mini-Norwegian) and
`r flow_n("battery_baseline")` a baseline cognitive battery. Of these,
`r flow_n("joint_predictor")` had both measures and were therefore candidates for the
joint predictor model, and `r flow_n("usable_rseeg")` also had usable baseline
resting-state EEG. Because the
joint model requires complete data on every predictor, the fitted raw-band model
comprises `r proj_n(projpath_raw)` of those participants. The difference is one
participant with usable resting-state EEG and no baseline Stroop score, whom listwise
deletion removes. The aperiodic variant
described below comprises `r proj_n(projpath_aper)` participants, the subset for which
the specparam re-parameterisation is defined. That decomposition returns no individual
alpha frequency for a spectrum in which it fits no alpha peak, and listwise deletion then
removes those participants too. This loss is therefore not independent of the alpha
measures themselves. The Session-1-versus-5 pre/post models of
executive function use all participants contributing data at either session
(`r prepost_ns_phrase()`), with the session effect estimated within participants through
the by-participant intercept. Participants tested only once inform the intercept but
contribute no within-participant change. @fig-flow traces
attendance and the resulting analysis samples across the six sessions.

Of the
`r pstat("n_lhq3", "value", 0)` participants with a Language History Questionnaire record
[LHQ3; @lhq3], the mean age was `r pstat("age_years")` years
($SD = `r pstat("age_years", "sd")`$, range `r pstat("age_years", "vmin", 0)`--`r pstat("age_years", "vmax", 0)`).
Of that group, `r pstat("sex_female", "value", 0)` were female,
`r pstat("sex_male", "value", 0)` male and
`r pstat("sex_nonbinary", "value", 0)` non-binary, and `r pstat("sex_undisclosed", "value", 0)`
did not disclose their gender. Most were right-handed
(`r pstat("hand_right", "value", 0)` of `r pstat("hand_reported_n", "value", 0)` who
reported handedness). English had been acquired early (age of first exposure
$M = `r pstat("eng_aoa_listen")`$ years) and used for many years, with high self-rated
 proficiency ($M = `r pstat("eng_l2_proficiency", "value", 2)`$ on a 0--1 scale). The project
 was registered with the Norwegian Agency for Shared Services in Education and Research
 (Sikt; project 2559611). All participants gave written informed consent. The authors assert that all
procedures contributing to this work comply with the ethical standards of the relevant
national and institutional committees on human experimentation and with the Helsinki
Declaration of 1975, as revised in 2008.

```{r}
#| label: fig-flow
#| fig-cap: "Participant Flow"
#| apa-note: "Each stage spans the full enrolled cohort. The solid block is the participants still present and the faint block completing the bar is those no longer present, so cumulative attrition reads as filled against unfilled area. Tapering bands connect consecutive stages, each transition annotated with the number lost and the percentage retained. In the upper tier, the solid block is split by mini-language. The predictor-model stages are not recorded by language and are drawn in neutral grey. The upper tier tracks attendance across all six sessions, since this paper draws on all of them: the cognitive battery at Sessions 1 and 5, the resting-state recording at Session 2, and the learning trajectory across the four judgement sessions. The lower tier is the predictor-model sample proper, which is nested, and its final stage is the sample the raw-band model is fitted to, after listwise deletion on the full predictor set. Bands are drawn only within a tier, because the analysis samples pool across sessions and are not a subset of final-session attendance. The repository's flow artefacts reproduce and cross-check every count shown."
#| fig-width: 6.5
#| fig-height: 5.4
if (!is.null(flow_tbl)) {
  fl <- flow_tbl[!is.na(flow_tbl$n_total), ]
  gv <- function(st) { v <- fl$n_total[fl$stage == st]; if (length(v)) v[1] else NA_integer_ }
  att <- fl[grepl("^(enrolled|attended)", fl$stage), ]
  att <- att[order(att$order), ]
  d <- rbind(
    data.frame(label = sub("\\s*\\(Paper [12]\\)", "", att$label), n = att$n_total,
               n_a = att$n_mini_english, n_b = att$n_mini_norwegian,
               tier = "Enrolment and attendance", note = NA_character_,
               stringsAsFactors = FALSE),
    data.frame(
      label = c("Contributed a learning trajectory",
                "...and a usable baseline battery",
                "...and usable baseline resting-state EEG",
                "...and complete data on every predictor (fitted raw-band model)"),
      # The final stage is the model's own fitted sample, read from the projpred
      # solution path rather than from the flow artefact, which records candidacy
      # stages and not the listwise deletion the joint model applies. Without it the
      # funnel stopped one stage short of the sample every Part A estimate rests on.
      n     = c(gv("trajectory"), gv("joint_predictor"), gv("usable_rseeg"),
                suppressWarnings(as.integer(proj_n(projpath_raw)))),
      n_a = NA_integer_, n_b = NA_integer_,
      tier = "Predictor-model sample",
      note = c(NA_character_,
               sprintf("%s had a usable baseline battery in total", gv("battery_baseline")),
               NA_character_,
               sprintf("the aperiodic variant is fitted to %s", proj_n(projpath_aper))),
      stringsAsFactors = FALSE))
  les_funnel(d, base_size = LES_FIG_BASE_SIZE)
} else {
  cat("*Participant-flow diagram will appear once `_sample_flow.csv` is present.*")
}
```

## Cognitive and neural indices

The cognitive battery was administered at Sessions 1 (baseline) and 5 (post) as
browser-based tasks on the Gorilla experiment platform [@anwylirvine2020], and records
from individuals who completed the battery without enrolling were excluded at extraction,
having no consent linkage to the participant key. Working memory was measured by digit
span (forward and backward, list lengths 3--8 with four items each,
`r task_spec("digit_span", "scored_trials")` scored trials, the index being the number
correct). Inhibitory control was measured by a four-colour Stroop task [@stroop1935]
(`r task_spec("stroop", "test_trials")` test trials,
`r task_spec("stroop", "congruent_trials")` of them congruent, the index being the
incongruent-minus-congruent mean reaction time, so that higher scores index poorer
inhibitory control). Statistical learning was
measured by an alternating serial reaction-time (ASRT) task [@howard1997], the
alternating-sequence variant of the serial reaction-time paradigm [@nissen1987], of
`r task_spec("asrt", "n_blocks")` blocks of `r task_spec("asrt", "trials_per_block")`
trials in which pattern and random trials alternate. The index is the
low-minus-high-probability-triplet mean reaction-time difference, with higher scores
indexing stronger learning. Reaction times were trimmed to 50--5000 ms and within-cell
outliers beyond three standard deviations were removed, with both indices averaged over
correct and incorrect responses alike. The Stroop index includes the practice block,
which the export records in the same response zone as the test trials and which no
filter removes at extraction. The trimming cells and the minimum trial counts are given
in the online supplementary material (Section S1).

Resting-state EEG (rs-EEG) was recorded once, at Session 2, with a 32-channel Brain
Products LiveAmp (500 Hz) in separate eyes-closed and eyes-open recordings of about five
minutes each, whose order was counterbalanced between participants. Preprocessing in
BrainVision Analyzer comprised visual inspection,
re-referencing to the average of the mastoids, zero-phase Butterworth filtering (0.1--30
Hz, no notch) and ocular correction by independent component analysis. The full five
minutes was retained for every participant in both conditions, and the acquisition and
preprocessing details are given in the online supplementary material (Section S2). Across
the laboratory sessions, participants were instructed to minimise movement and blinking,
especially during the presentation of words. During pauses in these long sessions,
experimenters monitored the signal and intervened when necessary to maintain low electrode
impedance and acceptable signal quality. Power
spectral density was estimated over eight occipito-parietal electrodes (O1, Oz, O2, P3,
Pz, P4, P7, P8) by Welch's method [@welch1967] and integrated over delta (1--4 Hz), theta
(4--8 Hz), alpha (8--13 Hz), beta (13--30 Hz) and gamma (30--45 Hz). The individual alpha
frequency (IAF) was the frequency of maximum power in the 7--13 Hz range [@klimesch1999],
an argmax for which @corcoran2018 recommend a smoothed alternative, so the specparam IAF
recovered in the aperiodic decomposition serves as the cross-check.

Because the one suggestive neural result rests on this measure, one failure mode of the
argmax matters here. For a spectrum that merely decays across the search range it returns
the lower bound, so a value at 7 Hz is not necessarily a peak. The argmax lies strictly
inside the range for `r rseeg_peak("resolvable")` of the
`r rseeg_peak("recorded")` eyes-closed recordings, and the aperiodic decomposition
described below fits a peak within that window in `r specparam_peak_n()` of the
`r specparam_rec_n()`. For the recordings in
which the decomposition fits no such peak, the individual alpha frequency should be read
as unresolved. Two of them are pinned at the 7 Hz floor with no fitted peak, and the
others carry interior argmax values. `r iafres_sentence()`

The gamma band (30--45 Hz) lies entirely above the 30 Hz corner of the offline low-pass,
so `r gamma_att_clause()`. The power observed there far exceeds what the filter would
transmit of the aperiodic background (`r gamma_att_obs_clause()`), consistent with mains
energy at 50 Hz persisting through a filter applied without a notch. The gamma predictor
is therefore not a valid index of gamma-band activity. It is retained only so that the
predictor set is not altered after the fact, we do not interpret it, and its null result
is not evidence about gamma-band activity. The transmission analysis is given in the
online supplementary material (Section S3).

The eyes-closed values serve as the neural predictors, since the alpha rhythm and its peak
are most reliably estimated with the eyes closed. The reduction reproduces the expected
posterior alpha-blocking on eye opening, a check on the recording whose absence would
have indicated a problem with the montage or the reference. All continuous predictors
were $z$-scored, so each coefficient is expressed per standard deviation of its predictor
and the predictors are directly comparable with one another [@brauer2018;
@schielzeth2010]. The cognitive predictors were standardised over every participant with
a Session 1 battery and the resting-state predictors over every eyes-closed recording, so
a unit is a standard deviation of the measured sample, a slightly wider frame than the
fitted one.

The observed elapsed duration of each home task and of each phase of the laboratory
sessions is shown in @fig-home-task-duration and @fig-lab-phase-duration. These are
descriptive timing checks, not additional outcomes: home-task duration runs from the
first to the last Gorilla event for each participant and task, whereas lab-phase duration
uses the OpenSesame phase markers and the final event in a phase when the next phase was
recorded in a separate export.
`r les_task_duration_median_text(task_duration_home, task_duration_lab)`

```{r}
#| label: fig-home-task-duration
#| fig-cap: "Duration of the Home Cognitive Tasks by Session"
#| apa-note: "Each box summarises the distribution across participants of elapsed time from the first to the last raw Gorilla event for one task and session; the centre line is the median, the box spans the interquartile range, and whiskers extend to 1.5 times the interquartile range. Sessions 1 and 5 are baseline and post-training, respectively. Durations therefore include instructions, practice and test material when those events are present in the export. The ordinate is logarithmic because five exports contain multi-hour idle gaps; those observations remain as outliers. Participants without a linked enrolled-participant key are excluded."
#| fig-width: 6.5
#| fig-height: 2.9
les_task_duration_home_plot(task_duration_home)
```

```{r}
#| label: fig-lab-phase-duration
#| fig-cap: "Duration of the Laboratory-Session Phases"
#| apa-note: "Each box summarises the distribution across participants of elapsed time in one laboratory-session phase and session; the centre line is the median, the box spans the interquartile range, and whiskers extend to 1.5 times the interquartile range. Phase duration is measured between the recorded phase-start markers, or from a phase start to the last phase-specific event when the next phase was exported separately. Sessions 2, 3 and 4 are training sessions and Session 6 is the retention session."
#| fig-width: 6.5
#| fig-height: 3.0
les_task_duration_lab_plot(task_duration_lab)
```

## Learning outcome and model structure (Part A)

The learning outcome is the longitudinal trajectory of grammaticality-judgement accuracy,
modelled jointly at trial level. The two-stage alternative most aptitude studies take, a
slope fitted per learner and then regressed on the predictors, treats a slope estimated
from few noisy trials as if it were measured without error, ignores the heterogeneous
precision of those estimates and uses the data inefficiently [@gelman2005]. Any
standardised association is attenuated in proportion to the slopes' unreliability
[@hedge2018], and where estimated slopes serve as predictors, as in the designs
@humberg2024 evaluate, the attenuation becomes a regression dilution of the coefficient
itself. We therefore fit a single Bayesian multilevel Bernoulli model of trial-level
accuracy. The learning *rate* is a by-participant random slope for session and the
predictors enter as interactions with session [@barr2013], so that under partial pooling a
learner whose slope is poorly determined contributes proportionately less to the predictor
estimate [@gelmanhill2007]:
$$
\begin{aligned}
\text{correct} \sim\; & \text{Session} \times \text{Property} \\
 & + \text{Session} \times (\text{WM} + \text{Inhibition} + \text{StatLearning} + \text{rs-EEG}) \\
 & + (1 + \text{Session} \mid \text{participant}).
\end{aligned}
$$

The model carries no by-item term. The judgement sentences are generated combinatorially
from each grammar, so each distinct sentence is presented to only a few participants and
item variance is thinly sampled, as the companion paper sets out [@lesstransfer2026].
That variance is absorbed into the residual.

Session was coded as a continuous standardised variable so that the rate is a single
slope. The four judgement sessions enter at equal steps, although Sessions 2 to 4 were
one week apart and Session 6 was an unrehearsed retention test about four months after
Session 4 (see the companion paper), so the slope is per session step, whatever the
elapsed time between sessions. Session was standardised over the pooled trial frame
before any listwise deletion, so one unit is close to one step of the session index and
zero falls at the trial-weighted mean session, which the cumulative design places
between Sessions 3 and 4. The gender-agreement model inherits that scaling, so its main
effects and its intercept are evaluated at the same session index, which lies later than
that stratum's own mean session. Grammatical
property entered as a three-level factor under treatment coding with differential object
marking, the first level alphabetically, as the reference, so the population-level
session term is that property's slope and the two session-by-property terms are the
other properties' departures from it. The predictor terms do not interact with property,
so their estimates do not depend on this choice. A linear effect of session is the
parsimonious default. A generalised additive form [@winterwieling2016; @soskuthy2017] was
held in reserve, since averaging across learners can distort the apparent form of a
learning curve [@heathcote2000; @murre2011], and it was not fitted. Because linearity
was not formally tested, the session term is interpreted as a linear approximation to
the trajectories in @fig-accuracy-trajectory, where the gender-agreement trajectories in
particular are not monotone.

We report both the learning rate (the predictor-by-session
interactions) and the attainment (the model-implied end-of-training accuracy), and assess
the reliability of the by-participant rate and of the by-participant level of accuracy,
the random intercept, before interpreting either as an individual difference
[@hedge2018; @hedge2021]. Attainment is the population-level predicted accuracy at
Session 6 for a learner at the mean of every predictor and at zero on both random
effects, averaged with equal weight over the three properties in the pooled model, so it
is a conditional prediction and carries no by-participant reliability. The reliabilities
are single-occasion, model-based separation reliabilities estimated within this sample,
computed as the variance of the shrunken by-participant posterior means relative to the
model's between-participant variance, $\mathrm{Var}(\hat{u}_i)/\tau^2$. They are not
test-retest reliabilities, which would require a second measurement wave
[@rouderhaaf2019].
Because session is standardised, a predictor's main effect is its association with
accuracy at the pooled-frame mean session. We label the main effects as overall accuracy
and reserve *attainment* for the model-implied end-of-training quantity. Pooled
accuracy is shaped by properties of differing difficulty entering cumulatively across
sessions: gender agreement from Session 2, differential object marking from Session 3 and
verb--object number agreement from Session 4 (see @lesstransfer2026 for the design).
Part A is therefore additionally fitted restricted to gender agreement, the property with
full session coverage, so that the predictor pattern can be read on an outcome free of the
cumulative design. The pooled model is the primary analysis and the gender-agreement
model its companion.

## Pre/post models of executive function (Part B)

Part B fits a Gaussian pre/post
model of each cognitive index (Session 1 versus Session 5) with a by-participant intercept and a
$\pm0.5$ pre/post contrast. Because the study has no no-training control group, a pre/post change is confounded with the generic threats that attend any uncontrolled retest [@shadish2002]. We add a joint model across the three $z$-scored indices in which each acts as an internal control for the others. With digit span as the reference level, the
session effect is its pre/post change and each measure-by-session term is how much less that
measure changed. The model can therefore identify a change concentrated in one index and set it
against the broad shift that practice, regression to the mean or expectancy would produce.
The three indices are standardised within measure and enter on their recorded polarity,
which is not aligned across them: higher digit span is better performance, higher Stroop
interference worse. A generic improvement therefore moves the two in opposite signed
directions, so the digit-span-versus-Stroop contrast confounds specificity with that
opposition and should be read as an upper bound on the separation. A sign-aligned refit,
in which the Stroop index is reversed before standardisation so that a positive change
means improvement on all three indices, is reported alongside it. Because resting-state
EEG was recorded only at baseline, no neural pre/post is estimated.

## Priors

Priors are informative and standardised on the logit scale, established from the
literature reviewed in the Introduction and following a selective, evidence-tiered scheme. The cited behavioural associations are *attainment* correlations, so each small informative prior is placed on a predictor's main effect and shrunk towards zero on its interaction with session. Working memory and statistical learning receive small
positive priors [@linck2014; @kurokawa2026; @zhou2024; @misyak2012]. Inhibitory control
receives a near-zero regularising prior, reflecting its weak-to-null evidence base
[@huensch2024; @linckweiss2015; @chen2022], and the resting-state EEG predictors receive
near-zero regularising priors. The band-power precedent [@prat2016; @prat2019] was
obtained over a different scalp region and for a different outcome, learning rate, so no
directional prior is encoded for the bands, and a null obtained under regularising priors
at this sample size cannot count against that precedent. Variance components receive
generous
half--Student-$t(3,0,2.5)$ priors and correlations an $\mathrm{LKJ}(2)$ prior
[@gelman2006; @lewandowski2009]. In a sensitivity analysis, the Part A predictors were
re-fit under weakly informative priors, $\mathcal{N}(0, 1.5)$ on the intercept and on
every population-level coefficient, with the variance and correlation priors unchanged
(see Results). The Part B pre/post models use weakly informative priors throughout
[@schad2021; @depaoli2017]. In brief, the Part A intercept has a $\mathcal{N}(1, 1)$
prior on the logit scale, any population-level coefficient without a term-specific prior
has $\mathcal{N}(0, 1)$ [@gelman2008], and the term-specific priors are small positive
normals for working memory and statistical learning and near-zero normals for inhibitory
control and every resting-state predictor. The full specification is given in the online
supplementary material (Section S6).

```{r prior-spec-check}
#| results: asis
# Render-time consistency check for the prior specification stated in prose above.
# The fit scripts are the source of truth; once the cluster job writes
# results/_prior_specification.csv (brms::prior_summary() rows per model), this chunk
# compares the artefact against the prose values, kept here in one named vector, and
# prints a visible warning marker on any divergence. Silent when the artefact is
# absent or everything matches.
.prior_spec_tbl <- .read_csv_if(paper2_results("_prior_specification.csv"))
if (!is.null(.prior_spec_tbl) &&
    all(c("class", "coef", "prior") %in% names(.prior_spec_tbl))) {
  .norm <- function(x) gsub("\\s+", "", tolower(x))
  # The prose values for the Part A pooled model: coefficient-specific slabs, the
  # intercept, and the default population-level prior (empty coef).
  # ".default" stands for the class-b prior with no coefficient (an empty coef in
  # prior_summary()); an empty string cannot be used as a name in a named vector.
  .expected_b <- c(
    "z_digit_span"                = "normal(0.1, 0.25)",
    "z_session_time:z_digit_span" = "normal(0.05, 0.15)",
    "z_asrt"                      = "normal(0.1, 0.2)",
    "z_session_time:z_asrt"       = "normal(0.05, 0.15)",
    "z_stroop"                    = "normal(0, 0.15)",
    "z_session_time:z_stroop"     = "normal(0, 0.1)",
    "z_alpha"                     = "normal(0, 0.25)",
    "z_session_time:z_alpha"      = "normal(0, 0.15)",
    ".default"                    = "normal(0, 1)")
  ps <- .prior_spec_tbl
  if ("model" %in% names(ps)) ps <- ps[ps$model == LES_P2_MODELS[["predictive"]], , drop = FALSE]
  .mism <- character(0)
  for (cf in names(.expected_b)) {
    is_default <- identical(cf, ".default")
    rows <- ps[ps$class == "b" &
                 (if (is_default) (is.na(ps$coef) | ps$coef == "") else ps$coef == cf), ,
               drop = FALSE]
    if (nrow(rows) < 1) next  # term absent from the artefact; nothing to compare
    if (!identical(.norm(rows$prior[1]), .norm(.expected_b[[cf]])))
      .mism <- c(.mism, sprintf("`%s`: fitted `%s` vs prose `%s`",
                                if (is_default) "(default b)" else cf,
                                rows$prior[1], .expected_b[[cf]]))
  }
  ri <- ps[ps$class == "Intercept", , drop = FALSE]
  if (nrow(ri) >= 1 && !identical(.norm(ri$prior[1]), .norm("normal(1, 1)")))
    .mism <- c(.mism, sprintf("Intercept: fitted `%s` vs prose `normal(1, 1)`", ri$prior[1]))
  if (length(.mism))
    cat("\n\n**[PRIOR-SPECIFICATION MISMATCH -- the fitted priors differ from the",
        " specification stated in the supplementary material: ", paste(.mism, collapse = "; "),
        ". Reconcile before submission.]**\n\n", sep = "")
}
```

## Sampler, diagnostics and inference

Models were fitted with `brms` [@burkner2017; @burkner2018] as a front end to Stan
[@carpenter2017], using the same sampler configuration as the companion paper
[@lesstransfer2026]: four chains
of 10,000 iterations each, the first 2,000 discarded as warm-up, giving 32,000 post-warm-up
draws, with `adapt_delta` = 0.95 and `max_treedepth` = 12. The convergence gate was
rank-normalised $\widehat{R} < 1.01$, bulk- and tail-ESS above 400, and zero divergent
transitions [@vehtari2021]. Posterior predictive checks for the Part A models were
written to the repository at fit time [@gabry2019], and inference was by posterior
medians, 95% credible intervals and
the probability of direction ($p_d$) [@makowski2019; @kruschke2018], every credible
interval being equal-tailed (2.5th to 97.5th percentiles of the draws). An ESS of 400
guarantees trustworthy diagnostics without guaranteeing the stability of an interval
limit [@kruschke2021; @vehtari2021]. As in the companion paper,
we describe an effect as credible when its 95% credible interval excludes zero, and
otherwise read the probability of direction as a continuous index of directional
certainty, never as a dichotomous test. The projection-predictive and aperiodic analyses
narrow the predictor space, and we apply no blanket multiplicity correction, which would
inflate Type II error without an inferential gain [@rubin2021]. Binary accuracy is
modelled with a Bernoulli (logit) likelihood throughout. Analyses were carried out in R
`r prov("R")`
[@rcoreteam] with `brms` `r prov("brms")` as a front end to CmdStan `r prov("CmdStan")`,
`projpred` `r prov("projpred")` for the projection-predictive selection and `loo`
`r prov("loo")` for the Pareto-smoothed importance-sampling leave-one-out (PSIS-LOO)
comparison, under the fixed seed `r prov("seed:LES_SEED", "20260607")`, in a
reproducible `here()`-anchored compendium [@marwick2018]. The models were fitted on a
Linux computing cluster, the versions given above are those of the fitting run, and the
platform, the compiler settings and the figure palette are given in the online
supplementary material (Section S7). The analysis and reporting code was written
with the assistance of Claude Code (version 2.1, Anthropic), running primarily the
Claude Opus and Claude Fable models, accessed through Anthropic's Claude service
(<https://claude.com/claude-code>) between July 2026 and September 2026. The tool
drafted and revised scripts under the authors' direction and checked them against the
method reported here. The authors reviewed and ran every script, and the deposited code
reproduces the analyses without the tool.

## Predictor selection and a 1/f-adjusted robustness check

Two further analyses probe the neural predictors. First, raw resting-state band power
conflates oscillatory activity with the aperiodic $1/f$ background, itself a substantive
individual difference [@donoghue2020; @ouyang2020]. We therefore re-parameterised each
participant's baseline spectrum with a specparam decomposition [@donoghue2020] into an
aperiodic component (exponent and offset), $1/f$-adjusted theta, alpha and beta power and
the specparam individual alpha frequency, the centre of the tallest fitted peak between 7
and 13 Hz inclusive. Part A was then re-fit with these quantities in place of the raw
bands.
The decomposition is our own R implementation, fitted over 2--40 Hz in fixed (knee-free)
mode, and its settings and the consequences of the 30 Hz low-pass for the fit are given
in the online supplementary material (Section S4), where they can be judged against the
recommendations of @gerster2022. Only tall peaks pass the threshold of two standard
deviations of the flattened spectrum (Section S4), so a theta peak was fitted in
`r n_peak_band("theta_adj")`, a peak within the individual-alpha window in
`r specparam_peak_n()` and a beta peak in `r n_peak_band("beta_adj")` of the
`r specparam_rec_n()` eyes-closed recordings. The adjusted theta and beta predictors are
therefore close to constant, and we do not interpret their coefficients or their
positions in the selection path. This aperiodic variant is a robustness check of the
raw-band model, specified after data collection like the rest of the analysis plan (see
*Transparency and preregistration*), and it is not a second pass over the predictor
space.

The two reference models are compared by leave-one-out cross-validation [@vehtari2017].
Because they differ only in participant-level predictors, which a held-out trial's own
participant intercept already absorbs, they are also compared by $K$-fold cross-validation
with folds blocked on participants ($K = 10$, the same folds for both models, participants
dealt round-robin over a seeded shuffle), the remedy @roberts2017 recommend wherever such
dependence exists. We report the difference in expected log predictive density (elpd) with a standard error
clustered on participants, summing the pointwise contributions within participant before
taking the standard error across participants, and read a difference as decisive only
when it exceeds twice that standard error. Why the trial-level criterion cannot settle
the comparison on its own, and the naive standard error alongside, are given in the
online supplementary material (Section S5).

Second, the joint Part A model retains every candidate predictor for inference, and
projection-predictive variable selection [@piironen2017; @piironen2020] asks which subset
accounts for its predictive performance by projecting the reference posterior onto
nested submodels, each inheriting the reference model's regularisation. Because the
reference model is a multilevel Bernoulli model with a correlated random slope, we use
the latent projection [@catalina2021; @mclatchie2025], which fits each projected submodel
on the linear-predictor scale and evaluates its performance on the response scale (the
reason the traditional projection fails for this model is given in Section S5). The
forward search is cross-validated in full over `r projpred_nfolds_word()`
participant-grouped folds [@roberts2017], and the suggested submodel size is projpred's
standard heuristic, the smallest size whose cross-validated elpd lies within one standard
error of the reference model's [@piironen2020]. A predictor that the minimal submodel
does not retain is not thereby shown to be null. The selection summarises which
predictors account for the reference model's predictive performance, and the full Part A
posterior remains the inferential statement [@piironen2020]. The raw-band and aperiodic
reference models' selections are reported side by side.

## Transparency and preregistration

The study design, materials and session protocol were preregistered before data
collection (<https://osf.io/tjr54>), with no statistical analysis plan, so the Bayesian
multilevel models, the informative priors, the specparam/aperiodic re-parameterisation
and the projection-predictive selection reported here were specified after the data were
collected and are exploratory [@nosek2018; @wagenmakers2012]. The preregistration
motivated the Session-1 battery as an index of attention relevant to the selection of a
source of transfer, and reframing the same three tasks as working-memory,
inhibitory-control and statistical-learning measures departs from that rationale. Data
collection was preregistered for two sites, Norway and Spain, and only the Norwegian
cohort (first-language Norwegian, second-language English) is analysed here, so the
findings generalise only to first-language-Norwegian learners. Finally, the preregistered
protocol required participants to score above 80% on a post-training behavioural test in
each grammar session before they could proceed. After a few participants had been tested,
we recognised that this criterion was too strict and modified the procedure: when the
first test score was below 80%, pretraining, training and the test were repeated; after
the second attempt, participants proceeded to the experiment regardless of their score.
Participants were shown their percentage correct at the end of each test attempt. The 80%
threshold therefore describes the first-attempt gate, not a requirement imposed on the
second attempt. The gate shapes which participant-sessions contribute learning data and
is part of the attrition described above.

# Results

```{r convergence}
#| echo: false
#| results: asis
# The static lead-in and the generated count are emitted together, as one paragraph. Split
# across a sentence of prose and a chunk they rendered as two paragraphs of one sentence
# each, which reads as a stub.
if (!results_ready) {
  cat("::: {.callout-note title=\"Reproducible draft\"}\n",
      "Model fits are not yet present. Every figure and statistic below populates ",
      "automatically once the pooled result files are produced by the modelling ",
      "pipeline.\n:::\n", sep = "")
} else {
  np <- sum(conv_tbl$passed, na.rm = TRUE)
  # The divergence clause is built conditionally so a future run with divergent
  # transitions cannot print a self-contradictory paragraph.
  div_clause <- if (all(!is.na(conv_tbl$n_divergent)) && all(conv_tbl$n_divergent == 0)) {
    "and no divergent transitions"
  } else {
    sprintf("and %d divergent transitions in total (see @tbl-convergence)",
            sum(conv_tbl$n_divergent, na.rm = TRUE))
  }
  cat(sprintf(paste0(
    "Convergence diagnostics for every model reported below are given in ",
    "@tbl-convergence, in the appendix, as realised values, since a realised ",
    "value distinguishes a model that converged comfortably from one that converged ",
    "marginally. All %d models were assessed against the same gate and %d passed it, ",
    "with rank-normalised $\\widehat{R}$ below 1.01, effective sample sizes above 400 ",
    "in both the bulk and the tail, %s.\n"),
    nrow(conv_tbl), np, div_clause))
}
```


```{r}
#| results: asis
# The pooled convergence artefact is written per fitted model, so any model that has not
# been run is silently absent rather than shown as failing. State that explicitly instead
# of letting the table imply the grid is complete.
if (!is.null(conv_tbl)) {
  .expected <- unname(c(LES_P2_MODELS[c("predictive", "predictive_gender",
                                        "predictive_aperiodic", "predictive_aperiodic_gender")],
                        .prepost_id("digit_span"), .prepost_id("stroop_interference"),
                        .prepost_id("asrt_learning"), LES_P2_MODELS[["cog_prepost_joint"]]))
  .missing <- setdiff(.expected, conv_tbl$model)
  if (length(.missing))
    cat("\nDiagnostics are not yet available for ", length(.missing),
        " of the models reported elsewhere in this paper (",
        paste0("`", .missing, "`", collapse = ", "),
        "), whose fits predate the current diagnostic artefact.\n", sep = "")
}
```

## Descriptive overview

Before the models, @fig-accuracy-trajectory plots the raw learning trajectory, the direct
check on the linear-in-session assumption the models make, @fig-predictor-corr shows the
baseline predictors' intercorrelations and @tbl-predictor-desc gives the same predictors
numerically. The per-measure distributions (Figure S1) and the resting-state spectra
behind the band-power and individual-alpha-frequency measures (Figure S2) are given in
the online supplementary material (Section S8). The frames of these displays are wider
than the `r proj_n(projpath_raw)` participants of the fitted raw-band model and are stated
in their notes.

One feature of the learning trajectory bears on how *rate* and *attainment* should be read,
since both terms would otherwise suggest improvement. Mean accuracy is highest
at each property's first session and lower at its last, in every property and both
mini-languages. The session term is credibly negative in both models. In the pooled
model the slope for the reference property, differential object marking, is
`r report_effect(LES_P2_MODELS[["predictive"]], "z_session_time")`, and the
gender-agreement and verb--object number-agreement slopes differ from it by
`r report_effect(LES_P2_MODELS[["predictive"]], "z_session_time:grammatical_propertygender_agreement")`
and
`r report_effect(LES_P2_MODELS[["predictive"]], "z_session_time:grammatical_propertyverb_object_number_agreement")`.
Within gender agreement fitted alone the slope is
`r report_effect(LES_P2_MODELS[["predictive_gender"]], "z_session_time")`. The pooled
decline partly reflects the cumulative design, in which later sessions sample harder
material (see *Learning outcome and model structure*). Within gender agreement, which is
sampled throughout, it is an unexplained feature of these data. Throughout, *rate* is the
slope of this trajectory and *attainment* its model-implied end point, whichever way the
trajectory runs.

```{r}
#| label: fig-accuracy-trajectory
#| fig-cap: "Observed Grammaticality-Judgement Accuracy Across Sessions, by Grammatical Property and Mini-Language"
#| apa-note: "Thin lines are individual participants, and the ribbon is a 95% confidence interval computed between participants, the sampling unit being participants and not trials. This is the raw trajectory the joint model summarises, and the direct check on the linear-in-session assumption. No judgements were collected at Session 5. The segment from Session 4 to Session 6 spans the unrehearsed retention interval of about four months, whereas Sessions 2 to 4 were one week apart, so the axis is not proportional to elapsed time."
#| fig-width: 6.5
#| fig-height: 3.4
if (!is.null(acc_traj_tbl)) {
  gc <- acc_traj_tbl[acc_traj_tbl$level == "group_cell" & !is.na(acc_traj_tbl$grammatical_property), ]
  bp <- acc_traj_tbl[acc_traj_tbl$level == "participant_cell" &
                       !is.na(acc_traj_tbl$grammatical_property), ]
  if (!nrow(bp)) bp <- NULL
  # Order the panels by the session in which each property was introduced (gender
  # agreement in Session 2, differential object marking in Session 3, verb-object
  # number agreement in Session 4). The column is character, so without an explicit
  # factor the panels would fall alphabetically and scramble the design. This must
  # precede the ggplot()/geom_line() calls below, which capture their data by value.
  .prop_order <- c("gender_agreement", "differential_object_marking",
                   "verb_object_number_agreement")
  prop_lab <- as_labeller(c(gender_agreement = "Gender agreement",
                            differential_object_marking = "Differential object marking",
                            verb_object_number_agreement = "Verb-object number agreement"))
  gc$grammatical_property <- factor(gc$grammatical_property, levels = .prop_order)
  if (!is.null(bp)) bp$grammatical_property <- factor(bp$grammatical_property,
                                                      levels = .prop_order)
  p <- ggplot(gc, aes(session, mean_accuracy, colour = mini_language))
  if (!is.null(bp)) p <- p + geom_line(data = bp,
                                       aes(session, mean_accuracy,
                                           group = interaction(participant_lab_ID, mini_language)),
                                       colour = "grey80", linewidth = 0.2, alpha = 0.5)
  p + geom_ribbon(aes(ymin = mean_accuracy - 1.96 * se, ymax = mean_accuracy + 1.96 * se,
                      fill = mini_language), alpha = 0.2, colour = NA) +
    geom_line(aes(linetype = mini_language), linewidth = 0.7) + geom_point(size = 1) +
    # Only the four judgement sessions carry a tick: a tick at Session 5, where no
    # judgements were collected, would imply a measured point there.
    scale_x_continuous(breaks = c(2, 3, 4, 6)) +
    facet_wrap(~ grammatical_property, labeller = prop_lab) +
    # Okabe-Ito blue and vermillion, the same pair fig-predictor-forest uses, so every
    # hard-coded scale in the paper draws on the colourblind-safe set the Method names.
    scale_colour_manual(values = c(`Mini-English` = "#0072B2", `Mini-Norwegian` = "#D55E00"),
                        aesthetics = c("colour", "fill")) +
    scale_linetype_manual(values = c(`Mini-English` = "solid", `Mini-Norwegian` = "22")) +
    labs(x = "Session", y = "Proportion correct", colour = NULL, fill = NULL, linetype = NULL) +
    les_theme() + theme(legend.position = "top")
} else cat("*Raw accuracy trajectory will appear once `_accuracy_trajectory_descriptives.csv` is present.*")
```

```{r}
#| label: fig-predictor-corr
#| fig-cap: "Pearson Correlations Among the Baseline Predictors"
#| apa-note: "Correlations are computed on the z-scored predictors, over the participants who have both a baseline cognitive battery and an eyes-closed resting-state spectrum, pairwise-complete, so the n behind each cell varies slightly. Only the lower triangle is shown, the redundant mirror and the unit diagonal being omitted. The resting-state band powers are expected to be positively intercorrelated through the shared aperiodic 1/f background, which motivates the specparam robustness check and is why the projpred ranking must be read against this structure."
#| fig-width: 5.6
#| fig-height: 5
if (!is.null(predictor_corr)) {
  ord <- c("digit_span", "stroop_interference", "asrt_learning", LES_P2_RS_MEASURES)
  corr_labs <- c(digit_span = "Digit span", stroop_interference = "Stroop",
                 asrt_learning = "ASRT", delta = "Delta", theta = "Theta",
                 alpha = "Alpha", beta = "Beta", gamma = "Gamma", iaf = "IAF")
  strip0 <- function(x, d = 2) sub("^(-?)0\\.", "\\1.", formatC(x, format = "f", digits = d))
  pc <- predictor_corr
  # Keep one triangle only: drop the redundant upper mirror and the r = 1 diagonal.
  pc <- pc[match(as.character(pc$var1), ord) < match(as.character(pc$var2), ord), ]
  pc$var1 <- factor(pc$var1, levels = ord); pc$var2 <- factor(pc$var2, levels = rev(ord))
  ggplot(pc, aes(var1, var2, fill = r)) +
    geom_tile(colour = "white", linewidth = 0.5) +
    geom_text(aes(label = strip0(r)), size = 3) +
    scale_fill_gradient2(low = les_div[1], mid = les_div[2], high = les_div[3], midpoint = 0,
                         limits = c(-1, 1),
                         labels = function(x) ifelse(x == 0, "0", strip0(x, 1))) +
    scale_x_discrete(labels = corr_labs) + scale_y_discrete(labels = corr_labs) +
    labs(x = NULL, y = NULL, fill = expression(italic(r))) +
    les_theme() +
    # Blank the specific grid elements, not just the parent: a theme that sets
    # panel.grid.major.x/y explicitly (as depictr's does) is not cleared by setting
    # panel.grid alone, and the grid would then show through the masked triangle.
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          panel.grid = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank())
} else cat("*Predictor correlation matrix will appear once `_predictor_correlations.csv` is present.*")
```

```{r}
#| label: tbl-predictor-desc
#| tbl-cap: "Descriptive Statistics for the Baseline Predictors"
#| apa-note: "The frame is the participants who have both a baseline cognitive battery and an eyes-closed resting-state spectrum, which is wider than the sample any Part A model is fitted to, and the $n$ column gives the number contributing to each row. Each row carries its own unit. Because a band power is the spectral density integrated over the band, its unit is squared microvolts, without the per-hertz denominator of a density."
if (!is.null(predictor_desc) && requireNamespace("knitr", quietly = TRUE)) {
  pd <- predictor_desc
  num <- vapply(pd, is.numeric, logical(1)); pd[num] <- lapply(pd[num], round, 2)
  # Map the pipeline's snake_case keys and lowercase headers to reader-facing labels,
  # using the same names the figures use, and carry the unit on each row so the table
  # is readable without the Method. Done defensively: if the artefact's columns ever
  # change, the raw frame is printed rather than the chunk failing.
  .pdl <- c(digit_span = "Working memory: digit span (items correct)",
            stroop_interference = "Inhibitory control: Stroop interference (ms)",
            asrt_learning = "Statistical learning: ASRT difference (ms)",
            delta = "Delta power (microvolts squared)", theta = "Theta power (microvolts squared)",
            alpha = "Alpha power (microvolts squared)", beta = "Beta power (microvolts squared)",
            gamma = "Gamma power (microvolts squared, not interpreted)",
            iaf = "Individual alpha frequency (Hz)")
  if (all(c("predictor", "n", "mean", "sd", "median", "min", "max") %in% names(pd))) {
    lab <- unname(.pdl[pd$predictor])
    # Statistical symbols set in italic (APA 7), and negatives typeset with a minus sign
    # that LaTeX will not break the number at.
    .num <- function(x) sub("^-", "-", formatC(x, format = "f", digits = 2))
    pd <- data.frame(Predictor = ifelse(is.na(lab), pd$predictor, lab),
                     `$n$` = pd$n, `$M$` = .num(pd$mean), `$SD$` = .num(pd$sd),
                     `$Mdn$` = .num(pd$median), Min = .num(pd$min), Max = .num(pd$max),
                     check.names = FALSE)
  }
  knitr::kable(pd, align = c("l", rep("r", ncol(pd) - 1)))
} else if (is.null(predictor_desc)) cat("*Predictor descriptive table pending `_predictor_descriptives.csv`.*")
```

## Part A: Predictors of the learning trajectory

Each predictor enters twice: as a main effect, indexing the overall level of accuracy, and
in interaction with session, indexing the learning *rate*. Because session is standardised,
the main effect is the association at the pooled-frame mean session (see Method). Working
memory was
the clearest predictor. Higher baseline digit span went with higher overall accuracy,
credibly so for the fully sampled gender-agreement property
(`r report_effect(LES_P2_MODELS[["predictive_gender"]], "z_digit_span")`) and positively,
though with an interval that narrowly included zero, in the pooled model
(`r report_effect(LES_P2_MODELS[["predictive"]], "z_digit_span")`). Inhibitory control
(`r report_effect(LES_P2_MODELS[["predictive"]], "z_stroop")`) and statistical learning
(`r report_effect(LES_P2_MODELS[["predictive"]], "z_asrt")`) were not credibly related to
accuracy, in line with their weak-to-null evidence base.

Of the baseline resting-state measures, the band powers yielded no credible coefficient
(@fig-predictor-forest), but that null must be read against how strongly they are
intercorrelated. Entering the bands simultaneously means each coefficient is the unique
contribution of that band with the others held constant, and the four low-frequency bands
share a great deal of variance (@fig-predictor-corr). Their variance inflation factors,
computed over the `r vif_n()` participants with complete predictor data in the
descriptive frame, are
`r vif_range(c("delta","theta","alpha","beta"))`, which widens their posterior intervals by
roughly `r vif_widen(c("delta","theta","alpha","beta"))` relative to an orthogonal
predictor. Their null coefficients are therefore weakly identified and uninformative
about whether those rhythms relate to learning. Two of the neural predictors escape this:
gamma
(VIF `r vif_of("gamma")`) and the individual alpha frequency (VIF `r vif_of("iaf")`) are
nearly uncorrelated with the rest, so their estimates stand on their own. The one
suggestive signal was the individual alpha frequency, whose association with
gender-agreement accuracy was
`r report_effect(LES_P2_MODELS[["predictive_gender"]], "z_iaf")`. The cross-check the
Method promised points the same way without settling it. In the aperiodic
re-parameterisation of the same stratum, fitted to a smaller sample, the
specparam-estimated individual alpha frequency was
`r report_effect(LES_P2_MODELS[["predictive_aperiodic_gender"]], "z_specparam_iaf")`, and the
aperiodic exponent, the quantity the decomposition adds, was
`r report_effect(LES_P2_MODELS[["predictive_aperiodic_gender"]], "z_aperiodic_exponent")`. The learning-*rate*
interactions were uniformly weak (for working memory,
`r report_effect(LES_P2_MODELS[["predictive"]], "z_session_time:z_digit_span")`),
consistent with priors that placed their small informative mass on the main effects and
shrank the rate terms towards zero. The largest was the statistical-learning-by-session
term in the gender-agreement model
(`r report_effect(LES_P2_MODELS[["predictive_gender"]], "z_session_time:z_asrt")`), with
the adjusted-alpha-by-session term of the aperiodic refit of that stratum comparable
(`r report_effect(LES_P2_MODELS[["predictive_aperiodic_gender"]], "z_session_time:z_alpha_adj")`).
Neither is interpreted, for the reason given below.

Under weakly informative priors, the regularising priors on the resting-state bands are
seen to have shrunk them towards zero and tightened their intervals (across all
coefficients, the weakly informative intervals were about `r prior_ci_widening()` wider
at the median). No band-power predictor was credible under either prior. The individual
alpha frequency, however, is prior-sensitive. Under the weakly informative prior, its
gender-agreement interval excluded zero
(`r sens_effect(LES_P2_MODELS[["predictive_gender"]], "z_iaf", "weak")`), whereas under the
regularising prior it did not
(`r sens_effect(LES_P2_MODELS[["predictive_gender"]], "z_iaf", "inf")`). The regularised
estimate is treated as primary, and we read the association as suggestive. The
working-memory main effect was, if anything, larger under the weakly informative prior,
and its pooled interval then excluded zero
(`r sens_effect(LES_P2_MODELS[["predictive"]], "z_digit_span", "weak")`). The
predictor-by-session interactions likewise remained non-credible under weakly informative
priors (largest $p_d$ = `r sens_max_rate_pd()`), so the level-not-rate asymmetry is a
property of the data and does not depend on the prior specification.

```{r p2-selection}
#| results: asis
if (!is.null(projpath_raw)) {
  cat(sprintf(paste0(
    "To establish which predictors account for the reference model's predictive ",
    "performance, projection-predictive selection was run on %s and on the aperiodic ",
    "re-parameterisation, and it returned the same parsimonious answer throughout.%s ",
    "In the raw-band model, the suggested size was %s: the design scaffold ",
    "(by-participant intercept and session slope, session, grammatical property and their ",
    "interaction) together with working memory (digit span). Digit span was the first ",
    "individual-difference predictor selected, entering at step %s of the forward search, ",
    "exactly the suggested size. Inhibitory control (step %s), statistical learning (step %s) ",
    "and the resting-state band predictors entered only beyond the suggested size, so they ",
    "are not among the terms the search retained. The forward search was capped at projpred's ",
    "default path length, so the lowest-ranked terms, including the individual alpha ",
    "frequency, reported above as a suggestive baseline signal, fell below that cap ",
    "and were not positively ruled out in that stratum.\n\n%s\n\n",
    "The aperiodic re-parameterisation reproduced this ",
    "ordering (suggested size %s), with the aperiodic exponent and offset entering at steps ",
    "%s and %s, well beyond the suggested size, so the $1/f$ background added no predictive ",
    "information the raw bands had missed. The same path cap left the adjusted beta power ",
    "and the specparam individual alpha frequency unreached, with the same caveat."),
    projpred_strata(), projpred_gender_caveat(),
    proj_size(projpath_raw), proj_rank(projpath_raw, "z_digit_span"),
    proj_rank(projpath_raw, "z_stroop"), proj_rank(projpath_raw, "z_asrt"),
    projpred_gender_result(),
    proj_size(projpath_aper), proj_rank(projpath_aper, "z_aperiodic_exponent"),
    proj_rank(projpath_aper, "z_aperiodic_offset")))
  if (!is.null(aper_loo_tbl)) {
    both_tie <- isTRUE(aper_loo_tie("pooled")) && isTRUE(aper_loo_tie("gender"))
    verdict <- if (both_tie) {
      "in each case smaller than twice its standard error"
    } else {
      sprintf("%s in the pooled stratum and %s for gender agreement",
              aper_loo("pooled", "verdict"), aper_loo("gender", "verdict"))
    }
    cat(sprintf(paste0(
      "\n\nThe two parameterisations were also compared by PSIS-LOO on the sample they ",
      "share. Because the specparam re-parameterisation is defined for %s of the %s participants, ",
      "the raw-band model was refit on that common subsample so the two are compared on ",
      "identical observations. The aperiodic-minus-raw difference in expected log predictive ",
      "density was %s (SE %s) in the pooled stratum and %s (SE %s) for gender agreement, %s ",
      "[@vehtari2017]. As noted in the Method, this leave-one-trial-out criterion cannot ",
      "detect participant-level predictors, so the near-zero difference is not read as ",
      "evidence that the two characterisations are equivalent.%s"),
      proj_n(projpath_aper), proj_n(projpath_raw),
      aper_loo("pooled", "elpd"), aper_loo("pooled", "se"),
      aper_loo("gender", "elpd"), aper_loo("gender", "se"), verdict,
      aper_kf_sentence()))
  }
  cat("\n")
} else {
  cat("*Projection-predictive selection results appear here once the projpred ",
      "solution-path CSVs are present in `paper_2_plasticity/results/`.*\n", sep = "")
}
```

```{r p2-derived}
#| results: asis
if (results_ready && !is.null(attainment_tbl) && !is.null(reliability_tbl)) {
  af <- function(m) { r <- attainment_tbl[attainment_tbl$model == m, ]
    if (nrow(r) == 1) sprintf("%.2f (95%% CrI [%.2f, %.2f])", r$median, r$ci_low, r$ci_high) else "NA" }
  # Reliabilities are bounded above by 1, so the leading zero is stripped, matching relv()
  # and predrel() elsewhere in the document rather than printing the same quantity two ways.
  rf <- function(m, e) { r <- reliability_tbl[reliability_tbl$model == m & reliability_tbl$effect == e, ]
    if (nrow(r) == 1) sub("^0", "", sprintf("%.2f", r$reliability)) else "NA" }
  cat(sprintf(paste0("Model-implied end-of-training attainment was %s pooled and %s for gender ",
              "agreement. The by-participant level at the pooled-frame mean session (the intercept) was ",
              "estimated with high reliability (pooled %s, gender-only %s). The learning-rate slope ",
              "was estimated less reliably (pooled %s, ",
              "gender-only %s), and because the gender-only rate is the least reliably estimated of ",
              "these quantities we do not interpret rate-based individual-difference claims from that ",
              "model.\n"),
      af(LES_P2_MODELS[["predictive"]]), af(LES_P2_MODELS[["predictive_gender"]]),
      rf(LES_P2_MODELS[["predictive"]], "Intercept"), rf(LES_P2_MODELS[["predictive_gender"]], "Intercept"),
      rf(LES_P2_MODELS[["predictive"]], "z_session_time"), rf(LES_P2_MODELS[["predictive_gender"]], "z_session_time")))
}
```

@tbl-reliability gives these outcome reliabilities in full, and
@tbl-predictor-reliability the internal consistency of the three baseline cognitive
indices. Both matter, because the association between a predictor and an outcome is
bounded by the reliability of each. Two of the three cognitive indices are moreover
difference scores, a form whose reliability is typically lower than that of its
components [@draheim2019; @hedge2018].

```{r}
#| label: fig-predictor-forest
#| fig-cap: "Part A Predictor Effects on Overall Accuracy and Learning Rate"
#| apa-note: "Points are posterior medians with 95% equal-tailed credible intervals for each baseline predictor's association with overall accuracy (the main effect, the association at the average session) and with learning rate (the predictor-by-session interaction), in the pooled and gender-agreement-only models. Estimates are on the logit scale: main effects are the change in log-odds of a correct response per SD of the predictor, interaction terms the change in the per-session slope per SD. The two panels share a common x axis so the families are directly comparable, and the rate-interaction terms carry deliberately tighter priors than the main effects. Predictors are ordered working memory, inhibitory control and statistical learning (for the Stroop-based inhibitory-control predictor, higher scores mean poorer inhibition), then the resting-state measures. Intervals excluding zero indicate effects whose direction is well determined."
#| fig-width: 6.5
#| fig-height: 6
if (results_ready) {
  preds <- c("z_digit_span", "z_stroop", "z_asrt", paste0("z_", LES_P2_RS_MEASURES))
  rate_terms <- paste0("z_session_time:", preds)
  fp <- post_tbl |>
    filter(model %in% LES_P2_MODELS[c("predictive", "predictive_gender")],
           parameter %in% c(preds, rate_terms)) |>
    mutate(Component = ifelse(grepl("^z_session_time:", parameter),
                              "Learning rate (x session)", "Overall accuracy (main effect)"),
           Predictor = factor(.pretty_term(gsub("^z_session_time:", "", parameter)),
                              levels = rev(.pretty_term(preds))),
           Model = ifelse(grepl("_gender$", model), "Gender only", "Pooled"))
  # Both are character, so without explicit levels the facets and the legend fall
  # alphabetically: the rate panel ahead of the main effects, and the gender-only
  # sensitivity model ahead of the primary pooled one, in both cases reversing the order
  # the caption and the surrounding text use.
  fp$Component <- factor(fp$Component,
                         levels = c("Overall accuracy (main effect)", "Learning rate (x session)"))
  fp$Model     <- factor(fp$Model, levels = c("Pooled", "Gender only"))
  if (nrow(fp)) {
    ggplot(fp, aes(x = median, y = Predictor, colour = Model, shape = Model)) +
      geom_vline(xintercept = 0, linetype = 2, colour = "grey50") +
      geom_pointrange(aes(xmin = ci_low, xmax = ci_high),
                      position = position_dodge(width = 0.5)) +
      facet_wrap(~ Component) +
      scale_colour_manual(values = c(Pooled = "#0072B2", `Gender only` = "#D55E00")) +
      scale_shape_manual(values = c(Pooled = 16, `Gender only` = 17)) +
      labs(x = "Posterior estimate (log-odds)", y = NULL, colour = NULL, shape = NULL) +
      les_theme(10) + theme(legend.position = "top",
                            plot.margin = margin(4, 8, 8, 8))
  }
}
```

```{r}
#| label: tbl-reliability
#| tbl-cap: "By-Participant Reliability of the Learning Trajectory"
#| apa-note: "Columns give the group standard deviation, mean posterior standard error, and reliability of the intercept (a participant's level at the mean session of the pooled standardisation frame) and of the session (rate) slope, for the pooled and gender-agreement-only models. The reliability column is the empirical-Bayes form, the variance of the shrunken by-participant posterior means over the model's between-participant variance. The group SD and mean posterior SE columns are shown for context and are not the terms of a formula."
if (!is.null(reliability_tbl)) {
  # Reliability is bounded above by 1, so its leading zero is stripped, matching relv()
  # and the Results prose. Group SD and the mean posterior SE are unbounded and keep theirs.
  .strip0 <- function(x) sub("^(-?)0\\.", "\\1.", formatC(x, format = "f", digits = 2))
  rel <- reliability_tbl |>
    transmute(Model = ifelse(grepl("_gender$", model), "Gender only", "Pooled"),
              Quantity = ifelse(effect == "Intercept", "Overall accuracy (intercept)", "Learning rate (session)"),
              `Group SD` = round(group_sd, 2),
              `Mean posterior SE` = round(mean_posterior_se, 2),
              Reliability = .strip0(reliability))
  # The reliability column is character (leading zero stripped), which kable would
  # otherwise left-align under a numeric header.
  knitr::kable(rel, align = c("l", "l", "r", "r", "r"))
}
```


## Part B: Change in executive function over training

Comparing the first and fifth sessions, working-memory span increased over training,
`r report_effect(.prepost_id("digit_span"), "session_prepost")`. The Stroop
interference effect
(`r report_effect(.prepost_id("stroop_interference"), "session_prepost")`) and the
statistical-learning effect
(`r report_effect(.prepost_id("asrt_learning"), "session_prepost")`) showed no
credible change. In the joint model that lets each index control the others, the pattern
pointed the same way. The joint-model digit-span change was
`r report_effect(LES_P2_MODELS[["cog_prepost_joint"]], "session_prepost")`, an interval that,
unlike the single-index model's, includes zero. The estimate is less precise in the
joint specification, which shares one residual standard deviation across the three
measures and adds a by-participant change slope. Relative to that change, the
pre/post shift was smaller for the Stroop interference effect
(`r report_effect(LES_P2_MODELS[["cog_prepost_joint"]], "session_prepost:measurestroop_interference")`)
and for the statistical-learning effect
(`r report_effect(LES_P2_MODELS[["cog_prepost_joint"]], "session_prepost:measureasrt_learning")`).
These specificity contrasts have credible intervals that include zero (probabilities of
direction `r pd_of(LES_P2_MODELS[["cog_prepost_joint"]], "session_prepost:measurestroop_interference")`%
and `r pd_of(LES_P2_MODELS[["cog_prepost_joint"]], "session_prepost:measureasrt_learning")`%), so
the evidence that the gain is working-memory-specific is suggestive without being decisive.
@fig-prepost shows these effects.

Two features of the comparison qualify it further. First, the digit-span-versus-Stroop
contrast is inflated by the polarity opposition described in the Method, since a generic
improvement pushes digit span up and Stroop interference down.
`r signaligned_sentence()` Second, the Stroop index has very low internal consistency here (see
*Limitations and future directions*), which leaves it a noisy control. In task units,
the digit-span change was from `r cog_mean("digit_span", 1)` items correct at Session
1 to `r cog_mean("digit_span", 5)` at Session 5, against
`r cog_mean("stroop_interference", 1)` to `r cog_mean("stroop_interference", 5)` ms of
Stroop interference and `r cog_mean("asrt_learning", 1)` to
`r cog_mean("asrt_learning", 5)` ms of ASRT learning.

```{r}
#| label: fig-prepost
#| fig-cap: "Pre/Post (Session 1 to 5) Change in the Executive-Function Indices"
#| apa-note: "Points are single-index session effects and, for the joint model, the digit-span change and the contrasts of each measure's change against digit span (the reference level). Filled points mark effects whose 95% credible interval excludes zero; open points do not."
#| fig-width: 6.5
#| fig-height: 3.5
if (results_ready) {
  pp <- post_tbl |>
    filter((model %in% c(.prepost_id("digit_span"),
                         .prepost_id("stroop_interference"),
                         .prepost_id("asrt_learning")) & parameter == "session_prepost") |
           (model == LES_P2_MODELS[["cog_prepost_joint"]] &
              parameter %in% c("session_prepost",
                               "session_prepost:measurestroop_interference",
                               "session_prepost:measureasrt_learning"))) |>
    mutate(Effect = dplyr::case_when(
      model == .prepost_id("digit_span") ~ "Digit span change",
      model == .prepost_id("stroop_interference") ~ "Stroop change",
      model == .prepost_id("asrt_learning") ~ "ASRT change",
      parameter == "session_prepost" ~ "Joint: digit-span change",
      parameter == "session_prepost:measurestroop_interference" ~ "Joint: Stroop vs digit span",
      parameter == "session_prepost:measureasrt_learning" ~ "Joint: ASRT vs digit span"),
      Block = ifelse(grepl("^Joint", Effect), "Joint model", "Single-index models"))
  pp$Block  <- factor(pp$Block, levels = c("Single-index models", "Joint model"))
  pp$Effect <- factor(pp$Effect, levels = c(
    "Joint: ASRT vs digit span", "Joint: Stroop vs digit span", "Joint: digit-span change",
    "ASRT change", "Stroop change", "Digit span change"))
  # Filled points mark effects whose 95% interval excludes zero, matching the
  # convention used in the companion paper's ERP forest.
  pp$Credible <- (pp$ci_low > 0) | (pp$ci_high < 0)
  ggplot(pp, aes(x = median, y = Effect)) +
    geom_vline(xintercept = 0, linetype = 2, colour = "grey50") +
    geom_pointrange(aes(xmin = ci_low, xmax = ci_high, shape = Credible)) +
    scale_shape_manual(values = c(`FALSE` = 1, `TRUE` = 16), guide = "none") +
    facet_grid(Block ~ ., scales = "free_y", space = "free_y") +
    labs(x = "Pre/post change (SD units)", y = NULL) +
    les_theme(10) + theme(plot.margin = margin(4, 8, 8, 8))
}
```

# Discussion

This study asked two questions: whether baseline cognitive and neural measures predict
how well adults learn a new grammar (Part A), and whether executive function changes over
the training period (Part B). The answer to the first is selective. Working memory was
associated with learners' overall level of accuracy, while inhibitory control,
statistical learning and the resting-state measures added little, and the predictors were
associated with that level, with no detectable association with the rate of change. The
answer to the second falls on the same capacity: working-memory span was the one index
that rose over training.

## Predictors of overall accuracy and learning rate

Working memory was the clearest cognitive predictor of overall accuracy. In the primary
pooled model its interval narrowly included zero, and in the gender-agreement model, the
one property sampled at every judgement session, the association was credible. This fits
the literature placing working memory as the best-evidenced domain-general cognitive
correlate of second-language outcomes [@linck2014]. Inhibitory control and statistical
learning, whose evidence base is weak or contested [@huensch2024; @linckweiss2015;
@chen2022; @zhou2024], added little. Digit span taps storage, whereas the executive or
complex span carries most of the working-memory--language association [@linck2014], and
it is the weaker of the phonological short-term memory tasks for vocabulary outcomes
[@kurokawa2026], so whether a complex-span measure would have shown a larger association
here is untested.

The baseline resting-state measures yielded no credible predictor, and for the four
low-frequency bands, which enter the model simultaneously and share most of their
variance, the joint model had little precision with which to resolve one (see Results).
Only the individual alpha
frequency hinted at a positive association with overall accuracy, in the direction the
cross-sectional literature would suggest (see Introduction), and only in the
gender-agreement model. It stands as a lead for confirmatory work,
and one to be held lightly, the more so as a larger test of that literature found the
frequency unrelated to working memory and general ability [@ociepka2022]. In the same
stratum, the projection-predictive search ranked the
individual alpha frequency last, and the estimate is prior-sensitive (see *Priors*). In a
small number of recordings, that measure is also a boundary value with no resolved peak
behind it (see *Cognitive and neural indices*). The null for the band powers contrasts
with reports that resting beta and low-gamma power predict L2 learning *rate* [@prat2016;
@prat2019]. The band-by-session interactions are in the model and none was credible,
though under weakly informative priors their intervals do not exclude effects of the
published size, and the divergence may reflect our posterior reduction where theirs was
right-temporoparietal, our regularising priors or our lower precision, which these data
do not separate. The low-gamma half of that precedent is untested here, because the gamma
predictor is not a valid measure (see *Cognitive and neural indices*). The aperiodic
re-parameterisation sharpens the neural null. Had the flat band coefficients merely
reflected the conflation of oscillatory power with the $1/f$ background [@donoghue2020],
de-confounding them should have uncovered signal, and it did not, a test that bears on
the adjusted alpha power, the one adjusted band with a fitted peak in most recordings.
The aperiodic exponent and offset entered the selection well beyond the suggested size,
and under participant-grouped cross-validation neither characterisation predicted a
held-out learner's trajectory detectably better than the other. Against a largely
cross-sectional literature relating the aperiodic exponent to cognitive function, in
which associations with specific abilities are inconsistent and the clearest signal is a
small one with general ability [@euler2024], this is a longitudinal test that such
designs cannot run, and it bounds the question without answering it. The exponent
carried no detectable incremental signal at this sample size, and the band-power nulls
are therefore not obviously attributable to periodic/aperiodic conflation, though a
comparison this blunt (see the minimum detectable difference in the Results) cannot
establish that the two characterisations are equivalent.

The predictors were associated with learners' overall *level* of accuracy, with no
detectable association with the *rate*, an asymmetry with a precedent in the study that
established the resting-EEG result, where working memory, executive function and fluid
intelligence were likewise uncorrelated with learning rate [@prat2016]. The one
interaction approaching credibility, the statistical-learning-by-session term in the
gender-agreement model, and its adjusted-alpha counterpart in the aperiodic refit of that
stratum, come from the model whose rate is the least reliably estimated quantity here
(reliability `r relv(LES_P2_MODELS[["predictive_gender"]], "z_session_time")`), so we do
not interpret either. The asymmetry is not an artefact of priors that shrink the rate
terms towards zero, since it held when those terms were re-fit under weakly informative
priors (see Results).

The overall pattern is consistent with a circumscribed advantage. Had a broad aptitude
driven learning, it should also have appeared in the statistical-learning index, which is
reliably measured and was assessed and modelled identically, and it did not. In the
event, the signal was concentrated in working memory, and the projection-predictive
selection retained nothing beyond it. The pattern does not, however, exclude a broad
advantage, since the statistical-learning interval still admits an association as large
as the one attributed to working memory (see *Limitations and future
directions*).`r contrast_sentence()` The neural and inhibitory-control comparisons carry
less weight in that argument, since the low-frequency bands are weakly identified, the
gamma predictor is not a valid measure and the Stroop index is too unreliable to have
registered an association (see *Limitations and future directions*), so the contrast the
data can draw is between working memory and statistical learning, the two indices
measured well enough to be compared.

Because the by-participant intercept and slope were estimated inside one multilevel
model, the uncertainty that a two-stage slope analysis would discard is carried into the
predictor estimates [@gelman2005; @humberg2024]. The by-participant level was estimated
more reliably than the learning rate
(`r relv(LES_P2_MODELS[["predictive"]], "Intercept")` against
`r relv(LES_P2_MODELS[["predictive"]], "z_session_time")` pooled, and
`r relv(LES_P2_MODELS[["predictive_gender"]], "z_session_time")` for the gender-only
rate), and that asymmetry bears on how aptitude findings indexed by a per-learner slope
should be read. If the rate is in general the less reliably estimated quantity, as it was
here even under trial-level joint modelling, the most favourable case, then slope-indexed
correlations will be attenuated, and the coefficients unstable, in proportion to that
unreliability, and disagreement among them is to be expected [@hedge2018; @humberg2024].

One feature of the outcome qualifies the words *learning* and *rate* throughout. Mean
accuracy was highest at each property's first session and lower at its last, and the
session term was credibly negative for the reference property in the pooled model and
within gender agreement (see *Descriptive overview*). The companion paper reports the same fall in the same judgements
and shows that discrimination between grammatical and ungrammatical items is present, and
for gender agreement improving, where the level falls. Three features of the design bear
on it. Properties accumulate across sessions, so a later judgement is made against a
grammar with more manipulated properties. The final session follows about four months
without exposure, over which some forgetting is expected, and the model codes it one step
after Session 4. A drift in response criterion cannot be separated from either, because
no condition holds task demands constant while sessions advance. The predictors are
therefore associated with the level at which a learner sustains accuracy across
training, and the rate terms are slopes of that trajectory, whichever way it runs.
Separating acquisition from decline would need a retention test scored apart from the
training sessions.

## Change in executive function over training

A behavioural change also emerged over training. Working-memory span rose between the first and fifth
sessions, while inhibitory control and statistical learning did not. This comparison lacks a
no-training control group, so on its own a pre/post gain is open to the familiar threats of
an uncontrolled retest: practice effects, regression to the mean, maturation and expectancy
[@shadish2002; @calamia2012; @barnett2005]. One feature of the analysis bounds those
threats, since of three indices measured and retested identically only working memory
moved. In the joint model, the pre/post change
was larger for digit span than for the co-administered statistical-learning measure
(probability of direction
`r pd_of(LES_P2_MODELS[["cog_prepost_joint"]], "session_prepost:measureasrt_learning")`%, and
`r pd_of(.signaligned_id, "session_prepost:measureasrt_learning")`%
with the Stroop index sign-aligned). The digit-span-versus-Stroop contrast is weaker
evidence, since aligning the Stroop polarity lowered its probability of direction from
`r pd_of(LES_P2_MODELS[["cog_prepost_joint"]], "session_prepost:measurestroop_interference")`%
to `r pd_of(.signaligned_id, "session_prepost:measurestroop_interference")`% (see Results)
and the index is in any case too unreliable here to serve as a comparator. The
specificity argument therefore rests on the statistical-learning comparator, as in Part
A. That contrast does not clear the credibility threshold, since its 95% interval includes
zero, and the joint-model digit-span change does not either, so the concentration remains
provisional. The threats of an uncontrolled retest act broadly across a battery, so a
change concentrated in a single index is not what they predict. The same working-memory
measure was also the one baseline predictor of overall accuracy, so the predictive and
pre/post analyses converge on it, but a practice effect on digit span would arise whether
or not digit span predicted learning, so the convergence does not discriminate a
training-related account from one. None of this licenses a causal claim. Practice effects are not
uniform across tasks [@scharfen2018], and even active control conditions do not fully
rule out differential expectancy [@boot2013; @simons2016], so attributing the gain to the
multilingual training specifically would require a matched no-training group. In an
11-week randomised trial in older adults, language learning produced no
working-memory gain over a relaxation control [@berggren2020], and a three-week
randomised course found no gain on updating or working memory over a waitlist
[@grossmann2023]. The result is consistent with a training-related,
working-memory-specific change, and no more than that.

## Limitations and future directions

The first limitation is precision. Although the models are fitted to tens of thousands of
trials, the predictor effects are *between*-participant quantities, so their effective sample
size is the number of participants. Only
`r proj_n(projpath_raw)` participants contribute between-participant variance to the
raw-band model and `r proj_n(projpath_aper)` to the aperiodic variant, and 18
between-participant coefficients (nine main effects and nine interactions with time) are
estimated from them. That is why even the clearest predictor effect carries a wide
interval. Effects that do
clear a 95% interval at this sample size are also, on average, overestimates of their true
magnitude [@gelmancarlin2014], so the working-memory coefficient should be read as
evidence of a positive association whose size remains uncalibrated.

A second limitation is that the study is silent on *neural* plasticity. Resting-state EEG
was recorded only at baseline, so the rs-EEG measures bear only on the prediction
question. Establishing whether intensive multilingual training reshapes resting-state
activity would need a matched post-training recording, of the kind that has shown resting
network organisation shifting after language training in older adults [@kliesch2022]. A
change in working-memory span is a change in what the task measures, and says nothing
about whether the resting rhythms underlying that capacity have shifted.

A third limitation is the reliability of the predictors themselves, which bounds any
association they can enter into. We estimated the internal consistency of each baseline
index by permutation split-half with a Spearman--Brown correction. The result separates two measures that are usually discussed together as
difference scores [@hedge2018]. Statistical learning is estimated highly reliably
(`r predrel("asrt_learning")`) and working-memory span adequately so
(`r predrel("digit_span")`), but Stroop interference is not. Its split-half reliability is
`r predrel("stroop_interference")`, which is close to zero. That is in keeping with the
finding, across 24 published inhibition tasks, that trial noise is about eight times the
size of the true variability between individuals [@rouder2023]. Almost none of its
between-participant variance is systematic, so the inhibitory-control null in Part A is
uninformative however large the true effect, and the index is a noisy internal control in
Part B. The statistical-learning null carries no such excuse. The ASRT is estimated highly
reliably, so its near-zero coefficient cannot be put down to measurement error. The ASRT
figure is in line with the acceptable internal consistency the task reaches in a large
sample when learning scores are computed from reaction times over
`r task_spec("asrt", "n_blocks")` blocks, although such estimates move with
pre-processing choices [@farkas2024]. Within-session split-half estimates also run well
above the test-retest reliability of serial reaction-time learning [@oliveira2023], so
the figure bears on the precision of the baseline index and does not by itself establish
it as a stable trait. Nor can the null be attributed with any confidence to timing. In
the two training studies that tracked their predictors by session, statistical learning
contributed most in the earliest exposure [@chen2022; @kenanidis2026]. A contribution
that fades across sessions would register here as a statistical-learning-by-session
interaction, and those interactions were weak (see *Predictors of overall accuracy and
learning rate*), although a linear term would not capture a contribution confined to the
first grammar session alone. Even so, a null detection leaves the absence of an
association undemonstrated. Under a prior centred above zero, the pooled estimate is
`r report_effect(LES_P2_MODELS[["predictive"]], "z_asrt")`. The interval's upper bound is
close to the working-memory median we treat as our clearest effect, so these data cannot
exclude a statistical-learning association of the size they attribute to working memory.
What they establish is that no such association was detected on a well-measured index,
which is why it serves as the comparator the specificity argument rests on where the
neural and inhibitory-control nulls cannot.

A further scope condition is that no language-specific aptitude test was administered,
so the contrast our data can draw is between a single circumscribed capacity and a broad
domain-general advantage. Whether an aptitude measure of the Modern Language Aptitude
Test (MLAT) and LLAMA tradition [@li2015] would outpredict working memory here is an open
question.

Finally, the artificial-language design trades ecological validity for control. Because
the grammar is new to every participant, prior exposure cannot confound the learning
curve. The cost is that the three properties enter at different points and
differ in difficulty, so the pooled trajectory averages over learners who are at different
stages of different problems. The pooled analysis is therefore best read alongside the
gender-only one, restricted to the single property sampled from the first grammar session
onwards, which the predictor pattern survived. Two further scope conditions apply. The
outcome is a single grammaticality-judgement task, so nothing here speaks to production or
to comprehension under time pressure, and the sample is first-language-Norwegian,
second-language-English bilinguals (see *Transparency and preregistration*), to whom the
findings are confined.

## Conclusion

Two lines of evidence point to working memory. It was the one baseline index that
predicted learners' overall accuracy on the new grammar, and the one index that rose
across the training period. Statistical learning, measured at least as reliably, did
neither, and the inhibitory-control measure was too unreliable here to test either way.
Both results therefore converge on working memory. Neither establishes that the capacity
is a specific one, since the statistical-learning interval admits an association of
similar size and the evidence that the pre/post gain is confined to working memory stops
short of being decisive.

# Data and code availability

The data that support the findings of this study are openly available in the Open
Science Framework at <https://osf.io/tq7vy>. All extraction, modelling and reporting
scripts, the computational environment specification and the machine-written summary
tables that populate this manuscript are archived in the same repository, together with
the online supplementary material. The raw electroencephalographic recordings are held
there too, being too large to distribute through an ordinary version-control repository.
The paper's analysis code and other open materials are also available at
<https://github.com/pablobernabeu/LESS-cognitive-predictors>. The
supplement holds
the baseline predictor distributions and the resting-state spectra, which are descriptive
companions to the analyses reported here and are not needed to interpret them. The
environment specification has two parts, because no single artefact covers the whole of it:
an `renv` lockfile pinning the R packages of the fitting library, snapshotted on the cluster,
and a provenance record written at fit time. The compiler and the optimisation pin described
in the supplementary material belong to neither, being properties of the cluster module,
which is why they are stated in prose. The provenance record covers the software versions in force
when the Part A predictor models were fitted, so those fits can be traced to the environment
that produced them. The pre/post models, the projection-predictive selection and the
cross-validation comparisons run from the same environment, but they write no record of
their own.
`r prov_record_sentence()`

# Supplementary material

For supplementary material accompanying this paper, visit *[URL assigned by the
journal]*. The supplement comprises the trimming rules and minimum trial counts of the
cognitive indices, the acquisition and preprocessing of the resting-state EEG, the
transmission analysis behind the gamma-band judgement, the settings of the aperiodic
decomposition, the predictive comparison and the projection in full, the prior
specification in full, the computational environment, Figure S1 (distributions of the
baseline predictors) and Figure S2 (grand-average resting-state power spectral density
over the eight occipito-parietal electrodes).

{{< pagebreak >}}

# Convergence diagnostics {#apx-convergence-diagnostics-p2}

Every model whose posterior estimates are reported in the main text was assessed against the same convergence criteria. @tbl-convergence gives the realised per-model values behind the summary in the Results. The common-sample refits behind the cross-validation comparisons are not listed. They exist only to score the two parameterisations on identical observations, carry no reported coefficient and do not pass through the diagnostic artefact.

```{r}
#| label: tbl-convergence
#| tbl-cap: "Convergence Diagnostics for Every Model Whose Posterior Estimates Are Reported"
#| apa-note: "Columns give the maximum rank-normalised split $\\widehat{R}$ and the minimum bulk and tail effective sample sizes (ESS), each taken over all parameters. Part A models come first, then the Part B pre/post models. The weak-prior rows are the prior-sensitivity refits described under Priors."
if (!is.null(conv_tbl) && requireNamespace("knitr", quietly = TRUE)) {
  # Row labels: the CSV carries pipeline identifiers, which are not reader-facing, and it
  # is written in whatever order the summaries job produced. Map to prose names and order
  # rows to follow the manuscript (Part A, then Part B). An identifier with no entry in
  # the map keeps its stripped form rather than being dropped, so a future model still
  # appears.
  .cml <- c(
    predictive_trajectory                     = "Part A: pooled",
    predictive_trajectory_weakprior           = "Part A: pooled (weak priors)",
    predictive_trajectory_gender              = "Part A: gender agreement",
    predictive_trajectory_gender_weakprior    = "Part A: gender agreement (weak priors)",
    predictive_trajectory_gender_iafres       = "Part A: gender agreement, excluding the two floor-pinned IAF recordings",
    predictive_trajectory_aperiodic           = "Part A: pooled, aperiodic",
    predictive_trajectory_aperiodic_gender    = "Part A: gender agreement, aperiodic",
    cognition_prepost_digit_span              = "Part B: digit span",
    cognition_prepost_stroop_interference     = "Part B: Stroop interference",
    cognition_prepost_asrt_learning           = "Part B: ASRT learning",
    cognition_prepost_joint                   = "Part B: joint across the three indices",
    cognition_prepost_joint_signaligned       = "Part B: joint across the three indices, sign-aligned")
  conv_tbl <- conv_tbl[order(match(sub("^p2_", "", conv_tbl$model), names(.cml)),
                             conv_tbl$model), , drop = FALSE]
  # Reordering leaves the source row numbers as row names, which kable would print as an
  # unlabelled leading column of pipeline row indices.
  rownames(conv_tbl) <- NULL
  ct <- conv_tbl |>
    dplyr::transmute(Model = ifelse(is.na(.cml[sub("^p2_", "", model)]),
                                    sub("^p2_", "", model),
                                    unname(.cml[sub("^p2_", "", model)])),
                     `Maximum $\\widehat{R}$` = round(max_rhat, 3),
                     `Minimum bulk ESS` = round(min_ess_bulk),
                     `Minimum tail ESS` = round(min_ess_tail),
                     Divergences = n_divergent,
                     Passed = ifelse(passed, "yes", "no"))
  # Drop a column that carries no information. Every model here has zero divergent
  # transitions and every one passes, so those two columns were a wall of "0" and "yes"
  # restating the sentence above the table, and they were what pushed the header past the
  # measure. They are dropped only when they are genuinely constant, so a future run in
  # which a model diverges or fails still shows it rather than hiding it.
  if (all(!is.na(ct$Divergences)) && all(ct$Divergences == 0)) ct$Divergences <- NULL
  if (all(ct$Passed == "yes"))                ct$Passed      <- NULL
  knitr::kable(ct)
} else {
  cat("*Convergence diagnostics will appear here once the models are fitted.*")
}
```

# Reliability of the baseline predictors {#apx-predictor-reliability}

@tbl-predictor-reliability gives the split-half internal consistency of the three
baseline cognitive indices, each index having been split `r predrel_splits()` times. The
values, and what they imply for the inhibitory-control null, are discussed under
*Limitations and future directions*.

```{r}
#| label: tbl-predictor-reliability
#| tbl-cap: "Internal-Consistency Reliability of the Three Baseline Cognitive Predictors at Session 1"
#| apa-note: "Reliability is estimated by permutation split-half: each participant's trials are repeatedly split in half, the index is recomputed on each half, the halves are correlated across participants, and the mean correlation is corrected to full test length by the Spearman--Brown formula. Each index was split the number of times given in the text. Two of the three indices (Stroop interference, ASRT learning) are difference scores."
if (!is.null(predrel_tbl) && requireNamespace("knitr", quietly = TRUE)) {
  .prl <- c(digit_span = "Working memory (digit span)",
            stroop_interference = "Inhibitory control (Stroop interference)",
            asrt_learning = "Statistical learning (ASRT)")
  # Correlations and reliabilities are bounded by 1, so their leading zeros are stripped,
  # matching predrel() and relv() in the running prose.
  .strip0 <- function(x) sub("^(-?)0\\.", "\\1.", formatC(x, format = "f", digits = 2))
  pr <- predrel_tbl |>
    dplyr::transmute(Predictor = unname(.prl[predictor]),
                     `Participants` = n_participants,
                     `Mean half-test $r$` = .strip0(r_half),
                     `Reliability (Spearman--Brown)` = .strip0(reliability_sb),
                     `Random splits` = n_splits)
  # A column that is constant across rows conveys nothing and crowds the predictor
  # labels; the count then goes in the note (see predrel_splits()).
  if (length(unique(pr$`Random splits`)) == 1) pr$`Random splits` <- NULL
  knitr::kable(pr, align = c("l", rep("r", ncol(pr) - 1)))
} else {
  cat("*Predictor reliability will appear once `scripts/11_extract_predictor_reliability.R` has run.*")
}
```
